### 2020-05-04 paper/paper.tex
Updating

\documentclass[11pt]{article}
\usepackage{cite}

\begin{document}

\title{Sammy's Sproj}
\author{Sammy Furr}
\date{Today}
\maketitle

Blablabla said Test~\cite{DBLP:journals/corr/OCallahanJFHNP17}.

\bibliography{sprojbib}{}
\bibliographystyle{plain}
\end{document}

### 2020-05-07 paper/midway.tex
Midway board paper is underway!

\documentclass[12pt, twocolumn]{article}
\usepackage{cite}

\author{Sammy Furr}
\title{Teaching Debugging Collaboratively---Midway Report}
\date{\today}

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

\section{Introduction}

Debugging is invaluable in writing and understanding code, yet it is
rarely formally taught.  We typically teach students programming
structures, concepts, and languages, but we leave them to learn the
tools they use to code by themselves.  This approach often works
well---the programmer's choice of editor is \textit{very} personal,
students figure out how to configure an individualized workflow.
Perhaps because debuggers are tools, they often get lumped into the
``teach yourself'' category.  Unlike editors or reference guides
however, effectively using a debugger requires a set of high-level,
platform agnostic, teachable skills.  Teaching these skills is
effective, and translates into better, faster, debugging and
programming.\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}\par

Teaching debugging is particuarly important in

\section{The Value of Teaching Debugging}

\section{Debugger: Low Level}

\section{Debugger: High Level}

\section{Next Steps}

\pagebreak
\bibliographystyle{acm}
\bibliography{sprojbib}{}
\end{document}

### 2020-05-07 paper/midway.tex
Rounding out the inro

@@
 \documentclass[12pt, twocolumn]{article}
+\usepackage{multicol}
 \usepackage{cite}
@@
 
-Teaching debugging is particuarly important in
+The use of a debugger is particuarly important in a low-level
+programming classe, which is often the first time that students
+encounter concepts like assembly instructions and memory addresses.
+Debuggers such as GCC offer incredibly powerful tools to step through
+and inspect running code.  Sadly, low level debugging tools are often
+less than intuitive, and provide little to no means for collaboration.
+Powerful tools such as RR exist that enable collaborative ``record and
+replay'' debugging\cite{DBLP:journals/corr/OCallahanJFHNP17}, but they
+lack tools to visualize address spaces and control flow of programs.
+Research shows that visualization of these previously unexplored
+spaces aids students taking low-level or systems programming
+classes\cite{10.1145/3328778.3366894}.\par
+
+The necessity for accessible, collaborative tools for teaching all
+aspects of computer science, not the least debugging, has been
+exemplified by the current COVID-19 crisis.
 

### 2020-05-08 paper/midway.tex
Intro finished!

@@
 
-The use of a debugger is particuarly important in a low-level
-programming classe, which is often the first time that students
+The use of a debugger is particularly important in a low-level
+programming classes, which is often the first time that students
 encounter concepts like assembly instructions and memory addresses.
@@
 aspects of computer science, not the least debugging, has been
-exemplified by the current COVID-19 crisis.
+exemplified by the current COVID-19 crisis.  Many students don't have
+access to a high powered computer (or even a computer at all) to run
+tools that allow for collaboration and replay by capturing the entire
+state of a virtual machine\cite{10.1145/2843859.2843867}.  Though the
+limitations of interacting with a debugger through a phone or
+low-powered laptop seem daunting, I believe they are actually
+constraints that can help better shape a visualized, collaborative
+approach.\par
+
+For my senior project, I want to develop a front-end interface to RR
+that enables collaborative debugging and process visualization on a
+varitey of non-traditional platforms.
 

### 2020-05-08 paper/midway.tex
We're half way there on the value section!

@@
-\documentclass[12pt, twocolumn]{article}
+\documentclass[12pt]{article}
 \usepackage{multicol}
@@
 Debugging is invaluable in writing and understanding code, yet it is
-rarely formally taught.  We typically teach students programming
-structures, concepts, and languages, but we leave them to learn the
-tools they use to code by themselves.  This approach often works
-well---the programmer's choice of editor is \textit{very} personal,
-students figure out how to configure an individualized workflow.
-Perhaps because debuggers are tools, they often get lumped into the
-``teach yourself'' category.  Unlike editors or reference guides
-however, effectively using a debugger requires a set of high-level,
-platform agnostic, teachable skills.  Teaching these skills is
-effective, and translates into better, faster, debugging and
-programming.\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}\par
+rarely formally taught\cite{doi:10.1080/08993400802114581}.  We
+typically teach students programming structures, concepts, and
+languages, but we leave them to learn the tools they use to code by
+themselves.  This approach often works well---the programmer's choice
+of editor is \textit{very} personal, students figure out how to
+configure an individualized workflow.  Perhaps because debuggers are
+tools, they often get lumped into the ``teach yourself'' category.
+Unlike editors or reference guides however, effectively using a
+debugger requires a set of high-level, platform agnostic, teachable
+skills.  Teaching these skills is effective, and translates into
+better, faster, debugging and
+programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.\par
 
@@
 that enables collaborative debugging and process visualization on a
-varitey of non-traditional platforms.
+variety of non-traditional platforms.
 
 \section{The Value of Teaching Debugging}
+
+Before creating a platform for teaching debugging it's of course
+important to make sure that teaching debugging is actually valuable.
+There is not a lot research specifically into the efficacy of teaching
+debugging for computer science students, despite a recent rise in the
+inclusion of debugging in ``computational thinking''
+curriculums\cite{10.1145/3361721.3361724}.  These curriculums attempt
+to teach skills in computer science classes that are useful in other
+subject areas: the UK's computer science curriculum considers
+debugging an essential ``transferable skill''\cite{10.1145/2602484}.
+Though there seems to be confidence that the problem-solving
+techniques used in debugging are widely applicable, I am more
+interested in whether systematically teaching debugging actually
+benefits computer science students.  Michaeli and Romeike conducted a
+good, albeit somewhat small, study on the efficacy of teaching a
+systematic debugging process to K12 students.  They found that
+students who have been taught a specific debugging framework performed
+better in debugging tests and were more confident in their own
+debugging skills\cite{10.1145/3361721.3361724}.  Their result is
+positive evidence towards the efficacy of teaching debugging, though
+their research doesn't include college or university students.\par
+
+As Michaeli and Romeike point out, there is a lack of research into
+the value of teaching debugging in higher education.
+
+Research into how to best teach debugging is self-admittedly sparse.
+Chan et al. allow that ``in general research on how to improve
+debugging is sporadic''---an observation that leads them to research a
+framework to reduce the complexity of teaching
+debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
+they split debugging knowledge into 5 categories: Domain, System,
+Procedural, Strategic, and Experiential.  They then review different
+debugging tools and teaching aids---from those that involve writing
+code to games---and map tools to the knowledge areas they seek to
+address.  After an evaluation of a host of different tools, they claim
+a few significant faults in current debugging teaching platforms.  The
+two of which I seek to address follow:
+\begin{itemize}
+\item A lack of tools addressing system knowledge (an understanding of
+  the program to be debugged).
+\item A lack of back-tracing ability/coverage.
+\end{itemize}
 

### 2020-05-08 paper/midway.tex
Section 2 done!

@@
 As Michaeli and Romeike point out, there is a lack of research into
-the value of teaching debugging in higher education.
+the value of teaching debugging in higher education.  None of the
+research they found placed much focus on explicitly teaching
+debugging.  Chmiel and Loui studied whether students who were provided
+with debugging tools and frameworks performed better on tests or spent
+less time on assignments than those who were
+not\cite{10.1145/971300.971310}.  They weren't able to find conclusive
+evidence towards better performance on tests or assignments, though
+they did find that students in the treatment group felt more confident
+in their debugging abilities.  Unfortunately their study didn't
+involve extended explicit teaching of debugging--use of the tools was
+voluntary, and variations in the students' individual abilities made
+the data difficult to evaluate.\par
 
-Research into how to best teach debugging is self-admittedly sparse.
-Chan et al. allow that ``in general research on how to improve
-debugging is sporadic''---an observation that leads them to research a
-framework to reduce the complexity of teaching
+Despite the lack of higher education-specific research, I feel that
+the value of teaching debugging has still been demonstrated.  The
+research discussed all finds that K12 and college students alike
+commonly resort to sporadic debugging techniques when beginning to
+learn.  Given this, I see no reason why the conclusive and positive
+research toward the benefit of explicitly teaching debugging to K12
+students should not apply to their collegiate counterparts.\par
+
+Research into how to best teach debugging is also self-admittedly
+sparse.  Chan et al. allow that ``in general research on how to
+improve debugging is sporadic''---an observation that leads them to
+research a framework to reduce the complexity of teaching
 debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
@@
 address.  After an evaluation of a host of different tools, they claim
-a few significant faults in current debugging teaching platforms.  The
-two of which I seek to address follow:
-\begin{itemize}
+to find a few significant faults in current debugging teaching
+platforms.  The two of which I seek to address follow:
+
+\begin{enumerate}
+\item A lack of back-tracing ability/coverage.
 \item A lack of tools addressing system knowledge (an understanding of
   the program to be debugged).
-\item A lack of back-tracing ability/coverage.
-\end{itemize}
+\end{enumerate}
+
+Basing my collaborative debugging platform on RR solves the problem of
+back-tracing ability and coverage.  RR is a ``time traveling''
+addition to GDB, that allows for unlimited recording and replaying
+code execution deterministically, both backwards and
+forwards\cite{DBLP:journals/corr/OCallahanJFHNP17}.  By integrating
+the debugger with tools to visualize the execution and code space in
+an accessible way, I hope to address the issue of system knowledge.
+Though the system knowledge gained may be primarily applicable to low
+level programming, skills such as visualizing code execution and
+memory space are applicable to all types of coding.  These solutions
+are discussed in depth in the following sections.
 

### 2020-05-10 paper/midway.tex
some background

@@
 \usepackage{cite}
+\usepackage{listings}
+
+\lstset{basicstyle=\ttfamily}
 
@@
 
+\subsection{Overview of RR}
+
+Time traveling debugging---stepping backwards and forwards through a
+program's execution---is supported naively by GDB through the use of
+commands such as \lstinline{reverse-continue} and
+\lstinline{reverse-stepi}.\cite{gdbman} What RR brings to the table is
+the ability to deterministically record and replay the execution of
+even very complex programs.  This allows single users to capture and
+debug events that may not happen every time a program runs and revisit
+debugging later.  Essentially for my planned collaborative debugger,
+execution can be replayed and debugged even \textit{on a different
+  computer} than the computer that recorded the original execution.
+In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
+that rely on capturing the entire state of of a virtual machine to
+replay execution, RR records and replays faster, produces far smaller
+files, and doesn't force execution inside of a
+VM.\cite{DBLP:journals/corr/OCallahanJFHNP17}
+
+\subsection{Limitations of RR}
+
+RR comes with two major limitations: it is only compatible with the
+Linux kernel, and it's deterministic recording and replay relies on a
+feature that is only found on modern \textit{Intel} x86 CPUs.  For my
+purposes the first limitation is fairly irrelevant--I expect students
+to be connecting to a central server in order to execute and replay
+code, and most systems classes are taught in Linux.  The second
+limitation poses more of a problem for instruction.  Because recording
+and replay is limited to x86 CPUs, classes cannot use a debugger based
+on RR to teach ARM assembly, which is becoming increasingly popular.
+Though this isn't a limitation that solutions like the aforementioned
+full-vm recorder PANDA face, RR still maintains the advantage of being
+a debugger first and foremost.  Students can learn RR, GDB, and
+debugging techniques simultaneously.  Platforms that require entire VM
+replay come with far greater performance and instructional overhead.
+Because of RR's ease of use and tight integration with GDB, I feel it
+is the best choice of platform to base a collaborative debugger on.
+
 \section{Debugger: High Level}

### 2020-05-11 paper/midway.tex
Adding some more expo

@@
 \usepackage{listings}
+\usepackage{tikz}
 
@@
 encounter concepts like assembly instructions and memory addresses.
-Debuggers such as GCC offer incredibly powerful tools to step through
+Debuggers such as GDB offer incredibly powerful tools to step through
 and inspect running code.  Sadly, low level debugging tools are often
@@
 
-\section{Debugger: Low Level}
+\section{Collaborative Debugger}
 
-\subsection{Overview of RR}
+\subsection{Overview}
+
+\begin{tikzpicture} [node distance=2em, start chain=going right,]
+
+\end{tikzpicture}
+
+\subsection{Low Level}
+
+\paragraph{Overview of RR}
 
@@
 
-\subsection{Limitations of RR}
+\paragraph{Limitations of RR}
 
@@
 Though this isn't a limitation that solutions like the aforementioned
-full-vm recorder PANDA face, RR still maintains the advantage of being
+full-VM recorder PANDA face, RR still maintains the advantage of being
 a debugger first and foremost.  Students can learn RR, GDB, and
@@
 
-\section{Debugger: High Level}
+\subsection{High Level}
+
+There are many existing tools to visualize parts of the process space
+during execution, such as Python Tutor's C Tutor\cite{pythontutor}.
+Likewise, there are excellent platforms like Matt Godbolt's Compiler
+Explorer\cite{godbolt} that allow students to easily see the mapping
+of C code to assembly.  Unfortunately there are no student friendly
+tools such as these that integrate fully with GDB.  There are plugins
+that enhance the features of GDB to offer some further visualization,
+like GEF\cite{gef}, but these are typically designed with experts in
+mind.  While I feel that it is essential for students to learn
+real-world tools like GDB, research shows that visualization tools
+designed with teaching in mind significantly increase students'
+understanding\cite{10.1145/3328778.3366894}.  I want to marry the
+teaching-oriented interface of tools like C Tutor and the Compiler
+Explorer with the real-world power of GDB and RR.
+
+\subsection{Mid Level}
 

### 2020-05-11 paper/midway.tex
in which we waste time making charts

@@
 \usepackage{tikz}
+\usetikzlibrary{positioning}
 
@@
 
-\begin{tikzpicture} [node distance=2em, start chain=going right,]
+\begin{tikzpicture}
+  [node distance = 2em, auto, every node/.style={align=left,
+    rectangle, minimum size=2em, rounded corners, fill=blue!20, inner sep=1em},
+  ml/.style = {fill=green!30}]
 
+  \node (ll) 
+  {\textbf{Low Level:}\\
+    RR\\
+    gdb
+  };
+  \node (ml) [ml, right =of ll]
+  {\textbf{Mid Level:}\\
+    RR\\
+    gdb
+  };
+  \node (hl) [right =of ml]
+  {\textbf{High Level:}\\
+    Process Vis Tools\\
+    Interface to RR/GDB
+  };
+  \draw[<->] (ll) -- (ml);
+  \draw[<->] (ml) -- (hl);
 \end{tikzpicture}

### 2020-05-12 paper/midway.tex
pre-proof

@@
 in their debugging abilities.  Unfortunately their study didn't
-involve extended explicit teaching of debugging--use of the tools was
+involve extended explicit teaching of debugging---use of the tools was
 voluntary, and variations in the students' individual abilities made
@@
 
-\subsection{Overview}
+% \subsection{Overview}
 
-\begin{tikzpicture}
-  [node distance = 2em, auto, every node/.style={align=left,
-    rectangle, minimum size=2em, rounded corners, fill=blue!20, inner sep=1em},
-  ml/.style = {fill=green!30}]
-
-  \node (ll) 
-  {\textbf{Low Level:}\\
-    RR\\
-    gdb
-  };
-  \node (ml) [ml, right =of ll]
-  {\textbf{Mid Level:}\\
-    RR\\
-    gdb
-  };
-  \node (hl) [right =of ml]
-  {\textbf{High Level:}\\
-    Process Vis Tools\\
-    Interface to RR/GDB
-  };
-  \draw[<->] (ll) -- (ml);
-  \draw[<->] (ml) -- (hl);
-\end{tikzpicture}
-
-\subsection{Low Level}
+\subsection{Preexisting Debugging Tools}
 
@@
 feature that is only found on modern \textit{Intel} x86 CPUs.  For my
-purposes the first limitation is fairly irrelevant--I expect students
+purposes the first limitation is fairly irrelevant---I expect students
 to be connecting to a central server in order to execute and replay
@@
 
-\subsection{High Level}
+\subsection{User Facing Tools}
 
@@
 
-\subsection{Mid Level}
+% \subsection{Mid Level}
 
 \section{Next Steps}
+
+Further research is needed before I start on the user-facing part of
+the collaborative debugger.  I feel that it's important to pick and
+design visualization tools based on research into what works most
+effectively for students.  I also want the visualization tools to be
+as general as possible.  Though I plan to use them to visualize x86
+programs running in RR, tools like those to visualize registers could
+easily be written more generally, allowing them to represent ARM or
+MIPS registers.  In order to make the debugger as flexible as possible
+I plan on splitting the package into three parts:
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}
+    [node distance = 2em, auto, every node/.style={align=left,
+      rectangle, minimum size=2em, rounded corners, fill=blue!20, inner
+      sep=1em}, ml/.style = {fill=green!30}]
+
+    \node (ll) 
+    {\textbf{Debugger Facing:}\\
+      hooks for specific\\
+      RR/GDB output\\
+      session management
+    };
+    \node (ml) [ml, right =of ll]
+    {\textbf{Middleware:}\\
+      server code\\
+      user management\\
+      and storage
+    };
+    \node (hl) [right =of ml]
+    {\textbf{User Facing:}\\
+      process vis tools\\
+      interface to\\
+      RR/GDB
+    };
+    \draw[<->] (ll) -- (ml);
+    \draw[<->] (ml) -- (hl);
+  \end{tikzpicture}
+  \caption{\textbf{Blue:} RR specific software. \textbf{Green:} general collaborative
+    debugging management}
+\end{figure}
+
+I'm not planning to directly extend RR or GDB to add collaborative
+debugging features.  Instead, I plan to write code that will hook into
+specific RR/GDB outputs and manage a single RR/GDB session.  This code
+will interface with server code that manages multiple users and
+coordinates collaborative debugging.  The server will pass on commands
+to user facing visualization tools, which will translate output from
+the debugger to visualizations and provide an interface to RR/GDB.
+Segmenting the code in this way allows the addition of different
+debugger backends or different front-end vis tools without changing
+the code that handles collaboration.  This makes the debugger
+extensible as methods and tools evolve.\par
+
+I plan to continue research and begin working on the software over the
+summer.  By the beginning of the Fall semester I hope to have made
+significant progress on the interface for a single user to RR/GDB.
 

### 2020-10-26 containers/rr-server/pytest_gdb.py
Working rr with sockets on DO managed kubernetes

@@ -1,0 +1,10 @@
+from pygdbmi.gdbcontroller import GdbController
+from pprint import pprint
+
+# Start gdb process
+gdbmi = GdbController(gdb_args =["--interpreter=mi2"])
+print(gdbmi.get_subprocess_cmd())  # print actual command run as subprocess
+response = gdbmi.write('-break-insert main')
+response = gdbmi.write('-exec-run')
+response = gdbmi.write('-exec-continue')
+pprint(response)

### 2020-10-26 containers/rr-server/server.py

@@ -1,0 +1,31 @@
+import eventlet
+import socketio
+import time
+
+sio = socketio.Server(cors_allowed_origins='http://localhost:3000', sync_mode='eventlet')
+app = socketio.WSGIApp(sio)
+
+@sio.on('rr command')
+def on_rr_command(sid, data):
+    print('Received rr command:')
+    data['sid'] = sid
+    print(data)
+    print('Emitting...')
+    sio.emit('rr command', data)
+
+@sio.on('rr response')
+def on_rr_response(sid, data):
+    print('Recieved rr response:')
+    print(data)
+    print('Emitting...')
+    sio.emit('rr response', data)
+
+@sio.event
+def connect(sid, environ):
+    print('connect ', sid)
+
+@sio.event
+def disconnect(sid):
+    print('disconnect ', sid)
+
+eventlet.wsgi.server(eventlet.listen(('', 8000)), app)

### 2020-10-26 containers/rr/rr-translation.py

@@ -1,0 +1,81 @@
+from pygdbmi.gdbcontroller import GdbController
+from pprint import pprint
+import socketio
+
+class RRInterface:
+    def __init__(self, session=None):
+        self.gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
+        self.init_message = self.console_output(self.get_rr_response())
+        self.timeline = []
+
+    def exited(self, response):
+        # Unsure if this will always be the second to last message, works for now.
+        return response[-2]['message'] == 'thread-group-exited'
+
+    def end(self, response):
+        last = response[-1]['message']
+        return last == 'stopped' or last == 'done'
+
+    def get_rr_response(self):
+        return self.gdbmi.get_gdb_response()
+
+    def get_full_rr_response(self, command):
+        response = self.gdbmi.write(command)
+        while(not(self.end(response) or self.exited(response))):
+            response.extend(self.get_rr_response())
+        return response
+
+    def console_output(self, response):
+        output = []
+        for r in response:
+            if r['type'] == 'console' or r['type'] == 'output':
+                output.append(r['payload'])
+        return ''.join(output).replace("\\n", "\n").replace("\\t", "\t")
+    
+    def write(self, command):
+        self.timeline.append(command)
+        return self.console_output(self.get_full_rr_response(command))
+
+# Start gdb process
+#gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
+
+sio = socketio.Client()
+rri = RRInterface()
+
+print(rri.init_message)
+
+@sio.event
+def connect():
+    print("I'm connected!")
+    
+@sio.on('rr command')
+def on_rr_command(data):
+    print('Recieved rr command:')
+    print(data)
+    print('Passing command to rr')
+    response = rri.write(data['command'])
+    print('Emitting...')
+    sio.emit('rr response', {'response': response, 'from': data['sid'], 'command': data['command']})
+    
+    
+@sio.event
+def connect_error():
+    print("The connection failed!")
+
+@sio.event
+def disconnect():
+    print("I'm disconnected!")
+
+sio.connect('http://rr-host-load-balancer:8000')
+
+# while (command := input('(rr) ')) != 'exit':
+#     print(rri.write(command))
+
+# print(rri.timeline)
+
+# rri2 = RRInterface()
+# print(rri2.init_message)
+
+# for command in rri.timeline:
+#     print('(rr) ' + command)
+#     print(rri2.write(command))

### 2020-10-26 containers/rr/startup.sh

@@ -1,0 +1,4 @@
+gcc -g -o cat cat.c
+rr record ./cat cat.c test.py
+hostname -I
+python3 rr-translation.py

### 2020-10-26 frontend/client/pages/index.js

@@ -1,0 +1,52 @@
+import Head from 'next/head'
+import React from 'react';
+import ReactDOM from 'react-dom';
+
+import socketIOClient from "socket.io-client";
+import styles from '../styles/Home.module.css'
+
+class RRForm extends React.Component {
+    constructor(props) {
+	super(props);
+	this.state = {value: ''};
+	this.socket = props.socket;
+	this.handleChange = this.handleChange.bind(this);
+	this.handleSubmit = this.handleSubmit.bind(this);
+    }
+    
+    handleChange(event) {
+	this.setState({value: event.target.value});
+    }
+
+    handleSubmit(event) {
+	this.socket.emit('rr command', {'command': this.state.value})
+	event.preventDefault();
+    }
+
+    render() {
+	return (
+	    <form onSubmit={this.handleSubmit}>
+	      <label>
+		Command:
+		<input type="text" value={this.state.value} onChange={this.handleChange} />
+	      </label>
+	      <input type="submit" value="send" />
+	    </form>
+	);
+    }
+}
+
+const socket = socketIOClient('http://157.230.64.84:8000');
+
+socket.on('rr response', (data) => {
+    if(data['from'] !== socket.id){
+	console.log(data['command'])
+    };
+    console.log(data['response']);
+});
+
+export default function Home() {
+  return (
+	  <RRForm socket={socket}/>
+  )
+}

### 2020-10-27 paper/paper.tex
Some service updates, adding updated paper... TODO: fix this nutty file structure we have going on

@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
+\usepackage{multicol}
 \usepackage{cite}
+\usepackage{listings}
+\usepackage{tikz}
+\usepackage[doublespacing]{setspace}
+
+\usepackage{titlesec}
+\newcommand{\sectionbreak}{\clearpage}
+\usetikzlibrary{positioning}
+
+\lstset{basicstyle=\ttfamily}
+
+\author{Sammy Furr}
+\title{The Development of a Collaborative Tool to Teach Debugging}
+\date{\today}
 
@@
 
-\title{Sammy's Sproj}
-\author{Sammy Furr}
-\date{Today}
+\begin{titlepage}
 \maketitle
+\end{titlepage}
 
-Blablabla said Test~\cite{DBLP:journals/corr/OCallahanJFHNP17}.
+\begin{abstract}
+TODO: write an abstract
+\end{abstract}
 
+\tableofcontents
+\pagebreak
+
+\section{Introduction}
+\subsection{Motivation}
+
+Debugging is invaluable in writing and understanding code, yet it is
+rarely formally taught\cite{doi:10.1080/08993400802114581}.  We
+typically teach students programming structures, concepts, and
+languages, but leave them to learn the tools they use to write code by
+themselves.  This approach often works well---a programmer's choice of
+tools is often \textit{very} personal and students figure out how to
+configure an individualized workflow.  Perhaps because debuggers are
+tools, students are often expected to learn them with minimal
+guidance.  Unlike editors or reference guides however, effectively
+using a debugger requires a set of high-level, platform agnostic,
+teachable skills.  Teaching these skills is effective, and translates
+into better, faster, debugging and
+programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.
+
+\subsubsection{The Value of Teaching Debugging}
+
+There is an unfortunate lack of research specifically into the
+efficacy of teaching debugging for computer science students, despite
+a recent rise in the inclusion of debugging in ``computational
+thinking'' curriculums\cite{10.1145/3361721.3361724}.  These
+curriculums attempt to teach skills in computer science classes that
+translate into other subject areas: the UK's computer science
+curriculum considers debugging an essential ``transferable
+skill''\cite{10.1145/2602484}.\par There seems to be confidence that
+the problem-solving techniques used in debugging are widely
+applicable, but of greater interest to computer science teachers is
+whether teaching debugging directly benefits student programmers.
+Michaeli and Romeike conducted a good, albeit somewhat small, study on
+the efficacy of teaching a systematic debugging process to K12
+students.  They found that students who have been taught a specific
+debugging framework performed better in debugging tests and were more
+confident in their own debugging skills\cite{10.1145/3361721.3361724}.
+Their result is positive evidence towards the efficacy of teaching
+debugging, though it doesn't include college or university
+students.\par
+
+As Michaeli and Romeike point out, there is a lack of research into
+the value of teaching debugging in higher education.  None of the
+research these authors found placed much focus on explicitly teaching
+debugging.  Chmiel and Loui studied whether students who were provided
+with debugging tools and frameworks performed better on tests or spent
+less time on assignments than those who were
+not\cite{10.1145/971300.971310}.  Though this research wasn't able to
+find conclusive evidence towards better performance on tests or
+assignments, it did find that students in the treatment group felt
+more confident in their debugging abilities.  Unfortunately Chmiel and
+Loui's study didn't involve extended explicit teaching of
+debugging---use of the tools was voluntary, and variations in the
+students' individual abilities made the data difficult to
+evaluate.\par
+
+Though there is a lack of higher-education research, the value of
+teaching debugging is still demonstrable.  The research discussed all
+finds that K-12 and college students alike commonly resort to sporadic
+debugging techniques when beginning to learn.  Since this pattern of
+behavior that explicitly teaching debugging corrects exists in college
+as well as in K-12 students, it seems logical that the benefit of
+explicitly teaching debugging to K-12 students should be realized
+equally by their collegiate counterparts.\par
+
+\subsubsection{Methods for Teaching Debugging}
+
+Similarly to research on the value of teaching debugging, research
+into how to best teach debugging is self-admittedly sparse.  Chan et
+al. allow that ``in general research on how to improve debugging is
+sporadic''---an observation that leads them to research a framework to
+reduce the complexity of teaching
+debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
+they split debugging knowledge into 5 categories: \textit{Domain},
+\textit{System}, \textit{Procedural}, \textit{Strategic}, and
+\textit{Experiential}.  They then review different debugging tools and
+teaching aids---from those that involve writing code to games---and
+map tools to the knowledge areas they seek to address.  After an
+evaluation of a host of different tools, they claim to find a few
+significant faults in current debugging teaching platforms. The
+primary two which this project seeks to address are as follows:
+
+\begin{enumerate}
+\item A lack of back-tracing ability/coverage.
+\item A lack of tools addressing system knowledge (an understanding of
+  the program to be debugged).
+\end{enumerate}
+
+\subsubsection{The Value of Collaborative Programming}
+
+TODO: find some basic research that backs up the claim that
+collaborative/pair programming is worthwhile.
+
+\subsubsection{Tools that Enable Collaborative Programming}
+
+TODO: write about glitch, repl.it, etc.
+
+\subsubsection{Collaborative Programming and Teaching Debugging}
+
+Debuggers exist at an intersection of tools and skills similar to
+programming languages themselves. By becoming familiar with a specific
+debugger, students learn techniques and paradigms necessary to use all
+debuggers effectively.
+
+\subsection{Project Components}
+
+\subsubsection{Mozilla's rr}
+
+\paragraph{Overview}
+
+rr is ``a lightweight tool for recording, replaying and debugging
+execution of applications''.\cite{rr-repo} rr allows a programmer to
+record the execution of a program on any compatible machine and replay
+the execution later.  This enhances GDB's ability to ``time-travel''
+when debugging, using commands such as \lstinline{reverse-continue}
+and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
+forwards through a program's execution.  Through a novel encapsulation
+of the execution space, rr is able to deterministically record and
+replay the execution of syscalls and other process behavior that
+differs run-to-run.  This is invaluable when trying to debug behavior
+that is not entirely dependent on the code being debugged.  A typical
+workflow in rr consists of recording an inexplicable error, replaying
+execution to find the area in which the error occurs, and then
+narrowing in on the bug not by re-running the entire program, but by
+progressing back and forth through execution in the problem area.
+
+\paragraph{Limitations}
+
+In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
+that rely on capturing the entire state of of a virtual machine to
+replay execution, rr records and replays faster, produces far smaller
+files, and doesn't force execution inside of a
+VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The tradeoff for these
+benefits are two major system limitations: rr is only compatible with
+the Linux kernel, and it's deterministic recording and replay relies
+on a feature that is only found on modern \textit{Intel} x86 CPUs.
+These limitations influenced the development of this project as a
+webapp similar to existing tools for collaborative programming.
+Luckily, the speed and size benefits of rr lend themselves well to
+non-local execution.
+
+\subsubsection{Kubernetes and Docker}
+
+TODO: Write about k8s and docker.
+
+\subsubsection{React}
+
+TODO: Write about react.
+
+\subsubsection{Xterm.js}
+
+TODO: Write about Xterm.js if I end up using it (which seems
+overwhelmingly likely ATM).
+
+\subsection{Project Overview}
+
+TODO: a brief overview linking aspects of the project to background
+research discussed above.
+
+\section{Design}
+
+TODO: the big piece of writing, detailing components of the project,
+the system as a whole, and motivations for design choices.
+
+\section{Next Steps}
+
+TODO: write about additional/necessary unimplemented features (ATM
+this is looking like auth stuff, unless OAuth2 proves to be less of a
+pain in the ass then when I've used it before in Javascript.  Maybe
+some of the vis stuff as well, though I'd really like to get the
+basics done here.)  Write about desire to actually do research and see
+whether the tool helps students learn debugging.
+
+\pagebreak
+\bibliographystyle{acm}
 \bibliography{sprojbib}{}
-\bibliographystyle{plain}
 \end{document}

### 2020-10-27 containers/rr-server/server.py

@@ -2,7 +2,7 @@ import eventlet
 import socketio
 import time
 
-sio = socketio.Server(cors_allowed_origins='http://localhost:3000', sync_mode='eventlet')
+sio = socketio.Server(cors_allowed_origins='*', sync_mode='eventlet')
 app = socketio.WSGIApp(sio)
 
 @sio.on('rr command')

### 2020-11-02 backend/client2.py
Adding container management... still TODO clean up god awful repo structure

@@ -1,0 +1,35 @@
+import socketio
+import asyncio
+
+# standard Python
+sio = socketio.Client()
+
+@sio.event
+def connect():
+    print("I'm connected!")
+    
+@sio.on('pod_created')
+def on_pod_created(data):
+    print(data)
+
+@sio.on('pod_delete')
+def on_pod_deleted(data):
+    print(data)
+    
+@sio.event
+def connect_error():
+    print("The connection failed!")
+
+@sio.event
+def disconnect():
+    print("I'm disconnected!")
+
+sio.connect('http://localhost:8001')
+
+# while (command := input('(rr) ')) != 'exit':
+#     sio.call('rr command', {'command': command}, timeout=60)
+
+# sio.emit('pod_create', {'names': ['sari', 'sammy']})
+sio.emit('pod_delete', {'channel': '77562'})
+# sio.emit('pod_delete', {'channel': '39640'})
+sio.wait()

### 2020-11-02 containers/controller/controller.py

@@ -1,0 +1,36 @@
+import sanic
+import socketio
+from podmanager import TranslationPodManager
+
+sio = socketio.AsyncServer(async_mode='sanic')
+app = sanic.Sanic(name='controller')
+sio.attach(app)
+tpm = TranslationPodManager(url='165.227.252.45', port=27017)
+
+# Receives a username from server requesting a pod be created
+# Returns a channel id that corresponds to the pod
+@sio.on('pod_create')
+async def on_pod_create(sid, data):
+    try:
+        channel = await tpm.create_pod(data['names'])
+        await sio.emit('pod_created', {'names': data['names'], 'channel': channel})
+    except:
+        print("error")
+
+@sio.on('pod_delete')
+async def on_pod_delete(sid, data):
+    try:
+        await tpm.delete_pod(data['channel'])
+        await sio.emit('pod_deleted', {'channel': data['channel']})
+    except:
+        print("error")
+        
+@sio.event
+async def connect(sid, environ):
+    print('connect ', sid)
+
+@sio.event
+async def disconnect(sid):
+    print('disconnect ', sid)
+
+app.run(port='8001')

### 2020-11-02 containers/controller/database.py

@@ -1,0 +1,93 @@
+from pymongo import MongoClient, errors
+from pprint import pprint
+import random
+
+class NoOpenChannelError(Exception):
+    """Raised when all channels are in use."""
+    pass
+
+class NoMatchingChannelError(Exception):
+    """Raised when no pod matches a provided channel"""
+    pass
+
+class DuplicateUserError(Exception):
+    """Raised when a duplicated username is attempted to be used"""
+    pass
+
+class DatabaseController:
+    """Provides interface to mongodb database"""
+    def __init__(self, url='localhost', port=27017):
+        self.client = MongoClient(url, port)
+        self.db = self.client.debugger
+        self.users = self.db.users
+        self.pods = self.db.pods
+        self.min_channel = 10000
+        self.max_channel = 99999
+        self.max_channels = self.max_channel - self.min_channel
+
+
+    # User Methods
+
+    def add_user(self, name):
+        """Given a name, attempt to add the user into the database"""
+        return str(self.users.insert_one({'name': name,
+                                          'pods': []}).inserted_id)
+
+    def delete_user(self, uid):
+        """Given a uid, delete a user"""
+        self.pods.update_many({}, {'$pull': {'users': uid}})
+        self.users.delete_one({'_id': uid})
+
+    def get_userids_by_name(self, names):
+        """Given a list of names, retrieve the mongodb ids for the users"""
+        users = self.users.find({'name': {'$in': names}}, {"_id": 1})
+        if users:
+            return [u['_id'] for u in users]
+        else:
+            return []
+
+    # Pod Methods
+    
+    def add_pod(self, uids=[]):
+        """Given a list of uids, creates a pod and returns a channel to use for communicating with the pod"""
+        # While there are open channels, try to create a new pod with
+        # an open channel.  Obviously in real use there should be a
+        # better method for generating channel numbers, but given the
+        # number of current users the likelyhood of a collision is
+        # extreemly low using 5 digit numbers
+        while self.pods.count_documents({'channel': {'$not': {'$eq': None}}}) <= self.max_channels:
+            try:
+                channel = str(random.randint(self.min_channel, self.max_channel))
+                pid = self.pods.insert_one({'channel': channel}).inserted_id
+                if uids:
+                    self.link_pod_to_users(pid, uids)
+                return channel
+            except errors.DuplicateKeyError:
+                pass
+        raise NoOpenChannelError
+    
[…]

### 2020-11-02 containers/controller/podmanager.py

@@ -1,0 +1,66 @@
+from kubernetes import client, config
+import database
+import asyncio
+import time
+
+class TranslationPodManager:
+    def __init__(self, url='localhost', port=27017):
+        self.dbc = database.DatabaseController(url, port)
+        
+        config.load_kube_config()
+        self.v1 = client.CoreV1Api()
+        
+    async def create_pod(self, names):
+        dep = {'apiVersion': 'v1',
+               'kind': 'Pod',
+               'metadata': {'labels': {'purpose': 'translate-rr'}},
+               'spec': {'containers': [{'image': 'registry.digitalocean.com/sproj/rr:translation',
+                                        'name': 'rr-test-container',
+                                        'securityContext': {'capabilities': {'add': ['SYS_PTRACE']}}}],
+                        'restartPolicy': 'OnFailure'}}
+        channel = self.dbc.add_pod(self.dbc.get_userids_by_name(names))
+        dep['metadata']['name'] = channel
+        resp = self.v1.create_namespaced_pod(
+            body=dep, namespace='default')
+        print("Pod created.  status='%s'" % resp.metadata.name)
+        return channel
+
+
+    async def delete_pod(self, channel):
+        """Given a channel, delete a pod"""
+        self.v1.delete_namespaced_pod(channel, 'default')
+        # Wait for double pod's maximum termination time (30 seconds).
+        # Waiting for termination eliminates the scenario where a pod
+        # is created with the same channel as a pod that is still
+        # terminating.  Creation of the new pod would fail.
+        await asyncio.sleep(60)
+        self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
+        # while True:
+        #     try:
+        #         print(channel)
+        #         print(self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses[0].state.terminated)
+        #         await asyncio.sleep(1)
+        #     except:
+        #         self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
+        #         break
+
+    def link_pod_to_users(self, channel, names):
+        pod = self.dbc.get_pod_by_channel(channel)
+        uids = self.dbc.get_userids_by_name(names)
+        self.dbc.link_pod_to_users(pod, uids)
+
+    def unlink_pod_from_users(self, channel, names):
+        pod = self.dbc.get_pod_by_channel(channel)
+        uids = self.dbc.get_userids_by_name(names)
+        self.dbc.unlink_pod_from_users(pod, uids)
+
[…]

### 2020-11-03 backend/client2.py
frontend stuff

@@ -12,7 +12,7 @@ def connect():
 def on_pod_created(data):
     print(data)
 
-@sio.on('pod_delete')
+@sio.on('pod_deleted')
 def on_pod_deleted(data):
     print(data)
     
@@ -29,7 +29,8 @@ sio.connect('http://localhost:8001')
 # while (command := input('(rr) ')) != 'exit':
 #     sio.call('rr command', {'command': command}, timeout=60)
 
-# sio.emit('pod_create', {'names': ['sari', 'sammy']})
-sio.emit('pod_delete', {'channel': '77562'})
+sio.emit('pod_create', {'names': ['sari', 'sammy']})
+sio.emit('pod_delete', {'channel': '60852'})
+sio.emit('pod_delete', {'channel': '28483'})
 # sio.emit('pod_delete', {'channel': '39640'})
 sio.wait()

### 2020-11-03 containers/controller/controller.py

@@ -11,19 +11,13 @@ tpm = TranslationPodManager(url='165.227.252.45', port=27017)
 # Returns a channel id that corresponds to the pod
 @sio.on('pod_create')
 async def on_pod_create(sid, data):
-    try:
-        channel = await tpm.create_pod(data['names'])
-        await sio.emit('pod_created', {'names': data['names'], 'channel': channel})
-    except:
-        print("error")
+    channel = await tpm.create_pod(data['names'])
+    await sio.emit('pod_created', {'names': data['names'], 'channel': channel})
 
 @sio.on('pod_delete')
 async def on_pod_delete(sid, data):
-    try:
-        await tpm.delete_pod(data['channel'])
-        await sio.emit('pod_deleted', {'channel': data['channel']})
-    except:
-        print("error")
+    await tpm.delete_pod(data['channel'])
+    await sio.emit('pod_deleted', {'channel': data['channel']})
         
 @sio.event
 async def connect(sid, environ):

### 2020-11-03 containers/controller/podmanager.py

@@ -23,6 +23,13 @@ class TranslationPodManager:
         resp = self.v1.create_namespaced_pod(
             body=dep, namespace='default')
         print("Pod created.  status='%s'" % resp.metadata.name)
+        # Wait to return the channel until the pod is live and read to
+        # recieve incomming communication
+        status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
+        # 
+        while status == None or status[0].state.running == None:
+            await asyncio.sleep(1)
+            status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
         return channel
 
 
@@ -35,6 +42,7 @@ class TranslationPodManager:
         # terminating.  Creation of the new pod would fail.
         await asyncio.sleep(60)
         self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
+        return
         # while True:
         #     try:
         #         print(channel)

### 2020-11-03 containers/database/database.py

@@ -24,13 +24,13 @@ class DatabaseController:
         self.min_channel = 10000
         self.max_channel = 99999
         self.max_channels = self.max_channel - self.min_channel
-
+        
     # User Methods
 
     def add_user(self, name):
         """Given a name, attempt to add the user into the database"""
         return str(self.users.insert_one({'name': name,
-                                          'pods': []}).inserted_id)
+                                          'pods': []}).inserted_id)        
 
     def delete_user(self, uid):
         """Given a uid, delete a user"""

### 2020-11-03 containers/frontend/react-flask-app/api/api.py

@@ -1,0 +1,29 @@
+import database
+from flask import Flask, request
+
+dbc = database.DatabaseController(url='165.227.252.45', port=27017)
+
+app = Flask(__name__)
+
+@app.route('/login', methods=['POST'])
+def login():
+    try:
+        name = request.get_json()['name']
+        print(name)
+        user = dbc.get_userids_by_name([name])
+        print(user)
+        if user == []:
+            return {'name': None}
+        return {'name': name}
+    except:
+        return {'error': 'Internal Error'}
+
+@app.route('/channel', methods=['POST'])
+def channel():
+    try:
+        channel = request.get_json()['channel']
+        print(channel)
+        dbc.get_pod_by_channel(channel)
+        return {'channel': channel}
+    except:
+        return {'error': 'Internal Error'}

### 2020-11-03 containers/frontend/react-flask-app/api/database.py

@@ -1,0 +1,1 @@
+../../../database/database.py

### 2020-11-03 containers/frontend/react-flask-app/src/App.js

@@ -1,0 +1,192 @@
+import React from 'react';
+import socketIOClient from "socket.io-client";
+import './App.css';
+
+async function post_request (url, data) {
+    const response = await fetch(url, {
+	method: 'POST',
+	body: JSON.stringify(data),
+	headers: new Headers({
+	    "content-type": "application/json"
+	})
+    }).then(res => res.json());
+    return response;
+}
+
+class RRForm extends React.Component {
+    constructor(props) {
+	super(props);
+	this.state = {value: ''};
+	this.socket = props.socket;
+	this.handleChange = this.handleChange.bind(this);
+	this.handleSubmit = this.handleSubmit.bind(this);
+    }
+    
+    handleChange(event) {
+	this.setState({value: event.target.value});
+    }
+
+    handleSubmit(event) {
+	this.socket.emit('rr command', {'command': this.state.value})
+	event.preventDefault();
+    }
+
+    render() {
+	return (
+	    <form onSubmit={this.handleSubmit}>
+	      <label>
+		Command:
+		<input type="text" value={this.state.value} onChange={this.handleChange} />
+	      </label>
+	      <input type="submit" value="send" />
+	    </form>
+	);
+    }
+}
+
+class LoginForm extends React.Component {
+   constructor(props) {
+	super(props);
+	this.state = {name: ''};
+	this.handleChange = this.handleChange.bind(this);
+	this.handleSubmit = this.handleSubmit.bind(this);
+   }
+
+    handleChange(event) {
+	this.setState({name: event.target.value});
+    }
+
+    handleSubmit(event) {
+	console.log(this.state.name);
+	post_request('/login', {name: this.state.name}).then(data => {
+	    console.log(data);
+	    if (data.name != null){
+		this.props.onLogin(data.name);
+	    }
+	});
+	event.preventDefault();
+    }
+
+    render() {
+	return (
+	    <form onSubmit={this.handleSubmit}>
+	      <label>
+		Login:
+		<input type="text" value={this.state.name} onChange={this.handleChange} />
+	      </label>
+	      <input type="submit" value="login" />
+	    </form>
+	);
+    }
+}
+
+class ChannelForm extends React.Component {
+   constructor(props) {
+	super(props);
+	this.state = {channel: ''};
+	this.handleChange = this.handleChange.bind(this);
[…]

### 2020-11-15 containers/frontend/react-flask-app/src/App.js
Some basic CSS

@@ -3,6 +3,7 @@ import socketIOClient from 'socket.io-client';
 import './App.css';
 import XTerm, {Terminal} from 'react-xterm';
 import 'xterm/css/xterm.css';
+import WebFont from 'webfontloader';
 
 async function post_request (url, data) {
     // Send a post request to the server, which will forward it to our
@@ -42,7 +43,7 @@ class RRForm extends React.Component {
 		Command:
 		<input type="text" value={this.state.value} onChange={this.handleChange} />
 	      </label>
-	      <input type="submit" value="send" />
+	      <input type="submit" value="send"><i className="material-icons">face</i></input>
 	    </form>
 	);
     }
@@ -76,12 +77,12 @@ class LoginForm extends React.Component {
 
     render() {
 	return (
-	    <form onSubmit={this.handleSubmit}>
-	      <label>
-		Login or Register:
-		<input type="text" value={this.state.name} onChange={this.handleChange} />
-	      </label>
-	      <input type="submit" value="login/register" />
+	    <form className='simple-form' onSubmit={this.handleSubmit}>
+	      <label>Login or Register:</label>
+	      <div className='simple-submit'>
+		<input type="text" value={this.state.name} onChange={this.handleChange}/>
+		<button type="submit"><i className="material-icons">arrow_forward_ios</i></button>
+	      </div>
 	    </form>
 	);
     }
@@ -116,12 +117,12 @@ class ChannelForm extends React.Component {
 
     render() {
 	return (
-	    <form onSubmit={this.handleSubmit}>
-	      <label>
-		Session ID:
+	    <form className='simple-form' onSubmit={this.handleSubmit}>
+	      <label>Session ID:</label>
+	      <div className='simple-submit'>
 		<input type="text" value={this.state.channel} onChange={this.handleChange} />
-	      </label>
-	      <input type="submit" value="join" />
+		<button type="submit"><i className="material-icons">arrow_forward_ios</i></button>
+	      </div>
 	    </form>
 	);
     }
@@ -219,22 +220,22 @@ class RRTerm extends React.Component {
 
 	this.term.on('key', (key, ev) => {
 	    const printable = !ev.altKey && !ev.ctrlKey && !ev.metaKey;
-//	    console.log(ev.keyCode);
-	    if (ev.keyCode === 13) {
[…]

### 2020-11-18 paper/paper.tex
paper changes

@@
 debuggers effectively.
+\pagebreak
 
@@
 
-\subsubsection{Mozilla's rr}
+\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+\tikzset{every picture/.style={baselineskip=1pt}}
+\tikzset{every picture/.style={/utils/exec={\ttfamily}}}
+
+\begin{figure}[h!]
+  \begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
+    % uncomment if require: \path (0,494); %set diagram left start at 0, and has height of 494
+
+    % Straight Lines [id:da7736831842096191] 
+    \draw [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=1.5]    (380,110) -- (400,110) ;
+    % Shape: Rectangle [id:dp04619309000609384] 
+    \draw  [color={rgb, 255:red, 136; green, 136; blue, 136 }  ,draw opacity=1 ][fill={rgb, 255:red, 136; green, 136; blue, 136 }  ,fill opacity=1 ] (420,174) ..
+controls (420,171.79) and (421.79,170) ..
+(424,170) -- (546,170) ..
+controls (548.21,170) and (550,171.79) ..
+(550,174) -- (550,236) ..
+controls (550,238.21) and (548.21,240) ..
+(546,240) -- (424,240) ..
+controls (421.79,240) and (420,238.21) ..
+(420,236) -- cycle ;
+    % Shape: Rectangle [id:dp1228490870487785] 
+    \draw   (410,74) ..
+controls (410,71.79) and (411.79,70) ..
+(414,70) -- (536,70) ..
+controls (538.21,70) and (540,71.79) ..
+(540,74) -- (540,156) ..
+controls (540,158.21) and (538.21,160) ..
+(536,160) -- (414,160) ..
+controls (411.79,160) and (410,158.21) ..
+(410,156) -- cycle ;
+    % Shape: Rectangle [id:dp17147478083974088] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (410,74) ..
+controls (410,71.79) and (411.79,70) ..
+(414,70) -- (536,70) ..
+controls (538.21,70) and (540,71.79) ..
+(540,74) -- (540,136) ..
+controls (540,138.21) and (538.21,140) ..
+(536,140) -- (414,140) ..
+controls (411.79,140) and (410,138.21) ..
+(410,136) -- cycle ;
+    % Shape: Rectangle [id:dp08735176678895862] 
+    \draw   (410,184) ..
+controls (410,181.79) and (411.79,180) ..
+(414,180) -- (536,180) ..
+controls (538.21,180) and (540,181.79) ..
+(540,184) -- (540,266) ..
+controls (540,268.21) and (538.21,270) ..
+(536,270) -- (414,270) ..
+controls (411.79,270) and (410,268.21) ..
+(410,266) -- cycle ;
+    % Shape: Rectangle [id:dp9595297060154465] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (410,184) ..
+controls (410,181.79) and (411.79,180) ..
+(414,180) -- (536,180) ..
+controls (538.21,180) and (540,181.79) ..
+(540,184) -- (540,246) ..
+controls (540,248.21) and (538.21,250) ..
+(536,250) -- (414,250) ..
+controls (411.79,250) and (410,248.21) ..
+(410,246) -- cycle ;
+    % Shape: Rectangle [id:dp8231056072369544] 
+    \draw   (410,294) ..
+controls (410,291.79) and (411.79,290) ..
+(414,290) -- (536,290) ..
+controls (538.21,290) and (540,291.79) ..
+(540,294) -- (540,376) ..
+controls (540,378.21) and (538.21,380) ..
+(536,380) -- (414,380) ..
+controls (411.79,380) and (410,378.21) ..
+(410,376) -- cycle ;
+    % Shape: Rectangle [id:dp6367028620504607] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (410,294) ..
+controls (410,291.79) and (411.79,290) ..
+(414,290) -- (536,290) ..
+controls (538.21,290) and (540,291.79) ..
+(540,294) -- (540,356) ..
+controls (540,358.21) and (538.21,360) ..
+(536,360) -- (414,360) ..
+controls (411.79,360) and (410,358.21) ..
+(410,356) -- cycle ;
+    % Shape: Rectangle [id:dp8426243961763082] 
+    \draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=1.5]  (40,74) ..
+controls (40,71.79) and (41.79,70) ..
+(44,70) -- (106,70) ..
+controls (108.21,70) and (110,71.79) ..
+(110,74) -- (110,106) ..
+controls (110,108.21) and (108.21,110) ..
+(106,110) -- (44,110) ..
+controls (41.79,110) and (40,108.21) ..
+(40,106) -- cycle ;
+    % Shape: Rectangle [id:dp3822373593536188] 
+    \draw  [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]  (40,134) ..
+controls (40,131.79) and (41.79,130) ..
+(44,130) -- (106,130) ..
+controls (108.21,130) and (110,131.79) ..
+(110,134) -- (110,166) ..
+controls (110,168.21) and (108.21,170) ..
+(106,170) -- (44,170) ..
+controls (41.79,170) and (40,168.21) ..
+(40,166) -- cycle ;
+    % Straight Lines [id:da30880840083209404] 
+    \draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (120,150) -- (210,150) ;
+    % Shape: Rectangle [id:dp5669106087613988] 
+    \draw  [line width=1.5]  (220,74) ..
+controls (220,71.79) and (221.79,70) ..
+(224,70) -- (366,70) ..
+controls (368.21,70) and (370,71.79) ..
+(370,74) -- (370,166) ..
+controls (370,168.21) and (368.21,170) ..
+(366,170) -- (224,170) ..
+controls (221.79,170) and (220,168.21) ..
+(220,166) -- cycle ;
+    % Straight Lines [id:da24300582818436978] 
+    \draw [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=1.5]    (120,90) -- (152.5,90) -- (210,90) ;
+    % Straight Lines [id:da9215040164115669] 
+    \draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (380,130) -- (400,130) ;
+    % Shape: Rectangle [id:dp3848745776815077] 
+    \draw  [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]  (40,224) ..
+controls (40,221.79) and (41.79,220) ..
+(44,220) -- (176,220) ..
+controls (178.21,220) and (180,221.79) ..
+(180,224) -- (180,286) ..
+controls (180,288.21) and (178.21,290) ..
+(176,290) -- (44,290) ..
+controls (41.79,290) and (40,288.21) ..
+(40,286) -- cycle ;
+    % Shape: Rectangle [id:dp9909839356430334] 
+    \draw  [line width=1.5]  (220,224) ..
+controls (220,221.79) and (221.79,220) ..
+(224,220) -- (366,220) ..
+controls (368.21,220) and (370,221.79) ..
+(370,224) -- (370,286) ..
+controls (370,288.21) and (368.21,290) ..
+(366,290) -- (224,290) ..
+controls (221.79,290) and (220,288.21) ..
+(220,286) -- cycle ;
+    % Straight Lines [id:da7242957436529691] 
+    \draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (190.5,250) -- (200,250) -- (210.5,250) ;
+    % Straight Lines [id:da6155515815274885] 
+    \draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (10,220) -- (10,290) -- (10,270) -- (30,270) ;
+    % Shape: Rectangle [id:dp04289765101859899] 
+    \draw  [line width=1.5]  (40,344) ..
+controls (40,341.79) and (41.79,340) ..
+(44,340) -- (176,340) ..
+controls (178.21,340) and (180,341.79) ..
+(180,344) -- (180,386) ..
+controls (180,388.21) and (178.21,390) ..
+(176,390) -- (44,390) ..
+controls (41.79,390) and (40,388.21) ..
+(40,386) -- cycle ;
+    % Straight Lines [id:da912738070320213] 
+    \draw [line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (0,300) -- (0,400) -- (560,400) -- (560,30) -- (390,30) -- (390,180) -- (0,180) -- (0,300) -- (290,300) -- (390,300) -- (390,400) ;
+    % Straight Lines [id:da9646857023949835] 
+    \draw [line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (390,180) -- (390,330) ;
+    % Straight Lines [id:da36286215804614264] 
+    \draw [line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (0,180) -- (0,30) -- (390,30) ;
+    % Straight Lines [id:da5365951688341024] 
+    \draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (10,340) -- (10,370) -- (30,370) ;
+    % Straight Lines [id:da2808441683655002] 
+    \draw [color={rgb, 255:red, 126; green, 211; blue, 33 }  ,draw opacity=1 ][line width=1.5]    (10,170) -- (10,120) -- (30,120) ;
+
+    % Text Node
+    \draw (411,141) node [anchor=north west][inner sep=0.75pt]   [align=left] {User A};
+    % Text Node
+    \draw (412,73) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textcolor[rgb]{1,1,1}{>}};
+    % Text Node
+    \draw (411,251) node [anchor=north west][inner sep=0.75pt]   [align=left] {User B};
+    % Text Node
+    \draw (412,183) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textcolor[rgb]{1,1,1}{>}};
+    % Text Node
+    \draw (411,361) node [anchor=north west][inner sep=0.75pt]   [align=left] {User C};
+    % Text Node
+    \draw (412,293) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textcolor[rgb]{1,1,1}{>}};
+    % Text Node
+    \draw (51,82) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{31.110000000000003pt}\setlength\topsep{0pt}
+        \begin{center}
+          12345
+        \end{center}
+
+      \end{minipage}};
+    % Text Node
+    \draw (401,42) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\large Users}};
+    % Text Node
+    \draw (11,42) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\large Debug Pods}};
+    % Text Node
+    \draw (51,142) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{31.110000000000003pt}\setlength\topsep{0pt}
+        \begin{center}
+          67890
+        \end{center}
+
+      \end{minipage}};
+    % Text Node
+    \draw (222,73) node [anchor=north west][inner sep=0.75pt]   [align=left] {Pod Communication\\Server};
+    % Text Node
+    \draw (11,192) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\large Pod \& User Control}};
+    % Text Node
+    \draw (42,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {API Server\\Pod Management\\User Management};
+    % Text Node
+    \draw (222,223) node [anchor=north west][inner sep=0.75pt]   [align=left] {Frontend Server};
+    % Text Node
+    \draw (11,312) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\large Database}};
+    % Text Node
+    \draw (42,343) node [anchor=north west][inner sep=0.75pt]   [align=left] {User Storage\\Pod Storage};
+    % Text Node
+    \draw (1,2) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{12.931356000000001pt}\setlength\topsep{0pt}
+        \begin{flushright}
+          {\large 2.}
+        \end{flushright}
+
+      \end{minipage}};
+    % Text Node
+    \draw (362,37) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{12.931356000000001pt}\setlength\topsep{0pt}
+        \begin{flushright}
+          {\large a.}
+        \end{flushright}
+
+      \end{minipage}};
+    % Text Node
+    \draw (391,2) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{12.931356000000001pt}\setlength\topsep{0pt}
+        \begin{flushright}
+          {\large 1.}
+        \end{flushright}
+
+      \end{minipage}};
+    % Text Node
+    \draw (362,187) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{12.931356000000001pt}\setlength\topsep{0pt}
+        \begin{flushright}
+          {\large b.}
+        \end{flushright}
+
+      \end{minipage}};
+    % Text Node
+    \draw (363,307) node [anchor=north west][inner sep=0.75pt]   [align=left] {\begin{minipage}[lt]{12.24pt}\setlength\topsep{0pt}
+        \begin{flushright}
+          {\large c.}
+        \end{flushright}
+
+      \end{minipage}};
+
+  \end{tikzpicture}
+
+  \caption{Overview of the Collaborative Debugger}
+  \label{debugger:overview}
+\end{figure}
+
+The next sections give an overview of the various tools used to create
+the collaborative debugger.  The debugger consists of:
+
+\begin{enumerate}
+\item A frontend web app built using React that presents a debugging
+  interface to the end user (\ref{react}).
+\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
+  \begin{enumerate}
+  \item A pod for each debugging instance which runs the rr debugger
+    (\ref{rr}).  These communicate directly with users through
+    websockets (\ref{socketio}).
+  \item Pods running a frontend server written in Node which works in
+    tandem with an API server created using Flask (\ref{flask/node}).
+    The API server manages creation and destruction of debug sessions,
+    as well as authentication.
+  \item A MongoDB database.
+  \end{enumerate}
+\end{enumerate}
+
+\subsubsection{Kubernetes}\label{k8s}
+
+\subsubsection{Mozilla's rr}\label{rr}
 
@@
 
-\subsubsection{Kubernetes and Docker}
+\subsubsection{socket.io}\label{socketio}
 
-TODO: Write about k8s and docker.
+\subsubsection{Flask and Node}\label{flask/node}
 
-\subsubsection{React}
-
-TODO: Write about react.
-
-\subsubsection{Xterm.js}
-
-TODO: Write about Xterm.js if I end up using it (which seems
-overwhelmingly likely ATM).
+\subsubsection{React and Xterm.js}\label{react}
 

### 2020-11-19 containers/rr/record.py
basic compilation library

@@ -1,0 +1,37 @@
+import subprocess
+import re
+
+class Recorder:
+    def __init__(self, gcc_command):
+        self.gcc_args = gcc_command.strip()[3:].strip()
+        self.program_name = self.parse()
+
+    def parse(self):
+        try:
+            # Match either '-o program_name' or '-output program_name'
+            return re.search('((?<=-o)\s+\S+|(?<=-output)\s+\S+)',
+                             self.gcc_args).group().strip()
+        except:
+            # If there's no match, use default output name
+            return 'a.out'
+
+    def build(self):
+        p = subprocess.Popen(['gcc'] + self.gcc_args.split(' '),
+                             stdout=subprocess.PIPE, 
+                             stderr=subprocess.PIPE)
+        stdout, stderr = p.communicate()
+        return [stdout, stderr]
+
+    def record(self):
+        p = subprocess.Popen(['rr', 'record', './' + self.program_name],
+                             stdout=subprocess.PIPE, 
+                             stderr=subprocess.PIPE)
+        stdout, stderr = p.communicate()
+        return [stdout, stderr]
+
+
+if __name__ == '__main__':
+    d = Recorder('gcc -g -o test test.c')
+    print(d.gcc_args, d.parse())
+    print(d.build())
+    print(d.record())

### 2020-11-23 containers/frontend/react-flask-app/src/App.js
adding the beginnings of file upload, which are more complicated than initially anticipated

@@ -128,6 +128,39 @@ class ChannelForm extends React.Component {
     }
 }
 
+class FileForm extends React.Component {
+    // Sends a channel api request to connect to an rr debug session
+    constructor(props) {
+	super(props);
+	this.state = {file: null};
+	this.handleFile = this.handleFile.bind(this);
+	this.handleCommand = this.handleCommand.bind(this);
+	this.handleSubmit = this.handleSubmit.bind(this);
+    }
+
+    handleFile(event) {
+	this.setState({file: event.target.value});
+    }
+
+    handleSubmit(event) {
+	console.log(this.state.file);
+	this.props.socket.emit('upload_file', {'channel': this.props.channel, 'file': this.state.file});
+	event.preventDefault();
+    }
+
+    render() {
+	return (
+	    <form className='simple-form' onSubmit={this.handleSubmit}>
+	      <label>Upload a program file</label>
+	      <div className='simple-submit'>
+		<input type="file" value={this.state.file} onChange={this.handleChange} />
+		<button type="submit"><i className="material-icons">arrow_forward_ios</i></button>
+	      </div>
+	    </form>
+	);
+    }
+}
+
 class RRTerm extends React.Component {
 
     constructor(props) {
@@ -301,7 +334,8 @@ class Debugger extends React.Component {
 	super(props);
 	this.state = {user: '',
 		      channel: '',
-		      pods: []};
+		      pods: [],
+		      newPod: null};
 	
 	this.socket = socketIOClient('http://157.230.64.84:8000');
 	
@@ -334,14 +368,14 @@ class Debugger extends React.Component {
 	post_request('/new', {name: this.state.user}).then(data => {
 	    console.log(data);
 	    if (data.channel != null){
-		this.setState({channel: data.channel});
+		this.setState({channel: data.channel, newPod: true});
 		this.socket.emit('join_channel', {'channel': this.state.channel});
 	    }
 	});
     }
 
     onChannel(c) {
-	this.setState({channel: c});
+	this.setState({channel: c, newPod: false});
 	this.socket.emit('join_channel', {'channel': this.state.channel});
     }
 
@@ -363,8 +397,7 @@ class Debugger extends React.Component {
 	let newButton;
 	if (channelSet) {
 	    channel = <span>{this.state.channel}</span>;
-	}
-	else if (isLoggedIn) {
+	} else if (isLoggedIn) {
 	    channel = <ChannelForm name={this.state.user} onChannel={this.onChannel}/>;
[…]

### 2020-11-23 containers/rr/record.py

@@ -3,7 +3,8 @@ import re
 
 class Recorder:
     def __init__(self, gcc_command):
-        self.gcc_args = gcc_command.strip()[3:].strip()
+        # Strip out 'gcc' and tell gcc to output messages in JSON
+        self.gcc_args = '-fdiagnostics-format=json ' + gcc_command.strip()[3:].strip()
         self.program_name = self.parse()
 
     def parse(self):
@@ -31,7 +32,7 @@ class Recorder:
 
 
 if __name__ == '__main__':
-    d = Recorder('gcc -g -o test test.c')
+    d = Recorder('gcc -g -Wall -o test test.c')
     print(d.gcc_args, d.parse())
     print(d.build())
     print(d.record())

### 2020-11-23 containers/rr/rr-translation.py

@@ -1,45 +1,14 @@
-from pygdbmi.gdbcontroller import GdbController
 from pprint import pprint
 import socketio
 import sys
 
-class RRInterface:
-    def __init__(self, session=None):
-        self.gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
-        self.init_message = self.console_output(self.get_rr_response())
-        self.timeline = []
-
-    def exited(self, response):
-        # Unsure if this will always be the second to last message, works for now.
-        return response[-2]['message'] == 'thread-group-exited'
-
-    def end(self, response):
-        last = response[-1]['message']
-        return last == 'stopped' or last == 'done'
-
-    def get_rr_response(self):
-        return self.gdbmi.get_gdb_response()
-
-    def get_full_rr_response(self, command):
-        response = self.gdbmi.write(command)
-        while(not(self.end(response) or self.exited(response))):
-            response.extend(self.get_rr_response())
-        return response
-
-    def console_output(self, response):
-        output = []
-        for r in response:
-            if r['type'] == 'console' or r['type'] == 'output':
-                output.append(r['payload'])
-        return ''.join(output).replace("\\n", "\n").replace("\\t", "\t")
-    
-    def write(self, command):
-        self.timeline.append(command)
-        return self.console_output(self.get_full_rr_response(command))
+import rrinterface
+import record
 
 sio = socketio.Client()
-rri = RRInterface()
+rri = rrinterface.RRInterface()
 channel = None
+ready = False
 
 print(rri.init_message)
 

### 2020-11-24 paper/paper.tex
lots of little QOL things'

@@
 \begin{figure}[h!]
+
   \begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
@@
 
+Kubernetes is the defacto standard in container orchestration
+software.  It provides a layer of abstraction on top of normal
+containers, like those created by Docker.  By bundling one or more
+closely linked containers into a ``pod'', Kubernetes is able to manage
+deployment and re-deployment of applications running inside
+containers.  It is trivial to create new pods, or to create multiple
+pods running the same application as needed within a Kubernetes
+cluster.  Kubernetes provides services to facilitate load balancing,
+manage storage volumes,
+
 \subsubsection{Mozilla's rr}\label{rr}

### 2020-11-24 containers/controller/podmanager.py

@@ -17,11 +17,11 @@ class TranslationPodManager:
         config.load_kube_config()
         self.v1 = client.CoreV1Api()
         
-    def create_pod(self, names):
+    def create_pod(self, names, image='registry.digitalocean.com/sproj/rr:translation-1.0'):
         dep = {'apiVersion': 'v1',
                'kind': 'Pod',
                'metadata': {'labels': {'purpose': 'translate-rr'}},
-               'spec': {'containers': [{'image': 'registry.digitalocean.com/sproj/rr:translation-channel',
+               'spec': {'containers': [{'image': image,
                                         'name': 'rr-test-container',
                                         'command': ['sh'],
                                         'args': ['startup.sh'],
@@ -36,7 +36,6 @@ class TranslationPodManager:
         # Wait to return the channel until the pod is live and read to
         # recieve incomming communication
         status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
-        # 
         while status == None or status[0].state.running == None:
             time.sleep(1)
             status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
@@ -53,14 +52,6 @@ class TranslationPodManager:
         time.sleep(60)
         self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
         return
-        # while True:
-        #     try:
-        #         print(channel)
-        #         print(self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses[0].state.terminated)
-        #         await asyncio.sleep(1)
-        #     except:
-        #         self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
-        #         break
 
     def link_pod_to_users(self, channel, names):
         pod = self.dbc.get_pod_by_channel(channel)
@@ -78,3 +69,9 @@ class TranslationPodManager:
         active = [p['channel'] for p in pods]
         inactive = []
         return {'active': active, 'inactive': inactive}
+
+    def get_users_by_pod(self, channel):
+        pid = self.dbc.get_pod_by_channel(channel);
+        uids = self.dbc.get_users_by_pod(pid);
+        users = [u['name'] for u in uids]
+        return users

### 2020-11-24 containers/database/database.py

@@ -56,6 +56,9 @@ class DatabaseController:
         user = self.users.find_one({'name': name})
         return user['_id']
 
+    def get_users_by_pod(self, pid):
+        return self.users.find({"pods": pid})
+
     # Pod Methods
     
     def add_pod(self, uids=[]):

### 2020-11-24 containers/frontend/react-flask-app/api/api.py

@@ -7,6 +7,10 @@ tpm = podmanager.TranslationPodManager(url='165.227.252.45', port=27017)
 um = usermanager.UserManager(url='165.227.252.45', port=27017)
 app = Flask(__name__)
 
+images = {'cat': 'registry.digitalocean.com/sproj/rr:translation-1.0',
+          'stack_smash': 'registry.digitalocean.com/sproj/rr:translation-1.0',
+          'threads': 'registry.digitalocean.com/sproj/rr:translation-1.0'}
+
 @app.route('/login', methods=['POST'])
 def login():
     try:
@@ -22,10 +26,10 @@ def login():
                 # This isn't redundant, the user could have been added
                 # by a different instance of api.py in between our
                 # initial get_user and add_user
-                return {'error': 'user_exists'}
+                return {'error': 'User already exists.'}
         return {'name': name}
     except:
-        return {'error': 'internal_error'}
+        return {'error': 'Internal Error'}
 
 @app.route('/channel', methods=['POST'])
 def channel():
@@ -36,9 +40,9 @@ def channel():
             tpm.link_pod_to_users(channel, [name])
             return {'channel': channel}
         except database.NoMatchingChannelError:
-            return {'error': 'channel_error'}
+            return {'error': 'There is no open session with id: "' + channel + '".'}
     except:
-        return {'error': 'internal_error'}
+        return {'error': 'Internal Error'}
 
 @app.route('/pods', methods=['POST'])
 def pods():
@@ -47,10 +51,29 @@ def pods():
         pods = tpm.get_pods_by_user(user)
         return pods
     except:
-        return {'error': 'internal_error'}
+        return {'error': 'Internal Error'}
 
 @app.route('/new', methods=['POST'])
 def new():
-    name = request.get_json()['name']
-    channel = tpm.create_pod([name])
-    return {'channel': channel}
+    try:
+        name = request.get_json()['name']
+        program = request.get_json()['program']
+        image = images[program]
+        channel = tpm.create_pod([name])
+        return {'channel': channel}
+    except:
+        return {'error': 'Internal Error'}
+
+@app.route('/delete', methods=['POST'])
+def delete():
+    try:
+        name = request.get_json()['name']
+        channel = request.get_json()['channel']
+        users = tpm.get_users_by_pod(channel)
+        if len(users) <= 1:
+            tpm.delete_pod(channel)
+        else:
+            tpm.unlink_pod_from_users(channel, [name])
+        return {'deleted': True}
+    except:
+        return {'error': 'Internal Error'}

### 2020-11-24 containers/frontend/react-flask-app/src/App.js

@@ -18,37 +18,6 @@ async function post_request (url, data) {
     return response;
 }
 
-class RRForm extends React.Component {
-    // Standin form to test communication with RR pods.
-    constructor(props) {
-	super(props);
-	this.state = {value: ''};
-	this.handleChange = this.handleChange.bind(this);
-	this.handleSubmit = this.handleSubmit.bind(this);
-    }
-    
-    handleChange(event) {
-	this.setState({value: event.target.value});
-    }
-
-    handleSubmit(event) {
-	this.props.socket.emit('rr command', {'command': this.state.value, 'channel': this.props.channel})
-	event.preventDefault();
-    }
-
-    render() {
-	return (
-	    <form onSubmit={this.handleSubmit}>
-	      <label>
-		Command:
-		<input type="text" value={this.state.value} onChange={this.handleChange} />
-	      </label>
-	      <input type="submit" value="send"><i className="material-icons">face</i></input>
-	    </form>
-	);
-    }
-}
-
 class LoginForm extends React.Component {
     // Sends a login api request.  As of right now, users who don't
     // exist are automatically created... and you don't need a
@@ -70,6 +39,8 @@ class LoginForm extends React.Component {
 	    console.log(data);
 	    if (data.name != null){
 		this.props.onLogin(data.name);
+	    } else if (data.error != null){
+		alert('Error: ' + data.error);
 	    }
 	});
 	event.preventDefault();
@@ -110,6 +81,8 @@ class ChannelForm extends React.Component {
 	    console.log(data);
 	    if (data.channel != null){
 		this.props.onChannel(data.channel);
+	    } else if (data.error != null){
+		alert('Error joining session: ' + data.error);
 	    }
 	});
 	event.preventDefault();
@@ -132,14 +105,12 @@ class RRTerm extends React.Component {
 
     constructor(props) {
 	super(props);
-	this.state = {command: '', /*history: [], historyIndex: 0*/};
+	this.state = {command: ''};
 	this.inputRef=React.createRef();
 	this.term = null;
 	this.runTerminal = this.runTerminal.bind(this);
 	this.updateCommand = this.updateCommand.bind(this);
 	this.sendCommand = this.sendCommand.bind(this);
-	// this.historyPrevious = this.historyPrevious.bind(this);
-	// this.historyAdd = this.historyAdd.bind(this);
     }
 
[…]

### 2020-11-24 containers/rr-server/server.py

@@ -9,27 +9,26 @@ app = socketio.WSGIApp(sio)
 def join_channel(sid, data):
     sio.enter_room(sid, data['channel'])
 
-@sio.on('rr command')
+@sio.on('rr_command')
 def on_rr_command(sid, data):
-    print('Received rr command:')
     data['sid'] = sid
-    print(data)
-    print('Emitting...')
-    sio.emit('rr command', data, room=data['channel'])
+    try:
+        sio.emit('rr_command', data, room=data['channel'])
+    except:
+        # If there's an error with the data or the channel, there's
+        # not much we can do
+        pass
 
-@sio.on('rr response')
+@sio.on('rr_response')
 def on_rr_response(sid, data):
-    print('Recieved rr response:')
-    print(data)
-    print('Emitting...')
-    sio.emit('rr response', data, room=data['channel'])
+        sio.emit('rr_response', data, room=data['channel'])
 
 @sio.event
 def connect(sid, environ):
-    print('connect ', sid)
+    print('Connect: ', sid)
 
 @sio.event
 def disconnect(sid):
-    print('disconnect ', sid)
+    print('Disconnect: ', sid)
 
 eventlet.wsgi.server(eventlet.listen(('', 8000)), app)

### 2020-11-24 containers/rr/rr-translation.py

@@ -1,70 +1,0 @@
-from pygdbmi.gdbcontroller import GdbController
-from pprint import pprint
-import socketio
-import sys
-
-class RRInterface:
-    def __init__(self, session=None):
-        self.gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
-        self.init_message = self.console_output(self.get_rr_response())
-        self.timeline = []
-
-    def exited(self, response):
-        # Unsure if this will always be the second to last message, works for now.
-        return response[-2]['message'] == 'thread-group-exited'
-
-    def end(self, response):
-        last = response[-1]['message']
-        return last == 'stopped' or last == 'done'
-
-    def get_rr_response(self):
-        return self.gdbmi.get_gdb_response()
-
-    def get_full_rr_response(self, command):
-        response = self.gdbmi.write(command)
-        while(not(self.end(response) or self.exited(response))):
-            response.extend(self.get_rr_response())
-        return response
-
-    def console_output(self, response):
-        output = []
-        for r in response:
-            if r['type'] == 'console' or r['type'] == 'output':
-                output.append(r['payload'])
-        return ''.join(output).replace("\\n", "\n").replace("\\t", "\t")
-    
-    def write(self, command):
-        self.timeline.append(command)
-        return self.console_output(self.get_full_rr_response(command))
-
-sio = socketio.Client()
-rri = RRInterface()
-channel = None
-
-print(rri.init_message)
-
-@sio.event
-def connect():
-    print("I'm connected!")
-    
-@sio.on('rr command')
-def on_rr_command(data):
-    print('Recieved rr command:')
-    print(data)
-    print('Passing command to rr')
-    response = rri.write(data['command'])
-    print('Emitting...')
-    sio.emit('rr response', {'response': response, 'from': data['sid'], 'command': data['command'], 'channel': channel})
-    
-@sio.event
-def connect_error():
-    print("The connection failed!")
-
-@sio.event
-def disconnect():
-    print("I'm disconnected!")
-
-if __name__ == '__main__':
-    channel = sys.argv[1]
-    sio.connect('http://rr-host-load-balancer:8000')
-    sio.emit('join_channel', {'channel': channel})

### 2020-11-24 containers/rr/startup.sh

@@ -1,4 +1,4 @@
 gcc -g -o cat cat.c
 rr record ./cat cat.c test.py
 hostname -I
-python3 rr-translation.py $1
+python3 rrtranslation.py $1

### 2020-11-24 containers/controller/async-podmanager.py
Cleaned up structure, removed useless code, and added support for viewing source code

@@ -1,74 +1,0 @@
-from kubernetes import client, config
-import database
-import asyncio
-import time
-
-class TranslationPodManager:
-    def __init__(self, url='localhost', port=27017):
-        self.dbc = database.DatabaseController(url, port)
-        
-        config.load_kube_config()
-        self.v1 = client.CoreV1Api()
-        
-    async def create_pod(self, names):
-        dep = {'apiVersion': 'v1',
-               'kind': 'Pod',
-               'metadata': {'labels': {'purpose': 'translate-rr'}},
-               'spec': {'containers': [{'image': 'registry.digitalocean.com/sproj/rr:translation',
-                                        'name': 'rr-test-container',
-                                        'securityContext': {'capabilities': {'add': ['SYS_PTRACE']}}}],
-                        'restartPolicy': 'OnFailure'}}
-        channel = self.dbc.add_pod(self.dbc.get_userids_by_name(names))
-        dep['metadata']['name'] = channel
-        resp = self.v1.create_namespaced_pod(
-            body=dep, namespace='default')
-        print("Pod created.  status='%s'" % resp.metadata.name)
-        # Wait to return the channel until the pod is live and read to
-        # recieve incomming communication
-        status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
-        # 
-        while status == None or status[0].state.running == None:
-            await asyncio.sleep(1)
-            status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
-        return channel
-
-
-    async def delete_pod(self, channel):
-        """Given a channel, delete a pod"""
-        self.v1.delete_namespaced_pod(channel, 'default')
-        # Wait for double pod's maximum termination time (30 seconds).
-        # Waiting for termination eliminates the scenario where a pod
-        # is created with the same channel as a pod that is still
-        # terminating.  Creation of the new pod would fail.
-        await asyncio.sleep(60)
-        self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
-        return
-        # while True:
-        #     try:
-        #         print(channel)
-        #         print(self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses[0].state.terminated)
-        #         await asyncio.sleep(1)
-        #     except:
-        #         self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
-        #         break
-
[…]

### 2020-11-24 containers/controller/controller.py

@@ -1,30 +1,0 @@
-import sanic
-import socketio
-from podmanager import TranslationPodManager
-
-sio = socketio.AsyncServer(async_mode='sanic')
-app = sanic.Sanic(name='controller')
-sio.attach(app)
-tpm = TranslationPodManager(url='165.227.252.45', port=27017)
-
-# Receives a username from server requesting a pod be created
-# Returns a channel id that corresponds to the pod
-@sio.on('pod_create')
-async def on_pod_create(sid, data):
-    channel = await tpm.create_pod(data['names'])
-    await sio.emit('pod_created', {'names': data['names'], 'channel': channel}, to=sid)
-
-@sio.on('pod_delete')
-async def on_pod_delete(sid, data):
-    await tpm.delete_pod(data['channel'])
-    await sio.emit('pod_deleted', {'channel': data['channel']})
-        
-@sio.event
-async def connect(sid, environ):
-    print('connect ', sid)
-
-@sio.event
-async def disconnect(sid):
-    print('disconnect ', sid)
-
-app.run(port='8001')

### 2020-11-24 containers/controller/database.py

@@ -1,92 +1,0 @@
-from pymongo import MongoClient, errors
-from pprint import pprint
-import random
-
-class NoOpenChannelError(Exception):
-    """Raised when all channels are in use."""
-    pass
-
-class NoMatchingChannelError(Exception):
-    """Raised when no pod matches a provided channel"""
-    pass
-
-class DuplicateUserError(Exception):
-    """Raised when a duplicated username is attempted to be used"""
-    pass
-
-class DatabaseController:
-    """Provides interface to mongodb database"""
-    def __init__(self, url='localhost', port=27017):
-        self.client = MongoClient(url, port)
-        self.db = self.client.debugger
-        self.users = self.db.users
-        self.pods = self.db.pods
-        self.min_channel = 10000
-        self.max_channel = 99999
-        self.max_channels = self.max_channel - self.min_channel
-
-    # User Methods
-
-    def add_user(self, name):
-        """Given a name, attempt to add the user into the database"""
-        return str(self.users.insert_one({'name': name,
-                                          'pods': []}).inserted_id)
-
-    def delete_user(self, uid):
-        """Given a uid, delete a user"""
-        self.pods.update_many({}, {'$pull': {'users': uid}})
-        self.users.delete_one({'_id': uid})
-
-    def get_userids_by_name(self, names):
-        """Given a list of names, retrieve the mongodb ids for the users"""
-        users = self.users.find({'name': {'$in': names}}, {"_id": 1})
-        if users:
-            return [u['_id'] for u in users]
-        else:
-            return []
-
-    # Pod Methods
-    
-    def add_pod(self, uids=[]):
-        """Given a list of uids, creates a pod and returns a channel to use for communicating with the pod"""
-        # While there are open channels, try to create a new pod with
-        # an open channel.  Obviously in real use there should be a
-        # better method for generating channel numbers, but given the
-        # number of current users the likelyhood of a collision is
-        # extreemly low using 5 digit numbers
-        while self.pods.count_documents({'channel': {'$not': {'$eq': None}}}) <= self.max_channels:
-            try:
-                channel = str(random.randint(self.min_channel, self.max_channel))
-                pid = self.pods.insert_one({'channel': channel}).inserted_id
-                if uids:
-                    self.link_pod_to_users(pid, uids)
-                return channel
-            except errors.DuplicateKeyError:
-                pass
-        raise NoOpenChannelError
-    
[…]

### 2020-11-24 containers/controller/podmanager.py

@@ -17,7 +17,7 @@ class TranslationPodManager:
         config.load_kube_config()
         self.v1 = client.CoreV1Api()
         
-    def create_pod(self, names, image='registry.digitalocean.com/sproj/rr:translation-1.0'):
+    def create_pod(self, names, image='registry.digitalocean.com/sproj/rr:translation-1.0.6'):
         dep = {'apiVersion': 'v1',
                'kind': 'Pod',
                'metadata': {'labels': {'purpose': 'translate-rr'}},

### 2020-11-24 containers/frontend/react-flask-app/api/api.py

@@ -7,9 +7,9 @@ tpm = podmanager.TranslationPodManager(url='165.227.252.45', port=27017)
 um = usermanager.UserManager(url='165.227.252.45', port=27017)
 app = Flask(__name__)
 
-images = {'cat': 'registry.digitalocean.com/sproj/rr:translation-1.0',
-          'stack_smash': 'registry.digitalocean.com/sproj/rr:translation-1.0',
-          'threads': 'registry.digitalocean.com/sproj/rr:translation-1.0'}
+images = {'cat': 'registry.digitalocean.com/sproj/rr:translation-1.0.6',
+          'stack_smash': 'registry.digitalocean.com/sproj/rr:translation-1.0.6',
+          'threads': 'registry.digitalocean.com/sproj/rr:translation-1.0.6'}
 
 @app.route('/login', methods=['POST'])
 def login():

### 2020-11-24 containers/frontend/react-flask-app/api/podmanager.py

@@ -1,0 +1,1 @@
+../../../controller/podmanager.py

### 2020-11-24 containers/frontend/react-flask-app/api/usermanager.py

@@ -1,0 +1,1 @@
+../../../controller/usermanager.py

### 2020-11-24 containers/frontend/react-flask-app/src/App.js

@@ -3,11 +3,13 @@ import socketIOClient from 'socket.io-client';
 import './App.css';
 import XTerm, {Terminal} from 'react-xterm';
 import 'xterm/css/xterm.css';
+import MonacoEditor from 'react-monaco-editor';
 import WebFont from 'webfontloader';
 
 async function post_request (url, data) {
     // Send a post request to the server, which will forward it to our
     // Flask API handler
+    
     const response = await fetch(url, {
 	method: 'POST',
 	body: JSON.stringify(data),
@@ -22,6 +24,7 @@ class LoginForm extends React.Component {
     // Sends a login api request.  As of right now, users who don't
     // exist are automatically created... and you don't need a
     // password...
+    
     constructor(props) {
 	super(props);
 	this.state = {name: ''};
@@ -34,10 +37,10 @@ class LoginForm extends React.Component {
     }
 
     handleSubmit(event) {
-	console.log(this.state.name);
 	post_request('/login', {name: this.state.name}).then(data => {
-	    console.log(data);
 	    if (data.name != null){
+		// Call the onLogin function provided by Debugger to
+		// actually log the user in
 		this.props.onLogin(data.name);
 	    } else if (data.error != null){
 		alert('Error: ' + data.error);
@@ -61,6 +64,7 @@ class LoginForm extends React.Component {
 
 class ChannelForm extends React.Component {
     // Sends a channel api request to connect to an rr debug session
+    
     constructor(props) {
 	super(props);
 	this.state = {channel: ''};
@@ -73,13 +77,13 @@ class ChannelForm extends React.Component {
     }
 
     handleSubmit(event) {
-	console.log(this.state.channel);
 	// If successful, associates the currently logged in user with
 	// a pod server side and sets the channel to begin sending
 	// socket messages
 	post_request('/channel', {channel: this.state.channel, name: this.props.name}).then(data => {
-	    console.log(data);
 	    if (data.channel != null){
+		// Call the onChannel function provided by Debugger to
+		// set the channel
 		this.props.onChannel(data.channel);
 	    } else if (data.error != null){
 		alert('Error joining session: ' + data.error);
@@ -102,18 +106,32 @@ class ChannelForm extends React.Component {
 }
 
 class RRTerm extends React.Component {
-
+    // This class contains two main components: a view of the source
[…]

### 2020-11-24 containers/rr-server/pytest_gdb.py

@@ -1,10 +1,0 @@
-from pygdbmi.gdbcontroller import GdbController
-from pprint import pprint
-
-# Start gdb process
-gdbmi = GdbController(gdb_args =["--interpreter=mi2"])
-print(gdbmi.get_subprocess_cmd())  # print actual command run as subprocess
-response = gdbmi.write('-break-insert main')
-response = gdbmi.write('-exec-run')
-response = gdbmi.write('-exec-continue')
-pprint(response)

### 2020-11-24 containers/rr-translation/rrtranslation.py

@@ -1,0 +1,115 @@
+from pygdbmi.gdbcontroller import GdbController
+from pprint import pprint
+import socketio
+import traceback
+import sys
+
+class DissallowedError(Exception):
+    """Raised when the user tries to execute a banned command."""
+    pass
+
+class RRInterface:
+    def __init__(self, session=None):
+        self.gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
+        self.init_message = self.console_output(self.get_rr_response())
+        self.timeline = []
+
+    def command_forbidden(self, command):
+        # Allowing shell commands is a security risk, useless for the
+        # purpose of the application, and annoying to handle
+        return command == 'shell'
+        
+    def exited(self, response):
+        return len(response) > 2 and response[-2]['message'] == 'thread-group-exited'
+
+    def end(self, response):
+        last = response[-1]['message']
+        return len(response) == 1 or last == 'stopped' or last == 'done' or last == 'error'
+
+    def get_rr_response(self):
+        return self.gdbmi.get_gdb_response()
+
+    def get_full_rr_response(self, command):
+        response = self.gdbmi.write(command)
+        while(not(self.end(response) or self.exited(response))):
+            response.extend(self.get_rr_response())
+        return response
+
+    def console_output(self, response):
+        output = []
+        for r in response:
+            if r['type'] == 'console' or r['type'] == 'output':
+                output.append(r['payload'])
+            if r['message'] == 'error':
+                output.append(r['payload']['msg'])
+        return ''.join(output).replace("\\n", "\n").replace("\\t", "\t")
+    
+    def write(self, command):
+        if self.command_forbidden(command):
+            raise DissallowedError
+        self.timeline.append(command)
+        return self.console_output(self.get_full_rr_response(command))
+
+    def source(self):
+        f = self.get_full_rr_response('-file-list-exec-source-file')[0]['payload']
+        return {'file': f['file'],
+                'line': f['line'],
+                'path': f['fullname']}
+
+sio = socketio.Client()
+rri = RRInterface()
+channel = None
+current_source = ''
+
+@sio.event
+def connect():
+    print("Connected.")
+    
+@sio.on('rr_command')
+def on_rr_command(data):
+
+    body = {'from': data['sid'],
+            'command': data['command'],
+            'channel': channel}
+    
+    try:
+        response = {'output': rri.write(data['command'])}
+        # Get the current source file.
+        s = rri.source()
+        if s['path'] == current_source:
[…]

### 2020-11-24 containers/rr-translation/startup.sh

@@ -1,4 +1,4 @@
 gcc -g -o cat cat.c
-rr record ./cat cat.c test.py
+rr record ./cat cat.c cat.c
 hostname -I
 python3 rrtranslation.py $1

### 2020-11-24 containers/rr/test.py

@@ -1,9 +1,0 @@
-from pygdbmi.gdbcontroller import GdbController
-from pprint import pprint
-
-# Start gdb process
-gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
-response = gdbmi.write('-break-insert main')
-response = gdbmi.write('-exec-run')
-response = gdbmi.write('-exec-continue')
-pprint(response)

### 2020-11-24 containers/rr/tolower.c

@@ -1,11 +1,0 @@
-#include <stdio.h>
-#include <ctype.h>
-
-int main(){
-  int c;
-
-  while ((c = getchar()) != EOF)
-    putchar(tolower(c));
-
-  return 0;
-}

### 2020-11-25 containers/controller/podmanager.py
Full working before upload to cluster

@@ -75,3 +75,10 @@ class TranslationPodManager:
         uids = self.dbc.get_users_by_pod(pid);
         users = [u['name'] for u in uids]
         return users
+
+    def get_examples(self):
+        examples = self.dbc.examples.find()
+        if examples == None:
+            return {}
+        else:
+            return {e['name']: e['container'] for e in examples}

### 2020-11-25 containers/database/database-setup.py

@@ -9,3 +9,4 @@ client = MongoClient(url, port)
 
 client.debugger.users.create_index('name', unique=True)
 client.debugger.pods.create_index('channel', unique=True)
+client.debugger.examples.create_index('name', unique=True)

### 2020-11-25 containers/database/database.py

@@ -25,6 +25,7 @@ class DatabaseController:
         self.db = self.client.debugger
         self.users = self.db.users
         self.pods = self.db.pods
+        self.examples = self.db.examples
         self.min_channel = 10000
         self.max_channel = 99999
         self.max_channels = self.max_channel - self.min_channel
@@ -107,3 +108,8 @@ class DatabaseController:
                                {'$pull': {'pods': pod}})
         self.pods.update_one({'_id': pod},
                              {'$pull': {'users': {'$in': uids}}})
+
+    # Example Methods
+
+    def get_examples(self):
+        return db.examples.find({})

### 2020-11-25 containers/frontend/react-flask-app/api/api.py

@@ -7,10 +7,6 @@ tpm = podmanager.TranslationPodManager(url='165.227.252.45', port=27017)
 um = usermanager.UserManager(url='165.227.252.45', port=27017)
 app = Flask(__name__)
 
-images = {'cat': 'registry.digitalocean.com/sproj/rr:translation-1.0.6',
-          'stack_smash': 'registry.digitalocean.com/sproj/rr:translation-1.0.6',
-          'threads': 'registry.digitalocean.com/sproj/rr:translation-1.0.6'}
-
 @app.route('/login', methods=['POST'])
 def login():
     try:
@@ -58,8 +54,10 @@ def new():
     try:
         name = request.get_json()['name']
         program = request.get_json()['program']
-        image = images[program]
-        channel = tpm.create_pod([name])
+        examples = tpm.get_examples()
+        container = examples[program]
+        print(container)
+        channel = tpm.create_pod([name], container)
         return {'channel': channel}
     except:
         return {'error': 'Internal Error'}
@@ -75,5 +73,13 @@ def delete():
         else:
             tpm.unlink_pod_from_users(channel, [name])
         return {'deleted': True}
+    except:
+        return {'error': 'Internal Error'}
+
+@app.route('/examples', methods=['POST'])
+def examples():
+    try:
+        examples = tpm.get_examples()
+        return {'examples': list(examples.keys())}
     except:
         return {'error': 'Internal Error'}

### 2020-11-25 containers/frontend/react-flask-app/src/App.js

@@ -194,6 +194,7 @@ class RRTerm extends React.Component {
 	// Register a new response handler with the socket defined in
 	// Debugger.  
 	this.props.socket.on('rr_response', (data) => {
+	    console.log(data)
 	    if(data.from !== this.props.socket.id){
 		// If the command originated from a different client,
 		// display it.  Disable input so the user doesn't mess
@@ -294,7 +295,7 @@ function NewDebugButton(props) {
     const program = props.program
     return (
 	<button onClick={() => {props.onClick(program)}}>
-	  <span>{program}</span>
+	  <span className={'button-label'}>{program}</span>
 	</button>
     );
 }
@@ -310,6 +311,14 @@ function DeleteButton(props) {
 	<button onClick={() => {props.onClick(pod)}}>
 	  {icon}
 	</button>
+    );
+}
+
+function ExampleList(props) {
+    const examples = props.examples;
+    const listItems = examples.map((example) => <li key={example}><NewDebugButton onClick={props.onClick} program={example} /></li>)
+    return (
+	<ul>{listItems}</ul>
     );
 }
 
@@ -329,7 +338,9 @@ class Debugger extends React.Component {
 	super(props);
 	this.state = {user: '',
 		      channel: '',
-		      pods: []};
+		      examples: [],
+		      pods: []
+		     };
 	
 	this.socket = socketIOClient('http://157.230.64.84:8000');
 	
@@ -350,6 +361,13 @@ class Debugger extends React.Component {
 		this.setState({pods: data.active});
 	    } else if (data.error != null) {
 		alert('Error fetching list of active pods: ' + data.error);
+	    }
+	});
+	post_request('/examples', {name: this.state.user}).then(data => {
+	    if(data.examples != null) {
+		this.setState({examples: data.examples});
+	    } else if (data.error != null) {
+		alert('Error fetching example pods: ' + data.error);
 	    }
 	});
     }
@@ -404,10 +422,12 @@ class Debugger extends React.Component {
 	}
 
 	const channelSet = this.state.channel !== '';
+	
 	// Channel to use with socket, identifies session pod.
 	let channel;
-	// Button to create a new pod.
-	let newButton;
+	// Example pods
+	let examples;
+	
 	if (channelSet) {
 	    channel = (
 		<div>
@@ -415,22 +435,15 @@ class Debugger extends React.Component {
 		  <span>{this.state.channel}</span>
 		</div>
 	    );
[…]

### 2020-11-25 containers/rr-translation/rrtranslation.py

@@ -3,6 +3,7 @@ from pprint import pprint
 import socketio
 import traceback
 import sys
+import re
 
 class DissallowedError(Exception):
     """Raised when the user tries to execute a banned command."""
@@ -12,19 +13,24 @@ class RRInterface:
     def __init__(self, session=None):
         self.gdbmi = GdbController(command=["rr", "replay", "--", "--nx", "--quiet", "--interpreter=mi2"])
         self.init_message = self.console_output(self.get_rr_response())
+        
+        # Disable hardware watchpoints
+        # self.gdbmi.write('set can-use-hw-watchpoints 0')
         self.timeline = []
 
     def command_forbidden(self, command):
         # Allowing shell commands is a security risk, useless for the
-        # purpose of the application, and annoying to handle
-        return command == 'shell'
+        # purpose of the application, and annoying to handle.
+        # Hardware watchpoints must stay disabled since too many users
+        # overwhelms the capability of the CPU
+        return command == 'shell' # or re.search('(set can-use-hw-watchpoints.*)', command) != None
         
     def exited(self, response):
         return len(response) > 2 and response[-2]['message'] == 'thread-group-exited'
 
     def end(self, response):
         last = response[-1]['message']
-        return len(response) == 1 or last == 'stopped' or last == 'done' or last == 'error'
+        return len(response) == 1 or last == 'stopped' or last == 'done' or last == 'error' or last == 'running'
 
     def get_rr_response(self):
         return self.gdbmi.get_gdb_response()
@@ -38,11 +44,13 @@ class RRInterface:
     def console_output(self, response):
         output = []
         for r in response:
-            if r['type'] == 'console' or r['type'] == 'output':
+            if r['type'] == 'console':
                 output.append(r['payload'])
+            if r['type'] == 'output':
+                output.append(r['payload'] + '\n')
             if r['message'] == 'error':
                 output.append(r['payload']['msg'])
-        return ''.join(output).replace("\\n", "\n").replace("\\t", "\t")
+        return ''.join(output).replace('\\n', '\n').replace('\\t', '\t')
     
     def write(self, command):
         if self.command_forbidden(command):
@@ -58,6 +66,7 @@ class RRInterface:
 
 sio = socketio.Client()
 rri = RRInterface()
+
 channel = None
 current_source = ''
 
@@ -85,12 +94,17 @@ def on_rr_command(data):
         else:
             # If it's different, read the contents of the file, and
             # send that along with the file name and line number
[…]

### 2020-11-25 example_programs/hash/hash.c

@@ -1,0 +1,95 @@
+#include <string.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+#define HASHSIZE 101
+
+struct pair {
+  struct pair *next;
+  char *key;
+  char *value;
+};
+
+static struct pair *hashtable[HASHSIZE];
+
+unsigned hash (char *s)
+{
+  unsigned hashval;
+  for (hashval = 0; *s != '\0'; s++)
+    hashval = *s + 31 * hashval;
+  return hashval % HASHSIZE;
+}
+
+struct pair *search(char *s)
+{
+  struct pair *pp;
+
+  for (pp = hashtable[hash(s)]; pp != NULL; pp = pp->next)
+    if (strcmp(s, pp->key) == 0)
+      return pp;
+  return NULL;
+}
+
+/* The bug here is that they forget to use strdup */
+struct pair *insert(char *key, char *value)
+{
+  struct pair *pp;
+  unsigned hashval;
+
+  if ((pp = search(key)) == NULL) {
+    pp = (struct pair *) malloc(sizeof(*pp));
+    if (pp == NULL)
+      return NULL;
+    pp->key = key;
+    hashval = hash(key);
+    pp->next = hashtable[hashval];
+    hashtable[hashval] = pp;
+  } else {
+    free((void *) pp->value);
+  }
+  pp->value = value;
+  return pp;
+}
+
+#define BUFFSIZE 1024
+
+int main() {
+  char * fname = "names";
+  FILE * f = fopen(fname, "r");
+  if (f == NULL){
+    printf("Error opening file: %s\n", fname);
+    return 1;
+  }
+
+  char kb[BUFFSIZE];
+  char vb[BUFFSIZE];
+  
+  while(fgets(kb, BUFFSIZE, f) != NULL && fgets(vb, BUFFSIZE, f) != NULL){
+    /* Strip out newline */
+    kb[strcspn(kb, "\n")] = 0;
+    vb[strcspn(vb, "\n")] = 0;
+    
+    if(insert(kb, vb) == NULL){
+      printf("Error inserting pair: %s, %s\n", kb, vb);
+      return 1;
+    }
+
+    printf("Inserted pair: %s, %s\n", kb, vb);
+  }
+
+  rewind(f);
+
+  char ignore[BUFFSIZE];
+  while(fgets(kb, BUFFSIZE, f) != NULL && fgets(ignore, BUFFSIZE, f) != NULL){
[…]

### 2020-11-25 example_programs/hash/startup.sh

@@ -1,0 +1,3 @@
+gcc -g -o hash hash.c
+rr record ./hash
+python3 rrtranslation.py $1

### 2020-11-25 containers/controller/podmanager.py
IT WORKS IN FULL ON DIGITAL OCEAN

@@ -17,7 +17,7 @@ class TranslationPodManager:
         config.load_kube_config()
         self.v1 = client.CoreV1Api()
         
-    def create_pod(self, names, image='registry.digitalocean.com/sproj/rr:translation-1.0.6'):
+    def create_pod(self, names, image):
         dep = {'apiVersion': 'v1',
                'kind': 'Pod',
                'metadata': {'labels': {'purpose': 'translate-rr'}},

### 2020-11-25 containers/frontend/react-flask-app/api/api.py

@@ -3,8 +3,8 @@ import podmanager
 import usermanager
 import database
 
-tpm = podmanager.TranslationPodManager(url='165.227.252.45', port=27017)
-um = usermanager.UserManager(url='165.227.252.45', port=27017)
+tpm = podmanager.TranslationPodManager(url='mongodb://database-load-balancer', port=27017)
+um = usermanager.UserManager(url='mongodb://database-load-balancer', port=27017)
 app = Flask(__name__)
 
 @app.route('/login', methods=['POST'])

### 2020-11-25 containers/frontend/react-flask-app/api/database.py

@@ -1,1 +1,115 @@
-../../../database/database.py
+from pymongo import MongoClient, errors
+from pprint import pprint
+import random
+
+class NoOpenChannelError(Exception):
+    """Raised when all channels are in use."""
+    pass
+
+class NoMatchingChannelError(Exception):
+    """Raised when no pod matches a provided channel"""
+    pass
+
+class DuplicateUserError(Exception):
+    """Raised when a duplicated username is attempted to be used"""
+    pass
+
+class NoMatchingUserError(Exception):
+    """Raised when a username isn't registered"""
+    pass
+
+class DatabaseController:
+    """Provides interface to mongodb database"""
+    def __init__(self, url='localhost', port=27017):
+        self.client = MongoClient(url, port)
+        self.db = self.client.debugger
+        self.users = self.db.users
+        self.pods = self.db.pods
+        self.examples = self.db.examples
+        self.min_channel = 10000
+        self.max_channel = 99999
+        self.max_channels = self.max_channel - self.min_channel
+        
+    # User Methods
+
+    def add_user(self, name):
+        """Given a name, attempt to add the user into the database"""
+        try:
+            return str(self.users.insert_one({'name': name,
+                                              'pods': []}).inserted_id)
+        except pymongo.DuplicateKeyError:
+            raise DuplicateUserError
+
+    def delete_user(self, uid):
+        """Given a uid, delete a user"""
+        self.pods.update_many({}, {'$pull': {'users': uid}})
+        self.users.delete_one({'_id': uid})
+        
+    def get_userids_by_name(self, names):
+        """Given a list of names, retrieve the mongodb ids for the users"""
+        users = self.users.find({'name': {'$in': names}}, {"_id": 1})
+        if users:
+            return [u['_id'] for u in users]
+        else:
+            return []
+
+    def get_userid_by_name(self, name):
+        user = self.users.find_one({'name': name})
+        return user['_id']
+
+    def get_users_by_pod(self, pid):
+        return self.users.find({"pods": pid})
+
+    # Pod Methods
+    
+    def add_pod(self, uids=[]):
+        """Given a list of uids, creates a pod and returns a channel to use for communicating with the pod"""
+        # While there are open channels, try to create a new pod with
+        # an open channel.  Obviously in real use there should be a
+        # better method for generating channel numbers, but given the
+        # number of current users the likelyhood of a collision is
+        # extreemly low using 5 digit numbers
[…]

### 2020-11-25 containers/frontend/react-flask-app/api/podmanager.py

@@ -1,1 +1,84 @@
-../../../controller/podmanager.py
+from kubernetes import client, config
+import database
+import time
+
+class NoOpenChannelError(Exception):
+    """Raised when all channels are in use."""
+    pass
+
+class NoMatchingChannelError(Exception):
+    """Raised when no pod matches a provided channel"""
+    pass
+
+class TranslationPodManager:
+    def __init__(self, url='localhost', port=27017):
+        self.dbc = database.DatabaseController(url, port)
+        
+        config.load_incluster_config()
+        self.v1 = client.CoreV1Api()
+        
+    def create_pod(self, names, image):
+        dep = {'apiVersion': 'v1',
+               'kind': 'Pod',
+               'metadata': {'labels': {'purpose': 'translate-rr'}},
+               'spec': {'containers': [{'image': image,
+                                        'name': 'rr-test-container',
+                                        'command': ['sh'],
+                                        'args': ['startup.sh'],
+                                        'securityContext': {'capabilities': {'add': ['SYS_PTRACE']}}}],
+                        'restartPolicy': 'OnFailure'}}
+        channel = self.dbc.add_pod(self.dbc.get_userids_by_name(names))
+        dep['metadata']['name'] = channel
+        dep['spec']['containers'][0]['args'].append(channel)
+        resp = self.v1.create_namespaced_pod(
+            body=dep, namespace='default')
+        print("Pod created.  status='%s'" % resp.metadata.name)
+        # Wait to return the channel until the pod is live and read to
+        # recieve incomming communication
+        status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
+        while status == None or status[0].state.running == None:
+            time.sleep(1)
+            status = self.v1.read_namespaced_pod_status(channel, 'default').status.container_statuses
+        return channel
+
+
+    def delete_pod(self, channel):
+        """Given a channel, delete a pod"""
+        self.v1.delete_namespaced_pod(channel, 'default')
+        # Wait for double pod's maximum termination time (30 seconds).
+        # Waiting for termination eliminates the scenario where a pod
+        # is created with the same channel as a pod that is still
+        # terminating.  Creation of the new pod would fail.
+        time.sleep(60)
+        self.dbc.delete_pod(self.dbc.get_pod_by_channel(channel))
+        return
+
+    def link_pod_to_users(self, channel, names):
[…]

### 2020-11-25 containers/frontend/react-flask-app/api/usermanager.py

@@ -1,1 +1,30 @@
-../../../controller/usermanager.py
+import database
+
+class NoMatchingUserError(Exception):
+    """Raised when a username isn't registered"""
+    pass
+
+class UserManager:
+    """Strictly for managing users (auth, creation, deletion, etc).  Pod
+    and user synchronization are handled by TranslationPodManager.
+    """
+    
+    def __init__(self, url='localhost', port=27017):
+        self.dbc = database.DatabaseController(url, port)
+
+    def get_user(self, name):
+        """Given a name, return a userid"""
+        try:
+            uid = self.dbc.get_userid_by_name(name)
+            return uid
+        except TypeError:
+            raise NoMatchingUserError
+
+    def add_user(self, name):
+        """Given a name, attempt to add the user into the database"""
+        self.dbc.add_user(name)
+
+
+    def delete_user(self, name):
+        """Given a name, delete a user"""
+        self.dbc.delete_user(self.dbc.get_userids_by_name([name])[0])

### 2020-11-25 containers/frontend/react-flask-app/startup.sh

@@ -1,0 +1,2 @@
+yarn start-api &
+yarn start

### 2020-11-27 paper/paper.tex
Added some tool writeups

--- 2020-10-27 paper/paper.tex
@@
 \usepackage{listings}
-\usepackage{tikz}
+\usepackage{graphicx}
 \usepackage[doublespacing]{setspace}
-
+\usepackage{hyperref}
 \usepackage{titlesec}
 \newcommand{\sectionbreak}{\clearpage}
-\usetikzlibrary{positioning}
 
 \lstset{basicstyle=\ttfamily}
+\graphicspath{{./images/}}
 
@@
 
-\subsubsection{Mozilla's rr}
+\begin{figure}[h!]
+
+  \includegraphics[scale=.8]{overall_system}
+  \centering
+  \caption{Overview of the Collaborative Debugger}
+  \label{debugger:overview}
+\end{figure}
+
+The next sections give an overview of the various tools used to create
+the collaborative debugger.  The debugger consists of:
+
+\begin{enumerate}
+\item A frontend web app built using React (\ref{react}) that presents a debugging
+  interface to the end user.
+\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
+  \begin{enumerate}
+  \item A pod for each debugging instance which runs the rr debugger
+    (\ref{rr}).  These communicate directly with users through
+    WebSocket server pods. (\ref{socketio}).
+  \item Pods running a frontend server written in Node which works in
+    tandem with an API server created using Flask (\ref{flask/node}).
+    The API server manages creation and destruction of debug sessions,
+    as well as authentication.
+  \item A MongoDB database.
+  \end{enumerate}
+\end{enumerate}
+
+\subsubsection{Kubernetes}\label{k8s}
+
+Kubernetes is the defacto standard in container orchestration
+software.  It provides a layer of abstraction on top of normal
+containers, like those created by Docker.  By bundling one or more
+closely linked containers into a ``pod'', Kubernetes is able to manage
+deployment and re-deployment of applications running inside
+containers.  It is trivial to create new pods, or to create multiple
+pods running the same application as needed within a Kubernetes
+cluster \cite{k8s}.  The speed at which even relatively large pods can be created
+and the inherent security provided by containerization drove the
+decision to create a new pod on the fly for each debugging instance in
+the collaborative debugger.
+\par
+
+Kubernetes also provides services to facilitate load balancing, manage
+storage volumes, and contain secrets.  The abstraction provided by
+these features, in tandem with the ease of Kubernetes deployment on a
+managed Kubernetes service\cite{do_managed_k8s} greatly accelerated
+development.
+
+\subsubsection{Mozilla's rr}\label{rr}
 
@@
 rr is ``a lightweight tool for recording, replaying and debugging
-execution of applications''.\cite{rr-repo} rr allows a programmer to
+execution of applications''\cite{rr-repo}. rr allows a programmer to
 record the execution of a program on any compatible machine and replay
@@
 progressing back and forth through execution in the problem area.
+\par
+
+rr is an ideal tool for teaching debugging because it allows
+instructors to record execution of a program and design a debugging
+example with the knowledge that normally non-deterministic events will
+be repeatable, and that any input they provide to the program will be
+exactly replicated.  With the collaborative debugger, teachers can
+record a program's execution and design a debugging lesson which
+students can work on together.  The repeatablity of rr means that
+students can focus on debugging, and teachers can create as specific
+examples as they please.
 
@@
 webapp similar to existing tools for collaborative programming.
+\par
+
 Luckily, the speed and size benefits of rr lend themselves well to
-non-local execution.
+non-local execution.  In conjunction with Kubernetes, it takes a few
+seconds to create a new container running rr and connect to web
+clients.
 
-\subsubsection{Kubernetes and Docker}
+\subsubsection{pygdbmi}
 
-TODO: Write about k8s and docker.
+\subsubsection{Socket.IO}\label{socketio}
 
-\subsubsection{React}
+To speed communication, the collaborative debugger uses WebSockets to
+directly connect web clients and the pods running rr.  Socket.IO is a
+library that extends WebSockets.  It provides backup incase a
+WebSocket connection cannot be established, enables automatic
+reconnection and disconnection detection, and adds support for
+namespaces \cite{socketio}.  The collaborative debugger uses the
+standard Javascript implementation of Socket.IO on the client side.
+Messages are passed through a server to individual debugging pods,
+both of which use the Python implementation of Socket.IO,
+python-socket.io \cite{python_socketio}.
 
-TODO: Write about react.
+\subsubsection{Flask and Node}\label{flask/node}
 
-\subsubsection{Xterm.js}
-
-TODO: Write about Xterm.js if I end up using it (which seems
-overwhelmingly likely ATM).
+\subsubsection{React and Xterm.js}\label{react}
 

### 2020-11-28 paper/paper.tex
finished draft of tool writeups

@@
 
-\subsection{Project Components}
+\subsection{Tools Used}
 
@@
     WebSocket server pods. (\ref{socketio}).
-  \item Pods running a frontend server written in Node which works in
-    tandem with an API server created using Flask (\ref{flask/node}).
-    The API server manages creation and destruction of debug sessions,
-    as well as authentication.
-  \item A MongoDB database.
+  \item Pods running a frontend server written in Node.js which works
+    in tandem with an API server created using Flask
+    (\ref{flask/node}).  The API server manages creation and
+    destruction of debug sessions, as well as authentication.
+  \item A MongoDB (\ref{mongodb}) database.
   \end{enumerate}
@@
 
+In order to ``support the development of systems which use the
+debugger as just one small component of a larger system'', GDB
+provides a machine-oriented interface called GDB/MI \cite{gdbman}. rr
+supports interaction through GDB/MI, and using the interface was a
+natural choice for the collaborative debugger.  In addition to being
+far easier to interact with from within a program, the structured,
+machine-friendly output of GDB/MI lends itself in particular to future
+development of visualization aids in the collaborative debugger.
+\par
+To parse rr output into Python dictionaries and to easily control rr
+as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging pod.
+pygdbmi's abstraction simplifies programatically controlling rr.  A
+pod can receive a command from the client, pass it to rr, and respond
+without having to deal with parsing GDB/MI output or managing the rr
+process.
+
 \subsubsection{Socket.IO}\label{socketio}
@@
 namespaces \cite{socketio}.  The collaborative debugger uses the
-standard Javascript implementation of Socket.IO on the client side.
+standard JavaScript implementation of Socket.IO on the client side.
 Messages are passed through a server to individual debugging pods,
@@
 
-\subsubsection{Flask and Node}\label{flask/node}
+\subsubsection{MongoDB}\label{mongodb}
 
-\subsubsection{React and Xterm.js}\label{react}
+The collaborative debugger uses a database to store information about
+users, pods, and example debugging sessions.  Due to it's speed of
+deployment and natural interaction with the object-oriented languages
+used to create the project, MongoDB was chosen as database software
+\cite{mongodb}.
+
+\subsubsection{Flask and Node.js}\label{flask/node}
+
+The primary server for the collaborative debugger is split into two
+sections: a simple Node.js \cite{node} server that serves the frontend
+webapp, and an API server created using Flask \cite{flask}.  While in
+development, the builtin React (see next section) development server
+is used to serve the frontend.  This makes debugging the frontend far
+easier.
+\par
+
+An API server is necessary to authenticate users and to provide a
+means to create/delete debugging sessions.  Since the rest of the
+backend was created using Python, Flask was chosen to create the API
+server.  Flask is a lightweight web application framework which lends
+itself perfectly to interacting with the Python MongoDB and Kubernetes
+APIs.
+
+\subsubsection{React}\label{react}
+
+React is JavaScript library that simplifies creating user interfaces
+and managing state \cite{react}.  React's state management is of
+particular importance to the collaborative debugger's frontend.  State
+constantly changes as users create/delete debugging sessions, join
+existing sessions, and communicate with rr.  React allows classes to
+encapsulate components such as a list of existing debugging sessions,
+a view of the current program's source code, and the terminal
+interface with rr.  Instances of these classes maintain state
+and update efficiently.
+\par
+The frontend makes extensive use of JSX, syntax which allows the
+inclusion of segments of HTML code within a React app written in
+JavaScript.  This makes it easy for each component of the one-page
+webapp to hide/show subcomponents as state changes.
+
+\subsubsection{Monaco and Xterm.js}\label{xtermjs/monaco}
+
+After joining or creating a debugging session, users spend most of
+their time interacting collaboratively with rr.  Their primary
+interface to rr is through Xterm.js, a frontend component that makes
+it easy to emulate terminal behavior in the browser \cite{xtermjs}.
+With a few control methods, it is simple to provide a terminal
+interface to rr that is virtually indistinguishable from a local
+session.  By using the Xterm.js based interface, students can learn to
+use rr (and by extension gdb) collaboratively, and directly translate
+that knowledge to individual work.
+\par
+
+In addition to the terminal interface, the frontend shows a view of
+the current source file being debugged.  The Monaco Editor
+\cite{monaco} is used to display this source view.  Though more
+complex than is strictly necessary to display code, Monaco makes it
+easy to format and syntax-highlight.  Using Monaco also simplifies the
+future addition of editing source code, should the need arise.
+React's state management allows updating text in the editor as
+efficiently as possible.
 

### 2020-11-29 paper/paper.tex
Added overview section and most of cluster/pod/service creation section

@@
 \usepackage{cite}
-\usepackage{listings}
+
 \usepackage{graphicx}
@@
 
-\lstset{basicstyle=\ttfamily}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{listingsutf8}
+
+% From: https://gordonlesti.com/custom-code-highlighting-in-latex/
+
+\lstdefinelanguage{Dockerfile}
+{
+  morekeywords={FROM, RUN, CMD, LABEL, MAINTAINER, EXPOSE, ENV, ADD, COPY,
+    ENTRYPOINT, VOLUME, USER, WORKDIR, ARG, ONBUILD, STOPSIGNAL, HEALTHCHECK,
+    SHELL},
+  morecomment=[l]{\#},
+  morestring=[b]"
+}
+
+\newcommand\YAMLcolonstyle{\color{red}\mdseries}
+\newcommand\YAMLkeystyle{\color{black}\bfseries}
+\newcommand\YAMLvaluestyle{\color{blue}\mdseries}
+
+% From: https://tex.stackexchange.com/questions/152829/how-can-i-highlight-yaml-code-in-a-pretty-way-with-listings
+
+\makeatletter
+
+% here is a macro expanding to the name of the language
+% (handy if you decide to change it further down the road)
+\newcommand\language@yaml{yaml}
+
+\expandafter\expandafter\expandafter\lstdefinelanguage
+\expandafter{\language@yaml}
+{
+  keywords={true,false,null,y,n},
+  keywordstyle=\color{darkgray}\bfseries,
+  basicstyle=\YAMLkeystyle,                                 % assuming a key comes first
+  sensitive=false,
+  comment=[l]{\#},
+  morecomment=[s]{/*}{*/},
+  commentstyle=\color{purple}\ttfamily,
+  stringstyle=\YAMLvaluestyle\ttfamily,
+  moredelim=[l][\color{orange}]{\&},
+  moredelim=[l][\color{magenta}]{*},
+  moredelim=**[il][\YAMLcolonstyle{:}\YAMLvaluestyle]{:},   % switch to value style at :
+  morestring=[b]',
+  morestring=[b]",
+  literate =    {---}{{\ProcessThreeDashes}}3
+                {>}{{\textcolor{red}\textgreater}}1     
+                {|}{{\textcolor{red}\textbar}}1 
+                {\ -\ }{{\mdseries\ -\ }}3,
+}
+
+% switch to key style at EOL
+\lst@AddToHook{EveryLine}{\ifx\lst@language\language@yaml\YAMLkeystyle\fi}
+\makeatother
+
+\newcommand\ProcessThreeDashes{\llap{\color{cyan}\mdseries-{-}-}}
+
+\lstset{basicstyle=\linespread{0.5}\ttfamily,
+  showstringspaces=false,
+  commentstyle=\color{red},
+  keywordstyle=\color{blue},
+  inputencoding=utf8,
+  extendedchars=true
+}
+
 \graphicspath{{./images/}}
@@
   \begin{enumerate}
-  \item A pod for each debugging instance which runs the rr debugger
+  \item A Pod for each debugging instance which runs the rr debugger
     (\ref{rr}).  These communicate directly with users through
-    WebSocket server pods. (\ref{socketio}).
+    WebSocket server Pods. (\ref{socketio}).
   \item Pods running a frontend server written in Node.js which works
@@
 containers, like those created by Docker.  By bundling one or more
-closely linked containers into a ``pod'', Kubernetes is able to manage
+closely linked containers into a ``Pod'', Kubernetes is able to manage
 deployment and re-deployment of applications running inside
-containers.  It is trivial to create new pods, or to create multiple
-pods running the same application as needed within a Kubernetes
-cluster \cite{k8s}.  The speed at which even relatively large pods can be created
+containers.  It is trivial to create new Pods, or to create multiple
+Pods running the same application as needed within a Kubernetes
+cluster \cite{k8s}.  The speed at which even relatively large Pods can be created
 and the inherent security provided by containerization drove the
-decision to create a new pod on the fly for each debugging instance in
+decision to create a new Pod on the fly for each debugging instance in
 the collaborative debugger.
@@
 record a program's execution and design a debugging lesson which
-students can work on together.  The repeatablity of rr means that
+students can work on together.  The repeatability of rr means that
 students can focus on debugging, and teachers can create as specific
@@
 files, and doesn't force execution inside of a
-VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The tradeoff for these
+VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The trade off for these
 benefits are two major system limitations: rr is only compatible with
@@
 To parse rr output into Python dictionaries and to easily control rr
-as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging pod.
+as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
 pygdbmi's abstraction simplifies programatically controlling rr.  A
-pod can receive a command from the client, pass it to rr, and respond
+Pod can receive a command from the client, pass it to rr, and respond
 without having to deal with parsing GDB/MI output or managing the rr
@@
 To speed communication, the collaborative debugger uses WebSockets to
-directly connect web clients and the pods running rr.  Socket.IO is a
+directly connect web clients and the Pods running rr.  Socket.IO is a
 library that extends WebSockets.  It provides backup incase a
@@
 standard JavaScript implementation of Socket.IO on the client side.
-Messages are passed through a server to individual debugging pods,
+Messages are passed through a server to individual debugging Pods,
 both of which use the Python implementation of Socket.IO,
@@
 The collaborative debugger uses a database to store information about
-users, pods, and example debugging sessions.  Due to it's speed of
+users, Pods, and example debugging sessions.  Due to it's speed of
 deployment and natural interaction with the object-oriented languages
@@
 
-\subsection{Project Overview}
-
-TODO: a brief overview linking aspects of the project to background
-research discussed above.
-
 \section{Design}
 
-TODO: the big piece of writing, detailing components of the project,
-the system as a whole, and motivations for design choices.
+The collaborative debugger consists of a distributed Kubernetes
+backend and frontend React webapp.  Kubernetes was chosen for the
+backend primarally so that a Pod could be created dynamically for each
+debugging session.  The design of the backend is heavily distributed,
+allowing individual components to be modified without the whole system
+needing to be reconfigured.
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=.9]{detailed_system}
+  \centering
+  \caption{Detailed Overview of the Collaborative Debugger}
+  \label{debugger:detailedoverview}
+\end{figure}
+
+Each backend component of the collaborative debugger runs in it's own
+individual Pod.  There are two different classes of Pods in the
+collaborative debugger:
+
+\begin{enumerate}
+\item Statically created Pods.  These are the \textit{RR Message
+    Server Pods}, the \textit{Frontend/API Server Pods}, and the
+  \textit{Database Pod}.  This class of Pods are manually created when
+  the cluster that will run the backend is first initialized.  The RR
+  Message Server Pods and Frontend/API Server Pods may be created
+  using Deployments \cite{k8s_docs} to allow later scaling, where
+  multiple Pods running the same application may be created to
+  facilitate increased load.
+\item Dynamically created Pods.  This class of Pod contains the
+  individual instances of \textit{RR Debug Pods} that are created on
+  request by the API server.  When a client requests a new debug
+  session, the API server uses the Kubernetes API to create a new Pod
+  based on an existing template, gives the Pod a unique identifier,
+  and associates it in the database with the requesting client.
+\end{enumerate}
+
+These Pods communicate with other Pods in the cluster and with the
+outside world through Services.  The Kubernetes documentation defines
+a Service as ``an abstraction which defines a logical set of Pods and
+a policy by which to access them'' \cite{k8s_docs}.  In the
+collaborative debugger, these Services manifest as:
+
+\begin{enumerate}
+\item The \textit{Database ClusterIP}: a ClusterIP, which exposes the
+  Database Pod only inside the cluster.  The only component that makes
+  use of this ClusterIP is the API server, which uses it primarily to
+  communicate information about users and RR Debug Pods with the
+  database.
+\item The \textit{RR Message Load Balancer}: a Load Balancer, which
+  exposes the RR Message Server Pods to the outside world.  Using
+  Socket.IO, clients send commands to and recieve responses from
+  individual RR Debug Pods through the RR Message Server Pods.
+\item The \textit{Frontend Load Balancer}: another Load Balancer,
+  which exposes the Frontend/API Server Pods to the outside world.
+  Clients request the frontend webapp and send api requests/recieve
+  api responses through this Load Balancer.
+\end{enumerate}
+
+The frontend webapp dynamically updates as the user requests new debug
+sessions, issues commands to rr, and visits new source files.  A user
+can be part of multiple debug sessions simultaneously.  Each debug
+session is assigned at 5 digit identifier at creation, which is used
+to join sessions in progress.
+\par
+
+Each component of the backend and
+frontend will be discussed in depth in the following sections.
+
+\subsection{Configuration and Setup}
+
+Configuration and setup of the collaborative debugger is relatively
+simple.  After a Kubernetes cluster is created (this is made easier by
+using a Managed Kubernetes service) Pods and Services are created
+using various configuration files.  Services should be created first,
+so that Pods can be created which rely on access to Services in order
+to run.  The following sections outline the process of creating a
+cluster, Pods, and Services, with a focus on statically created Pods.
+The process of building images for building dynamically created Pods
+is identical, but the process of starting the Pods is more
+complicated.  This process will be discussed in-depth in the section
+on the API server \ref{api}.
+
+\subsubsection{Cluster Selection and Configuration}
+
+Though Docker and by extension Kubernetes aim to be largely
+platform-agnostic, the requirements of rr impose some restrictions on
+cluster setup and configuration.  Clusters, even those running inside
+a VM, must be run on machines using relatively modern Intel x86 CPUs
+(Nehalem and beyond).  The clusters must run on an operating system
+using Linux kernel version 3.11 or higher \cite{rr-repo}.  The reasons
+for these restrictions are discussed in the section on rr (\ref{rr}).
+Finally, in order for rr to be able to work efficiently, the
+\lstinline{kernel.perf_event_paranoid} parameter must be set to 1
+\cite{rr-repo}.  This should be done on every node in the cluster
+which will run RR Debug Pods.  For the purposes of development, it has
+been set to 1 on all nodes in the collaborative debugger cluster.
+
+\subsubsection{Creating Pods}
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=.9]{pod_creation}
+  \centering
+  \caption{The Pod Creation Process}
+  \label{podcreation:overview}
+\end{figure}
+
+Pods are created in five steps:
+
+\begin{enumerate}
+\item A \lstinline{Dockerfile} is used to build a new Docker container
+  image from various pieces of source code, scripts, and a base image
+  (such as the official MongoDB image or official Ubuntu image).  The
+  Dockerfile also contains instructions to install necessary packages,
+  run build scripts, and create file structures in the image.
+\item The Docker image is tagged and uploaded to a private container
+  registry.
+\item A Pod configuration schema is defined/updated with details of
+  the corresponding image's tag and any necessary Pod-specific
+  settings/startup commands.
+\item The Pod configuration schema is applied, either statically or
+  dynamically.  When the schema is applied, Kubernetes pulls the image
+  from the container registry and creates a new Pod according to the
+  schema, running any startup commands if provided.
+\item If Pod creation is successful, the result is a new Pod running
+  in the cluster.
+\end{enumerate}
+
+\paragraph{1. Docker Container Creation}
+
+Docker containers are created using a Dockerfile.  The Dockerfile used
+to create the container image for the RR Message
+Server is shown below.\\
+
+\begin{lstlisting}[language=Dockerfile,caption={RR Message Server Dockerfile},captionpos=b]
+# Base Image
+FROM  ubuntu:latest
+
+# Package Installation
+WORKDIR /tmp/
+ENV DEBIAN_FRONTEND="noninteractive"
+RUN apt-get update && apt-get install -y \
+python3-pexpect python3-pip
+
+# User Creation
+RUN useradd -ms /bin/bash rrserver
+USER rrserver
+
+# File structure creation/app setup
+RUN pip3 install requests python-socketio \
+eventlet
+WORKDIR /home/rrserver/
+RUN mkdir app
+WORKDIR /home/rrserver/app/
+COPY server.py .
+COPY startup.sh .
+
+# Startup command
+CMD ["sh", "startup.sh"]
+\end{lstlisting}
+
+The build process for each collaborative debugger Docker image follows
+the same structure as the one outlined in the Dockerfile above:
+
+\begin{enumerate}
+\item The base image is defined. The RR Message Server and RR Debug
+Pod images are based on the latest Ubuntu image.  This is particuarly
+necessary for the RR Debug Pod image, as rr's low-level nature
+facilitates frequent updates as changes are made to the Linux kernel.
+The Frontend/API Server image is based on the latest Node image, and
+the Database image is based on the latest MongoDB image.
+\item Second, any necessary packages are installed.  For the RR Debug
+  Pod image, rr is compiled from source and installed.
+\item A non-root user is created if necessary.
+\item Program files are copied over and a file structure is created.
+  Packages that don't rely on the base images builtin package manager
+  are installed here.
+\item A startup command is defined.
+\end{enumerate}
+
+Each line in a Dockerfile corresponds to a layer in the built image.
+This build order minimizes the amount of rebuilding necessary by
+placing the items that are most likely to change towards the end of
+the build process.
+
+\paragraph{2. Container Registry Upload}
+
+Most Managed Kubernetes services come with the option to create a
+private container registry.  With proper authentication, this allows
+Docker and Kubernetes to access user-created images as easily as if
+they were in a public registry.  Images built with Docker are uploaded
+to a private container registry for use in the collaborative debugger.
+
+\paragraph{3. Pod Configuration Schema}
+
+The Pod configuration schema for most Pods in the collaborative
+debugger is fairly generic.  It consists of a \lstinline{name}, an
+\lstinline{image} sourced from the container registry, and in the case
+of pods that need to interact with a load balancer, an
+\lstinline{app}.
+
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Pod Schema},captionpos=b]
+apiVersion: v1
+kind: Pod
+metadata:
+  name: rr-translation
+  labels:
+    purpose: translate-rr
+spec:
+  containers:
+  - name: rr-test-container
+    image: # Redacted
+    securityContext:
+      capabilities:
+        add:
+        - SYS_PTRACE
+  restartPolicy: OnFailure
+\end{lstlisting}
+
+A notable exception is the RR Debug Pod Schema, which adds the
+\lstinline{SYS_PTRACE} capability to the Pod.  This is necessary for
+rr to properly trace system calls.
+
+\paragraph{4 \& 5. Pod Creation}
+
+For statically created Pods, the \lstinline{kubectl apply} command is
+used to create new Pods.  Kubernetes pulls the container image defined
+in the schema from the container registry and starts the Pod with any
+necessary commands.  The database Pod is connected to a long-term
+storage volume at this time.  Upon successful creation, the Pod is
+ready to interact with any necessary load balancers.
+
+\subsection{Frontend/API Server}\label{api}
+
+We just need this in here for the ref for now. TODO.
 
@@
 
-TODO: write about additional/necessary unimplemented features (ATM
-this is looking like auth stuff, unless OAuth2 proves to be less of a
-pain in the ass then when I've used it before in Javascript.  Maybe
-some of the vis stuff as well, though I'd really like to get the
-basics done here.)  Write about desire to actually do research and see
-whether the tool helps students learn debugging.
+TODO
 

### 2020-11-30 paper/paper.tex
Tiny bit of code cleanup in rrtranslation, and paper stuff

@@
 \usepackage[doublespacing]{setspace}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage{titlesec}
@@
 \section{Introduction}
+
+The goal of this project is to create a collaborative debugger to aid
+in the teaching of debugging.  The collaborative debugger aims to help
+teachers integrate this undertaught skill into their classes.  It
+realizes the benefits of collaborative programming and teaching
+debugging by providing a platform that is genuinely useful to students
+and teachers.
+
 \subsection{Motivation}
@@
 programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.
+Teaching these debugging skills collaboratively will likely offer the
+same confidence and correctness benefits realized by teaching
+programming collaboratively
+\cite{10.1145/1026487.1008043}\cite{10.1145/1145287.1145293}.
 
@@
 curriculum considers debugging an essential ``transferable
-skill''\cite{10.1145/2602484}.\par There seems to be confidence that
-the problem-solving techniques used in debugging are widely
-applicable, but of greater interest to computer science teachers is
-whether teaching debugging directly benefits student programmers.
-Michaeli and Romeike conducted a good, albeit somewhat small, study on
-the efficacy of teaching a systematic debugging process to K12
-students.  They found that students who have been taught a specific
-debugging framework performed better in debugging tests and were more
-confident in their own debugging skills\cite{10.1145/3361721.3361724}.
-Their result is positive evidence towards the efficacy of teaching
-debugging, though it doesn't include college or university
-students.\par
+skill''\cite{10.1145/2602484}.\par
+
+There seems to be confidence that the problem-solving techniques used
+in debugging are widely applicable, but of greater interest to
+computer science teachers is whether teaching debugging directly
+benefits student programmers.  Michaeli and Romeike conducted a good,
+albeit somewhat small, study on the efficacy of teaching a systematic
+debugging process to K12 students.  They found that students who have
+been taught a specific debugging framework performed better in
+debugging tests and were more confident in their own debugging
+skills\cite{10.1145/3361721.3361724}.  Their result is positive
+evidence towards the efficacy of teaching debugging, though it doesn't
+include college or university students.\par
 
@@
 significant faults in current debugging teaching platforms.  The
-primary two which this project seeks to address are as follows:
-
-\begin{enumerate}
-\item A lack of back-tracing ability/coverage.
-\item A lack of tools addressing system knowledge (an understanding of
-  the program to be debugged).
-\end{enumerate}
+collaborative debugger primarily seeks to address the lack of
+back-tracing ability/coverage found in their research.
 
@@
 
-TODO: find some basic research that backs up the claim that
-collaborative/pair programming is worthwhile.
+Research into the value of collaborative programming is overwhelmingly
+positive.  McDowell et. al. found that not only does collaborative
+programming significantly boost student confidence and the retention
+of students in computer science majors, but that it demonstrably
+improves student's work \cite{10.1145/1145287.1145293}.  These
+benefits of confidence and correctness are present when paired and
+non-paired students are given identical assignments
+\cite{10.1145/1026487.1008043}, further indicating that the simple act
+of collaboration definitively benefits computer science students.\par
 
-\subsubsection{Tools that Enable Collaborative Programming}
+It seems that the benefits in confidence that result from teaching
+debugging should be magnified by teaching debugging collaboratively.
+There are similarities between the introductory nature of teaching
+fundemental debugging skills and the nature of teaching fundemental
+programming skills covered in the classes studied in
+\cite{10.1145/1026487.1008043} and \cite{10.1145/1145287.1145293}.
+Introductory programming classes typically use a specific programming
+language in order to introduce widely applicable programming
+principles.  By using a specific collaborative platform to introduce
+debugging principles, students may realize the same benefits in both
+confidence \textit{and} correctness that they do from pair programming
+in introductory comptuer science classes.\par
+
+The collaborative debugger aims to provide such a platform.
+
+\subsection{Existing Tools that Enable Collaborative Programming} \label{exisitingcollab}
 
 TODO: write about glitch, repl.it, etc.
-
-\subsubsection{Collaborative Programming and Teaching Debugging}
 
@@
 debuggers effectively.
-\pagebreak
 
-\subsection{Tools Used}
+\section{Tools Used}
 
@@
 
-\subsubsection{Kubernetes}\label{k8s}
+\subsection{Kubernetes}\label{k8s}
 
@@
 Pods running the same application as needed within a Kubernetes
-cluster \cite{k8s}.  The speed at which even relatively large Pods can be created
-and the inherent security provided by containerization drove the
-decision to create a new Pod on the fly for each debugging instance in
-the collaborative debugger.
+cluster \cite{k8s}.  The speed at which even relatively large Pods can
+be created and the inherent security provided by containerization
+drove the decision to create a new Pod on the fly for each debugging
+instance in the collaborative debugger.  This allows the debugger to
+provide a similar level of convinience to exisiting collaborative
+tools, such repl.it (\ref{exisitingcollab}).
 \par
@@
 
-\subsubsection{Mozilla's rr}\label{rr}
+\subsection{Mozilla's rr}\label{rr}
 
-\paragraph{Overview}
+\subsubsection{Overview}
 
@@
 students can focus on debugging, and teachers can create as specific
-examples as they please.
+examples as they please.  The use of rr is the most significant step
+the collaborative debugger takes to addressing the lack of
+back-tracing ability/coverage found in existing tools for teaching
+debugging\cite{10.1145/3286960.3286970}.
 
-\paragraph{Limitations}
+\subsubsection{Limitations}
 
@@
 
-\subsubsection{pygdbmi}
+\subsection{pygdbmi}
 
@@
 Pod can receive a command from the client, pass it to rr, and respond
-without having to deal with parsing GDB/MI output or managing the rr
-process.
+without having to deal with parsing GDB/MI output or with directly
+managing the rr process.
 
-\subsubsection{Socket.IO}\label{socketio}
+\subsection{Socket.IO}\label{socketio}
 
@@
 
-\subsubsection{MongoDB}\label{mongodb}
+\subsection{MongoDB}\label{mongodb}
 
@@
 
-\subsubsection{Flask and Node.js}\label{flask/node}
+\subsection{Flask and Node.js}\label{flask/node}
 
@@
 webapp, and an API server created using Flask \cite{flask}.  While in
-development, the builtin React (see next section) development server
+development, the builtin React (\ref{react}) development server
 is used to serve the frontend.  This makes debugging the frontend far
@@
 
-\subsubsection{React}\label{react}
+\subsection{React}\label{react}
 
@@
 
-\subsubsection{Monaco and Xterm.js}\label{xtermjs/monaco}
+\subsection{Monaco and Xterm.js}\label{xtermjs/monaco}
 
@@
 \item Dynamically created Pods.  This class of Pod contains the
-  individual instances of \textit{RR Debug Pods} that are created on
-  request by the API server.  When a client requests a new debug
-  session, the API server uses the Kubernetes API to create a new Pod
-  based on an existing template, gives the Pod a unique identifier,
-  and associates it in the database with the requesting client.
+  individual instances of \textit{RR Debug Session Pods} that are
+  created on request by the API server.  When a client requests a new
+  debug session, the API server uses the Kubernetes API to create a
+  new Pod based on an existing template, gives the Pod a unique
+  identifier, and associates it in the database with the requesting
+  client.
 \end{enumerate}
@@
   use of this ClusterIP is the API server, which uses it primarily to
-  communicate information about users and RR Debug Pods with the
+  communicate information about users and RR Debug Sessions with the
   database.
@@
   Socket.IO, clients send commands to and recieve responses from
-  individual RR Debug Pods through the RR Message Server Pods.
+  individual RR Debug Session Pods through the RR Message Server Pods.
 \item The \textit{Frontend Load Balancer}: another Load Balancer,
@@
 
-Each component of the backend and
-frontend will be discussed in depth in the following sections.
+Each component of the backend and frontend will be discussed in depth
+in the following sections.
 
@@
 using various configuration files.  Services should be created first,
-so that Pods can be created which rely on access to Services in order
-to run.  The following sections outline the process of creating a
+so that Pods which rely on access to Services function properly on
+creation.  The following sections outline the process of creating a
 cluster, Pods, and Services, with a focus on statically created Pods.
 The process of building images for building dynamically created Pods
-is identical, but the process of starting the Pods is more
-complicated.  This process will be discussed in-depth in the section
-on the API server \ref{api}.
+is similar, but the process of starting the Pods is more complicated.
+This process will be discussed in-depth in the section on the API
+server (\ref{api}).
 
@@
 
-Though Docker and by extension Kubernetes aim to be largely
-platform-agnostic, the requirements of rr impose some restrictions on
-cluster setup and configuration.  Clusters, even those running inside
-a VM, must be run on machines using relatively modern Intel x86 CPUs
-(Nehalem and beyond).  The clusters must run on an operating system
-using Linux kernel version 3.11 or higher \cite{rr-repo}.  The reasons
-for these restrictions are discussed in the section on rr (\ref{rr}).
-Finally, in order for rr to be able to work efficiently, the
+Though Kubernetes aims to be largely platform-agnostic, the
+requirements of rr impose some restrictions on cluster setup and
+configuration.  Clusters, even those running inside a VM, must be run
+on machines using relatively modern Intel x86 CPUs (Nehalem and
+beyond).  The clusters must run on an operating system using Linux
+kernel version 3.11 or higher \cite{rr-repo}.  Finally, in order for
+rr to be able to work efficiently, the
 \lstinline{kernel.perf_event_paranoid} parameter must be set to 1
 \cite{rr-repo}.  This should be done on every node in the cluster
-which will run RR Debug Pods.  For the purposes of development, it has
-been set to 1 on all nodes in the collaborative debugger cluster.
+which will run RR Debug Sessions.  For the purposes of development, it
+has been set to 1 on all nodes in the collaborative debugger cluster.
 
@@
 to create the container image for the RR Message
-Server is shown below.\\
+Server is shown below:\\
 
@@
 \item The base image is defined. The RR Message Server and RR Debug
-Pod images are based on the latest Ubuntu image.  This is particuarly
-necessary for the RR Debug Pod image, as rr's low-level nature
-facilitates frequent updates as changes are made to the Linux kernel.
-The Frontend/API Server image is based on the latest Node image, and
-the Database image is based on the latest MongoDB image.
+  Session images are based on the latest Ubuntu image.  This is
+  particuarly necessary for the RR Debug Session image, as rr's
+  low-level nature necessitates somewhat frequent updates as changes
+  are made to the Linux kernel.  The Frontend/API Server image is
+  based on the latest Node image, and the Database image is based on
+  the latest MongoDB image.
 \item Second, any necessary packages are installed.  For the RR Debug
@@
   Packages that don't rely on the base images builtin package manager
-  are installed here.
+  are installed at this time.  In the example above, these include
+  Python packages.
 \item A startup command is defined.
@@
 of pods that need to interact with a load balancer, an
-\lstinline{app}.
+\lstinline{app}.\\
 
-\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Pod Schema},captionpos=b]
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Session Schema},captionpos=b]
 apiVersion: v1
@@
   - name: rr-test-container
-    image: # Redacted
+    image: example-container-registry.com/sproj/...
     securityContext:
@@
 
-A notable exception is the RR Debug Pod Schema, which adds the
+A notable exception is the RR Debug Session Schema, which adds the
 \lstinline{SYS_PTRACE} capability to the Pod.  This is necessary for
@@
 
+\subsubsection{Service Creation}
+
+Services are the first components of the collaborative debugger
+created after the cluster is initiallized.  The three Services used by
+the debugger are all statically created.  Like statically created
+Pods, schemas that define Services are applied manually using the
+\lstinline{kubectl apply} command.\\
+
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={Frontend Load Balancer Schema},captionpos=b]
+apiVersion: v1
+kind: Service
+metadata:
+  name: frontend-load-balancer
+spec:
+  selector:
+    app: frontend
+  type: LoadBalancer
+  ports:
+    - port: 3000
+    targetPort: 3000
+\end{lstlisting}
+
+The \lstinline{app} field in the above schema corresponds to the
+\lstinline{app} field defined in the metadata of the Frontend/API
+Server configuration schema.  Traffic to the Load Balancer's external
+IP address on port 3000 is redirected automatically by Kubernetes to
+any Pod running the Frontend/API Server.  The Load Balancer determines
+which Pod is the most suitable given current demands on the system.
+
+\subsection{RR Debug Sessions \& RR Message Server}
+
+At the heart of the collaborative debugger is the RR Debug Session.
+Every time a user wants to debug a new program execution, an RR Debug
+Session Pod is dynamically created by the API server.  The new Pod is
+assigned a unique five digit identification number when it is created.
+This five digit number, is used as the channel for the RR Debug
+Session, seperating its communication from other RR Debug Session
+Pods.  Clients, RR Message Server Pods, and the RR Debug Session Pod
+connect through the RR Message Load Balancer.  Clients and RR Debug
+Sessions send messages using the channel assigned to the pod at
+creation time.  A diagram of one full communication cycle between two
+clients and an RR Debug Session Pod is outlined in the diagram below:
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=1]{rr_detailed}
+  \centering
+  \caption{RR Message Server Communication Cycle}
+  \label{rr:detailed}
+\end{figure}
+
+The process of sending a command to an RR Debug Session and receiving
+a response is as follows:
+
+\begin{enumerate}
+\item Client A, already having connected to the channel that
+  corresponds to it's current RR Debug Session (channel 12345 in the
+  example), sends an rr command.
+\item The RR Message Server recieves the command and emits a message
+  that RR Debug Session Pods are equiped to recieve on the same
+  channel.  In practice, since only one RR Debug Session Pod is ever
+  on a channel, this equates to emmiting a message to the Pod
+  directly.
+\item The RR Debug Session Pod recieves the message, and passes the
+  command to its instance of RRInterface.  When it recieves a
+  response, it emits the output from rr along with other debugging
+  information and the channel.
+\item The RR Message Server recieves the response, and emits a
+  response message that clients are equipped to recieve on the
+  corresponding channel.
+\item Clients A and B are connected to the corresponding channel, so
+  they both recieve the response.
+\end{enumerate}
+
+\subsubsection{RR Message Server}
+
+The RR Message server is remarkably simple, consisting of just 3
+important functions:\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Message Server},captionpos=b]
+@sio.on('join_channel')
+def join_channel(sid, data):
+    sio.enter_room(sid, data['channel'])
+
+@sio.on('rr_command')
+def on_rr_command(sid, data):
+    data['sid'] = sid
+    try:
+        sio.emit('rr_command', data,
+                 room=data['channel'])
+    except:
+        pass
+
+@sio.on('rr_response')
+def on_rr_response(sid, data):
+    sio.emit('rr_response', data,
+             room=data['channel'])
+\end{lstlisting}
+
+Even in a fully production ready version of the collaborative debugger
+with more security features enabled, the RR Message Server is unlikely
+to become much more complex.  It exists purely to pass messages
+between clients and RR Debug Session Pods and to manage channels.
+Since the design of other aspects of the collaborative debugger ensure
+that all members of a given channel exist only during the lifetime of
+it's corresponding RR Debug Session Pod, the
+\lstinline{'join_channel'} event handler simply adds socket.io clients
+to a specified channel on request.
+\par
+
+The two message processing functions, \lstinline{on_rr_command} and
+\lstinline{on_rr_response}, are equally simple.  When the RR Message
+Server recieves an \lstinline{'rr_command'} message, it passes the
+message data along to the corresponding RR Debug Session Pod by
+emitting an \lstinline{'rr_command'} namespaced to the correct channel
+with \lstinline{room=data['channel']}.  Since the socket.io client
+running in RR Message Server Pods has an event handler defined for
+\lstinline{'rr_command'}s but frontend clients do not, only RR Message
+Server Pods recieve the command.  The same process happens in reverse
+for \lstinline{'rr_response'}s, with only the frontend socket.io
+clients having a handler defined for \lstinline{'rr_response'}.
+
+\subsubsection{RR Debug Session}
+
+\paragraph{Building Example Pods}
+
+The process for building RR Debug Session example images is differs
+slightly from other images used by the collaborative debugger.
+Currently, three example RR Debug Session images are availiable for
+use.  Example container images are based on the primary RR Debug
+Session image, RR Translation.  The build process for this image,
+based on the lasted official Ubuntu container image, installs rr,
+python, and all necessary packages as well as the app that will run on
+the final image, \lstinline{rrtranslation.py}.\\
+
+\begin{lstlisting}[language=Dockerfile,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
+FROM example-container-registry.com/sproj/...
+
+WORKDIR /home/debug/app/
+COPY hash.c .
+COPY names .
+COPY startup.sh .
+\end{lstlisting}
+
+The Dockerfile shown above is for a simple hash program RR Debug
+Session example.  The build process copies over any necessary files,
+as well as a startup script:\\
+
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Example Startup Script},captionpos=b]
+gcc -g -o hash hash.c
+rr record ./hash
+python3 rrtranslation.py $1
+\end{lstlisting}
+
+This script is used to start the \lstinline{'rrtranslation.py'} program
+when the Pod is dynamically created.  The script will be passed the
+Pod's channel as argument \lstinline{$1} by the Kubernetes API on
+startup.
+\par
+
+Current example pod startup scripts compile the program to be debugged
+and record execution when the pod is created.  It is easy to create a
+more specific example to be replayed, if exact reproduction of
+syscalls or other non-deterministic behavior is desired.  The
+example's creator can record an rr session locally ahead of building
+the example Pod and copy the recording into the appropriate folder
+during the build process.  Since rr recordings are portable, the
+example will replay without issue in the RR Debug Session Pod.
+
+\paragraph{The RR Translation Program}
+
+The RR Translation program consists of two main components: a socket.io
+client, \lstinline{sio}, and an instance of the
+\lstinline{RRInterface} class, \lstinline{rri}, that controls the rr
+subprocess and parses interactions.  Simple code to read information
+about the current source file currently exists within
+\lstinline{sio}'s \lstinline{'rr_command'} event handler, but should
+be broken out into its own class if more complexity is added.
+\par
+
+The RR Translation program's socket.io client instance begins by
+connecting to the RR Message Server.  It immediately emits a
+\lstinline{'join_channel'} message, using the channel passed in from
+the Kubernetes API call on it's creation. \\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Translation Main},captionpos=b]
+if __name__ == '__main__':
+    channel = sys.argv[1]
+    sio.connect('http://rr-message-server-load-balancer:8000')
+    sio.emit('join_channel', {'channel': channel})
+\end{lstlisting}
+
+After joining the appropriate channel, \lstinline{sio} waits to
+recieve an \lstinline{'rr_command'}.  The first part of
+\lstinline{sio}'s \lstinline{'rr_command'} event handler is shown below:\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Command Event Handler},captionpos=b]
+@sio.on('rr_command')
+def on_rr_command(data):
+
+    body = {'from': data['sid'],
+            'command': data['command'],
+            'channel': channel}
+
+    try:
+        response = {'output': rri.write(data['command'])}
+\end{lstlisting}
+
+The handler first builds part of the response body, passing back data
+about the channel and originator of the message that will be important
+for the RR Message Server and client later.  Next, it calls the
+function necessary to pass a message to rr and recieve a response,
+\lstinline{rri.write()}.
+\par
+
+When the program starts, an instance of the \lstinline{RRInterface}
+class, \lstinline{rri}, is initialized in the same scope as
+\lstinline{sio}.  \lstinline{RRInterface} contains an instance of the
+\lstinline{GDBController} class from pygdbmi to interface with rr and
+control the rr subprocess, as well as a collection of methods to parse
+rr output, the three most important of which are shown below: \\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RRInterface},captionpos=b]
+def get_full_rr_response(self, command):
+    response = self.gdbmi.write(command)
+    while(not(self.end(response) or self.exited(response))):
+        response.extend(self.get_rr_response())
+    return response
+
+def write(self, command):
+    if self.command_forbidden(command):
+        raise DissallowedError
+    self.timeline.append(command)
+    return self.console_output(
+        self.get_full_rr_response(command))
+
+def source(self):
+    f = self.get_full_rr_response(
+        '-file-list-exec-source-file')[0]['payload']
+    return {'file': f['file'],
+            'line': f['line'],
+            'path': f['fullname']}
+\end{lstlisting}
+
+When \lstinline{write} is called, it determines if the command to be
+executed is forbidden.  Currently only one gdb command,
+\lstinline{shell}, is disallowed.  Parsing the output from
+\lstinline{shell} is difficult, it is of virtually no use since
+\textit{recordings}, not currently executing programs that
+\lstinline{shell} could effect are being debugged, and it opens the
+door to security issues.  Though containerization means that the most
+users could probably do is render their own debug session useless, the
+downsides of \lstinline{shell} far outweigh the benefits.
+\par
+
+Next, write appends the command to the \lstinline{RRInterface}'s
+\lstinline{timline} instance variable.  \lstinline{timeline} is a list
+of all commands the debug session has executed.  Though unused at the
+moment, it can serve in the future to impliment a shared history
+between all clients and to facilitate the saving of debug sessions
+in-progress.  By issuing all commands in \lstinline{timeline} to a new
+instance of \lstinline{RRInterface} with the same recording, it is
+possible to restore a debug session to an identical previous state.
+To save a debug session, the recording and \lstinline{timeline} can be
+stored in a database, and be used to initialize a new Pod with
+identical state to a previous RR Debug Session Pod.  This appears to
+clients as a seemless restoration of a previous Debug Session.
+\par
+
+Since the frontend currently consists of a terminal interface to rr
+and a view of the current source code, write does not need to return
+any information from rr other than console output.  Future versions of
+the collaborative debugger that support visualization tools could use
+gdbmi commands such as \lstinline{-stack-list-frames} to retrieve
+information to be used by frontend visualization tools.
+\lstinline{console_output} extracts user-relevant output from the
+dictionary returned by \lstinline{GDBController}'s \lstinline{write}
+method.  The \lstinline{get_full_rr_response} method ensures that all
+relavant output has been recieved from the rr subprocess before
+returning.  Write reduces the lengthy dictionary that would be
+produced by a command as simple as \lstinline{break main} into a few
+lines of user-relevant output.
+\par
+
+After the \lstinline{'rr_command'} event handler has recieved a
+response from \lstinline{rri}, it calls \lstinline{RRInterface}'s
+\lstinline{source} function to retrieve information about the current
+source file being debugged.  Most often the source file has not
+changed, and the only relevant piece of information that needs to be
+passed back to the client is the new line number in the source file.
+If the source file has changed, the handler reads its contents,
+updates \lstinline{current_file}, and emits the line number, file
+name, and contents.  If an error occurs at any point in the process, a
+detailed trace is emmitted for debugging purposes.  The trace should
+be ommited in production.
+\par
+
+Most of the complexity in the RR Translation program stems from
+parsing rr output.  The flow of data in the program is quite simple: a
+command is recieved, it is passed to rr, rr's response is processed,
+and a response message is emmited.  This flow would remain unchanged
+even with the addition of functions to retrieve information for data
+visualization.  Since the process of turning rr responses into a
+terminal interface takes place in the frontend, rather than the RR
+Translation program providing some sort of REPL over WebSockets,
+updates can be made to the frontend webapp without requiring backend
+changes.  The whole process is also quite snappy, with only a slight
+delay from the client's perspective compared to running a debugger
+natively.
+\par
+
+Effort is taken in the design of the RR Debug Session and RR Message
+Server to ensure students recieve a pleasant and near-identical
+experience to debugging locally in the future.  The aim is to provide
+students with an experience they can directly reference when using rr
+or gdb in the future, as well as one which is seamless to encourage
+the acquisition of debugging skills.  The system of inheritance used
+to create example debug session pods is designed to be as simple as
+possible, though a more fully featured frontend could simplify this
+process further in the future.
+
 \subsection{Frontend/API Server}\label{api}
@@
 We just need this in here for the ref for now. TODO.
+
+\subsection{Webapp}\label{webapp}
+
+TODO
 

### 2020-11-30 containers/rr-translation/rrtranslation.py

@@ -70,6 +70,11 @@ rri = RRInterface()
 channel = None
 current_source = ''
 
+def read_source(path):
+    with open(path, 'r') as f:
+        text = f.read()
+        return text
+
 @sio.event
 def connect():
     print("Connected.")
@@ -95,12 +100,11 @@ def on_rr_command(data):
             # If it's different, read the contents of the file, and
             # send that along with the file name and line number
             try:
-                with open(s['path'], 'r') as f:
-                    text = f.read()
-                    # current_source = s['path']
-                    response['source'] = {'file_name': s['file'],
-                                          'current_line': s['line'],
-                                          'contents': text}
+                text = read_source(s['path'])
+                current_source = s['path']
+                response['source'] = {'file_name': s['file'],
+                                      'current_line': s['line'],
+                                      'contents': text}
             except:
                 response['source'] = {'file_name': None,
                                       'current_line': '0',

### 2020-12-02 paper/paper.tex
Touch of program cleanup discovered when writing the paper, and lots of writing!

@@
 
-\paragraph{3. Pod Configuration Schema}
+\paragraph{3. Pod Configuration Schema} \label{podschema}
 
@@
 
-\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Session Schema},captionpos=b]
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Debug Session Schema},captionpos=b]
 apiVersion: v1
@@
 
-\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={Frontend Load Balancer Schema},captionpos=b]
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend Load Balancer Schema},captionpos=b]
 apiVersion: v1
@@
 
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Message Server},captionpos=b]
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Message Server},captionpos=b]
 @sio.on('join_channel')
@@
 
-\paragraph{Building Example Pods}
+\paragraph{Building Example Pods} \label{buildingchannel}
 
@@
 
-\begin{lstlisting}[language=Dockerfile,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
+\begin{lstlisting}[language=Dockerfile,basicstyle=\small,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
 FROM example-container-registry.com/sproj/...
@@
 
-\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Example Startup Script},captionpos=b]
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Example Startup Script},captionpos=b]
 gcc -g -o hash hash.c
@@
 
-\paragraph{The RR Translation Program}
+\paragraph{The RR Translation Program} \label{joiningchannel}
 
@@
 
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Translation Main},captionpos=b]
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Translation Main},captionpos=b]
 if __name__ == '__main__':
@@
 
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Command Event Handler},captionpos=b]
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Command Event Handler},captionpos=b]
 @sio.on('rr_command')
@@
 
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RRInterface},captionpos=b]
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RRInterface},captionpos=b]
 def get_full_rr_response(self, command):
@@
 
-We just need this in here for the ref for now. TODO.
+The second major backend system consists of the Frontend/API Server
+and database.  This system is responsible for serving the frontend
+webapp, as well as processing all API requests, such as those to
+authenticate users and create new RR Debug Sessions.  The builtin
+React development server is used to serve the frontend webapp and
+redirect API requests to the Flask API server during development.  For
+both security and performance reasons this should be changed to a
+combination of a more robust solution like Nginx \cite{nginx} and a
+production suitable web server for production.  The relationships
+between components and overall process will remain unchanged after
+this migration.
+\par
+
+Since the entire webapp is one page, the process for serving it is
+completely standard.  The server recieves a request for the index page
+of the website, and returns the compiled React webapp that makes up
+the homepage.  The process for API requests is more involved, and an
+example is outlined below:
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=1]{login_request}
+  \centering
+  \caption{API Request Process---Login}
+  \label{rr:detailed}
+\end{figure}
+
+\begin{enumerate}
+\item The client makes a POST request to the \lstinline{'/login'} URL.
+  This request contains the necessary data to log in the user.
+\item The frontend server forwards the request of a URL it does not
+  recognize to the API server running in the same Pod on a different
+  port.
+\item The API server recognizes the \lstinline{'/login'} route, and
+  communicates with the database to attempt to log in the user.
+\item The API server returns the result of the login request, which is
+  returned by the frontend server to the client.
+\end{enumerate}
+
+\subsubsection{Frontend/API Server Configuration}
+
+The Frontend/API Server uses Yarn \cite{yarn} to manage packages.
+When a Frontend/API Server Pod is deployed, the Pod's startup script
+uses the \lstinline{yarn start} and \lstinline{yarn start-api}
+commands to start the frontend server and API server respectively.
+These scripts are defined in the server's \lstinline{package.json}
+configuration file:\\
+
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
+  "start": "react-scripts start",
+  "start-api": "cd api && flask run --no-debugger",
+  .
+  .
+  .
+  "proxy": "http://localhost:5000",
+\end{lstlisting}
+
+Also of note in \lstinline{package.json} is the instruction to proxy
+the API server, which is running on port 5000.  This achieves the
+automatic forwarding of API requests to the API server described
+above.  Since API requests are proxied through the same address
+serving the frontend webapp, there are no issues with cross-origin
+requests.
+
+\subsubsection{The API Server}
+
+The API server is implimented using Flask.  The server program
+consists of handlers for the various API routes and instances of two
+classes which communicate with the database/Kubernetes API:
+\lstinline{TranslationPodManager} and \lstinline{UserManager}.  The
+full structure of the API Server program, \lstinline{api.py} is shown
+below:
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=1]{api_structure}
+  \centering
+  \caption{Structure of the API Server}
+  \label{rr:detailed}
+\end{figure}
+
+Care was taken to abstract as much as possible in the design of the
+API server and the classes it instantiates.  The two classes
+instantiated by \lstinline{api.py}, \lstinline{TranslationPodManager}
+and \lstinline{UserManager}, each instantiate another class which
+interacts with the database, \lstinline{DatabaseController}, as an
+instance variable.  \lstinline{TranslationPodManager} further
+instantiates the Kubernetes Core V1 Python API to communicate with the
+cluster in order to create/delete pods.  This process of abstraction
+ensures that the database, underlying cluster structure, and
+authentication methods can all change without significant changes
+needing to be made to \lstinline{api.py}.
+\par
+
+The most significant improvement that could be made to this structure
+would be to break the \lstinline{UserManager} and
+\lstinline{TranslationPodManager} instances out into seperate programs
+running on their own Pods in the cluster.  This would allow changes to
+be made in those classes, say to update the process of deleting Pods,
+without needing to update the entire Frontend/API Server Pod.  For the
+time being, the relative simplicity of the API means that the added
+complexity and overhead of implementing a method to communicate
+between dedicated \lstinline{TranslationPodManager} and
+\lstinline{UserManager} Pods and Pods running the Frontend/API Server
+doesn't seem worthwhile.  This may change in the future.
+\par
+
+Currently supported API requests are as follows:
+
+\begin{enumerate}
+\item \lstinline{'/login'}: Logs in the user given by
+  \lstinline{'name'}, or, (given the rather lax development security)
+  creates a user if none matching \lstinline{'name'} exists.  Returns
+  the name of the user.
+\item \lstinline{'/channel'}: How users join RR Debug Sessions.  Binds
+  the user given by \lstinline{'name'} to an RR Debug Session Pod
+  matching \lstinline{'channel'} in the database, if such a Pod
+  exists.  Returns the channel.
+\item \lstinline{'/pods'}: How the client gets a list of RR Debug
+  Session Pods the user is currently participating in.  Returns the
+  channels for all RR Debug Session Pods who the user, given by
+  \lstinline{'name'}, is currently bound to in the database.
+\item \lstinline{'/examples}: How the client gets a list of example RR
+  Debug Session Pod names.  Returns all names of example RR Debug
+  Session Pods that exist in the database.
+\item \lstinline{'/new'}: Creates a new RR Debug Session Pod based on
+  the name given by \lstinline{'program'}.  Binds the user, given by
+  \lstinline{'name'}, to the Pod.  Returns the channel of the Pod.
+  This request is the most complex, and is covered in more detail
+  below.
+\item \lstinline{'/delete'}: Deletes a binding between the user given
+  by \lstinline{'name'} and the RR Debug Session Pod given by
+  \lstinline{'channel'}.  If there are no bindings left between the
+  Pod and users (if all the users have left the debug session),
+  deletes the Pod from the database and the cluster.  Returns
+  \lstinline{True}.
+\end{enumerate}
+
+All the API request handlers listed above return an error message in
+the case of an error occuring during the processing of an API request.
+Since most errors that could occur would either be the result of
+unanticipated issues on the part of the user, or of some unforseen
+catostrophic internal error, informing users of error specifics would
+be unhelpful.  Some \lstinline{TranslationPodManager} and
+\lstinline{UserManager} functions throw specific errors in the event
+of duplicate usernames, channels that do not exist, etc.  These errors
+are caught and meaningful error messages are returned to the frontend
+webapp, which can then pass them on to the user.
+\par
+
+To examine the process of processing an API request, it makes sense to
+look at the most complex example, creating a new RR Debug Session Pod.
+This example shows the process for processing an API POST request,
+interacting with the database, and dynamically creating a Pod.  All
+API requests follow a similar procedure.
+
+\paragraph{Processing a POST Request---Extracting Information}
+
+Below is the route handler for the \lstinline{'/new'} URL, as well as
+the instantiation of \lstinline{TranslationPodManager} and
+\lstinline{UserManager}.\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={API Server New RR Debug Session Event Handler},captionpos=b]
+tpm = podmanager.TranslationPodManager(
+                 url='mongodb://database-load-balancer',
+                 port=27017)
+um = usermanager.UserManager(
+                 url='mongodb://database-load-balancer',
+                 port=27017)
+app = Flask(__name__)
+
+@app.route('/new', methods=['POST'])
+def new():
+    try:
+        name = request.get_json()['name']
+        program = request.get_json()['program']
+        channel = tpm.create_pod([name], program)
+        return {'channel': channel}
+    except:
+        return {'error': 'Internal Error'}
+\end{lstlisting}
+
+All API routes only support POST requests.  The route handler first
+extracts the relavant information from the POST request body, in this
+case \lstinline{name} and \lstinline{program}.  \lstinline{name}
+always corresponds the user who is making the request's username, and
+\lstinline{program} corresponds to the name of the example RR Debug
+Session image to be used.
+\par
+
+\lstinline{new} then calls \lstinline{TranslationPodManager}'s
+\lstinline{create_pod} function to create a new RR Debug Session Pod.
+\lstinline{create_pod} can bind multiple users to a Pod when it is
+created, so \lstinline{name} is passed inside a list.
+
+\paragraph{The Database} Before drilling into \lstinline{create_pod},
+it may be helpful to examine the MongoDB database which
+\lstinline{TranslationPodManager} and \lstinline{UserManager} interact
+with.  The database consists of three collections: \lstinline{users},
+\lstinline{pods}, and \lstinline{examples}.  Below is an example
+record from each collection:
+
+\begin{figure}[h!]
+
+  \includegraphics[scale=1]{database_example}
+  \centering
+  \caption{Database Record Examples}
+  \label{rr:detailed}
+\end{figure}
+
+\lstinline{pods} and \lstinline{users} have a many-to-many
+relationship.  Each user can be bound to an unlimited number of pods,
+and vice versa.  The \lstinline{DatabaseController} class includes
+functions to efficiently retrieve \lstinline{users} and
+\lstinline{pods} given one another, and to create/destroy bindings
+between \lstinline{users} and \lstinline{pods}.
+\par
+
+\lstinline{examples} are independent of \lstinline{users} and
+\lstinline{pods}.  The continer registry URLs of RR Debug Session
+example images are stored in the database so that new images can be
+easily added to the collaborative debugger upon creation.
+\par
+
+MongoDB keeps \lstinline{_id}s unique.  In addition, the collaborative
+debugger defines uniqueness constrains on \lstinline{users:name},
+\lstinline{pods:channel}, and \lstinline{examples:name} when the
+database is created.
+
+% \begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
+% users:            pods:               examples:
+%   _id ObjectId      _id ObjectId        _id ObjectId
+%   name String       channel String      name String
+%   pods Array        users Array         container String
+
+% Example user:
+% {
+% 	"_id" : ObjectId("5fc14c622dc45448d67189b3"),
+% 	"name" : "sammy",
+% 	"pods" : [
+% 		ObjectId("5fc14d5b2dc45448d67189b4"),
+% 		ObjectId("5fc59acb2dc45448d67189b6")
+% 	]
+% }
+
+% Example pod:
+% {
+% 	"_id" : ObjectId("5fc14d5b2dc45448d67189b4"),
+% 	"channel" : "19452",
+% 	"users" : [
+% 		ObjectId("5fc14c622dc45448d67189b3"),
+% 		ObjectId("5fc317d22dc45448d67189b5")
+% 	]
+% }
+
+% Example example:
+% {
+% 	"_id" : ObjectId("5fc14d50b92a752776d3fd7d"),
+% 	"name" : "hash",
+% 	"container" : "example-container-registry.com/sproj/..."
+% }
+% \end{lstlisting}
+
+\paragraph{Processing a POST Request---Dynamic Pod Creation}
+
+To dynamically create a Pod, \lstinline{create_pod} gets the container
+registry URL of the image to base the Pod on, creates the new Pod,
+binds users to the Pod, and returns the channel the Pod was created
+with.  The first part of the function is shown below:\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 1},captionpos=b]
+def create_pod(self, names, program):
+  examples = self.get_examples()
+  image = None
+  try:
+    image = examples[program]
+  except:
+    raise NoSuchExampleError
+\end{lstlisting}
+
+\lstinline{create_pod} calls \lstinline{TranslationPodManager}'s
+\lstinline{get_examples} method to get a list of RR Debug Session
+example image names and container registry URLs from the database.  In
+the event that the user has somehow pased a spurious image name or has
+passed a name when there are no images, an exception is raised.
+Otherwise, a RR Debug Session Pod schema is created using the image:\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 2},captionpos=b]
+  dep = {'apiVersion': 'v1',
+         'kind': 'Pod',
+         'metadata': {'labels':
+                      {'purpose': 'translate-rr'}},
+         'spec': {'containers':
+                  [{'image': image,
+                    'name': 'rr-test-container',
+                    'command': ['sh'],
+                    'args': ['startup.sh'],
+                    'securityContext':
+                    {'capabilities':
+                     {'add': ['SYS_PTRACE']}}}],
+                    'restartPolicy': 'OnFailure'}}
+\end{lstlisting}
+
+This schema, represented as a dictionary in Python, is identical to
+the YAML representation of the RR Debug Session Pod creation schema
+shown in (\ref{podschema}) with the exception of \lstinline{args}
+which will be updated further in a later step.  Finally,
+\lstinline{create_pod} does the
+work necessary to dynamically create a Pod:\\
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 3},captionpos=b]
+  channel = self.dbc.add_pod(
+            self.dbc.get_userids_by_name(names))
+  dep['metadata']['name'] = channel
+  dep['spec']['containers'][0]['args'].append(channel)
+  resp = self.v1.create_namespaced_pod(
+         body=dep, namespace='default')
+  # Wait to return the channel until the pod is live
+  # and read to recieve incomming communication
+  status = self.v1.read_namespaced_pod_status(
+           channel, 'default').status.container_statuses
+  while status == None or status[0].state.running == None:
+    time.sleep(1)
+    status = self.v1.read_namespaced_pod_status(
+             channel, 'default').status.container_statuses
+  return channel
+\end{lstlisting}
+
+First, \lstinline{DatabaseController}'s \lstinline{add_pod} method is
+called to insert a new \lstinline{pods} record into the database.
+\lstinline{add_pod} randomly generates a new five digit
+\lstinline{channel} for the Pod, using the uniqueness constraint
+imposed on \lstinline{pods:channel} to ensure an unused
+\lstinline{channel} is generated.  If no channels are open an error is
+raised.  Otherwise, the users passed to \lstinline{add_pod} are bound
+to the new Pod in the database, and \lstinline{channel} is returned.
+This process has the effect of limiting the number of simultaneous RR
+Debug Sessions that can be in use at the same time to 89999, and of
+slowing as more channels are used.  Given the few current users, the
+likelyhood of a collision when generating a new channel is low enough
+that a more advanced method is unnecessary.
+\par
+
+Next, the name of the container is updated to \lstinline{channel}.  In
+addition, the startup arguments are updated to include
+\lstinline{channel}.  This is how \lstinline{channel} is passed to the
+\lstinline{startup.sh} script and eventually used to join a channel on
+the RR Message Server in \lstinline{rrtranslation.py}
+(\ref{buildingchannel} \& \ref{joiningchannel}).
+\par
+
+The Kubernetes API is then finally used to create the Pod.  The Pod is
+created in the default namespace.  This could be changed to logically
+isolate dynamically and statically created Pods with minimal hassle,
+since Kubernetes' DNS resolution can connect pods accros namespaces.
+If Pod creation fails, an error will be thrown.  Otherwise, a loop is
+used to wait until \lstinline{rrtranslation.py} is running inside the
+Pod.  Once the Pod is ready, \lstinline{channel} is returned.
+
+\paragraph{Processing a POST Request---Returning a Response}
+
+Most request handlers, \lstinline{'/new'} included, simply return a
+JSON serializable dicitionary containing whatever results the API
+Server's instance of \lstinline{TranslationPodManager} or
+\lstinline{UserManager} returned, or an error message.  The
+\lstinline{'/new'} handler returns \lstinline{'channel': channel} so
+that the frontend webapp's socket.io client can join the channel
+corresponding to the RR Debug Session Pod that was just created.
 

### 2020-12-02 containers/frontend/react-flask-app/api/api.py

@@ -54,10 +54,7 @@ def new():
     try:
         name = request.get_json()['name']
         program = request.get_json()['program']
-        examples = tpm.get_examples()
-        container = examples[program]
-        print(container)
-        channel = tpm.create_pod([name], container)
+        channel = tpm.create_pod([name], program)
         return {'channel': channel}
     except:
         return {'error': 'Internal Error'}

### 2020-12-02 containers/frontend/react-flask-app/api/podmanager.py

@@ -10,6 +10,10 @@ class NoMatchingChannelError(Exception):
     """Raised when no pod matches a provided channel"""
     pass
 
+class NoSuchExampleError(Exception):
+    """Raised when no example image is found for a given image name"""
+    pass
+
 class TranslationPodManager:
     def __init__(self, url='localhost', port=27017):
         self.dbc = database.DatabaseController(url, port)
@@ -17,7 +21,13 @@ class TranslationPodManager:
         config.load_incluster_config()
         self.v1 = client.CoreV1Api()
         
-    def create_pod(self, names, image):
+    def create_pod(self, names, program):
+        examples = self.get_examples()
+        image = None
+        try:
+            image = examples[program]
+        except:
+            raise NoSuchExampleError
         dep = {'apiVersion': 'v1',
                'kind': 'Pod',
                'metadata': {'labels': {'purpose': 'translate-rr'}},

### 2020-12-04 paper/paper.tex
Other example programs and some more writing

@@
 
+\lstset{frame=tb,
+  aboveskip=8mm,
+  columns=flexible,
+  numbers=none,
+  tabsize=3
+}
+
 \graphicspath{{./images/}}
@@
 Similarly to research on the value of teaching debugging, research
-into how to best teach debugging is self-admittedly sparse.  Chan et
-al. allow that ``in general research on how to improve debugging is
-sporadic''---an observation that leads them to research a framework to
-reduce the complexity of teaching
-debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
-they split debugging knowledge into 5 categories: \textit{Domain},
-\textit{System}, \textit{Procedural}, \textit{Strategic}, and
-\textit{Experiential}.  They then review different debugging tools and
-teaching aids---from those that involve writing code to games---and
-map tools to the knowledge areas they seek to address.  After an
-evaluation of a host of different tools, they claim to find a few
-significant faults in current debugging teaching platforms.  The
-collaborative debugger primarily seeks to address the lack of
-back-tracing ability/coverage found in their research.
+into how to best teach debugging is sparse.  Chan et al. allow that
+``in general research on how to improve debugging is sporadic''---an
+observation that leads them to research a framework to reduce the
+complexity of teaching debugging\cite{10.1145/3286960.3286970}.  To
+organize their framework, they split debugging knowledge into 5
+categories: \textit{Domain}, \textit{System}, \textit{Procedural},
+\textit{Strategic}, and \textit{Experiential}.  They then review
+different debugging tools and teaching aids---from those that involve
+writing code to games---and map tools to the knowledge areas they seek
+to address.  After an evaluation of a host of different tools, they
+claim to find a few significant faults in current debugging teaching
+platforms.  The collaborative debugger primarily seeks to address the
+lack of back-tracing ability/coverage found in their research.
 
@@
 \cite{10.1145/1026487.1008043}, further indicating that the simple act
-of collaboration definitively benefits computer science students.\par
+of collaboration definitively benefits computer science students.
+\par
 
@@
 confidence \textit{and} correctness that they do from pair programming
-in introductory comptuer science classes.\par
-
-The collaborative debugger aims to provide such a platform.
-
-\subsection{Existing Tools that Enable Collaborative Programming} \label{exisitingcollab}
-
-TODO: write about glitch, repl.it, etc.
+in introductory comptuer science classes.
+\par
 
@@
 programming languages themselves. By becoming familiar with a specific
-debugger, students learn techniques and paradigms necessary to use all
-debuggers effectively.
+debugger, students may learn techniques and paradigms necessary to use
+all debuggers effectively.  The collaborative debugger aims to provide
+the optimal platform for students to learn debugging skills.
+
+\section{Requirments}
+
+In order to provide the best platform for both teachers and students,
+the collaborative debugger must fufill two key requirements:
+
+\begin{enumerate}
+\item It must encourage collaboration through a seamless experience.
+\item It must make it easier to enhance students' understanding of
+  debugging and programs being debugged.
+\end{enumerate}
+
+\subsection{Encouraging Collaboration}
+
+A large number of tools exist to facilitate collaborative programming.
+COVID19 has greatly increased the demand for tools that not only make
+collaboration easier, but make it easy remotely.  The tool that was
+most influential in the design of the collaborative debugger,
+Replit\cite{replit}, enables remote collaborative programming.
+
+Replit provides a simple browser-based IDE for over 50 languages.
+Programmers choose a language and can edit and run code
+collaboratively inside the Replit webapp.  Similarly to other
+collaborative text-editors, input and output from all users is synced.
+\par
+
+The collaborative debugger aims to provide a similar experience to
+Replit.  Students should be able to start a debugging session, easily
+invite other students, and interact with the debugger in a way that
+makes it easy to share their knowledge.  By allowing multiple users to
+interact with the same debugging session together in real-time, the
+collaborative debugger lets students share experience and work through
+problems together.  The debugger also makes it easy for teachers to
+demonstrate debugging techniques remotely to multiple students, any of
+whom can also interact with the debugger.  Since debugging sessions
+are hosted remotely, users can start a session in one location and
+resume it later from a different computer.
+\par
+
+\subsection{Enhancing Student's Understanding}
+
+Apart from the benefits to understanding that are realized through
+collaboration, the collaborative debugger should enhance students'
+understanding of debugging in ways that a traditional debugger cannot.
+It aims to do this by allowing teachers to design and distribute
+lessons in the form of debugging sessions.  Each debugging session is
+a lesson consisting of a deterministic recording of program execution,
+which students can debug repeatedly in order to build debugging skills
+and to learn about the execution process.
+\par
+
+The aim is to provide students with an experience they can directly
+reference when using rr or gdb in the future, as well as one which is
+seamless to encourage the acquisition of debugging skills.
+
+\subsection{Existing Tools that Enable Collaborative Programming} \label{exisitingcollab}
 
@@
 \begin{enumerate}
-\item A frontend web app built using React (\ref{react}) that presents a debugging
-  interface to the end user.
-\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
+\item A frontend web app built using React (\ref{react}) that presents
+  a debugging interface to the end user.
+\item A distributed backend managed by Kubernetes (\ref{k8s}), split
+  into three parts:
   \begin{enumerate}
@@
 to create the container image for the RR Message
-Server is shown below:\\
+Server is shown below:
 
@@
 of pods that need to interact with a load balancer, an
-\lstinline{app}.\\
-
-\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Debug Session Schema},captionpos=b]
+\lstinline{app}.
+
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Session Schema},captionpos=b]
 apiVersion: v1
@@
 Pods, schemas that define Services are applied manually using the
-\lstinline{kubectl apply} command.\\
-
-\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend Load Balancer Schema},captionpos=b]
+\lstinline{kubectl apply} command.
+
+\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={Frontend Load Balancer Schema},captionpos=b]
 apiVersion: v1
@@
 The RR Message server is remarkably simple, consisting of just 3
-important functions:\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Message Server},captionpos=b]
+important functions:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Message Server},captionpos=b]
 @sio.on('join_channel')
@@
 python, and all necessary packages as well as the app that will run on
-the final image, \lstinline{rrtranslation.py}.\\
-
-\begin{lstlisting}[language=Dockerfile,basicstyle=\small,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
+the final image, \lstinline{rrtranslation.py}.
+
+\begin{lstlisting}[language=Dockerfile,basicstyle=,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
 FROM example-container-registry.com/sproj/...
@@
 Session example.  The build process copies over any necessary files,
-as well as a startup script:\\
-
-\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Example Startup Script},captionpos=b]
+as well as a startup script:
+
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Example Startup Script},captionpos=b]
 gcc -g -o hash hash.c
@@
 \lstinline{'join_channel'} message, using the channel passed in from
-the Kubernetes API call on it's creation. \\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Translation Main},captionpos=b]
+the Kubernetes API call on it's creation.
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Translation Main},captionpos=b]
 if __name__ == '__main__':
@@
 recieve an \lstinline{'rr_command'}.  The first part of
-\lstinline{sio}'s \lstinline{'rr_command'} event handler is shown below:\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Command Event Handler},captionpos=b]
+\lstinline{sio}'s \lstinline{'rr_command'} event handler is shown below:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Command Event Handler},captionpos=b]
 @sio.on('rr_command')
@@
 control the rr subprocess, as well as a collection of methods to parse
-rr output, the three most important of which are shown below: \\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RRInterface},captionpos=b]
+rr output, the three most important of which are shown below:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RRInterface},captionpos=b]
 def get_full_rr_response(self, command):
@@
 Server to ensure students recieve a pleasant and near-identical
-experience to debugging locally in the future.  The aim is to provide
-students with an experience they can directly reference when using rr
-or gdb in the future, as well as one which is seamless to encourage
-the acquisition of debugging skills.  The system of inheritance used
-to create example debug session pods is designed to be as simple as
-possible, though a more fully featured frontend could simplify this
-process further in the future.
+experience to debugging locally in the future.  The system of
+inheritance used to create example debug session pods is designed to
+be as simple as possible, though a more fully featured frontend could
+simplify this process further in the future.
 
@@
 These scripts are defined in the server's \lstinline{package.json}
-configuration file:\\
-
-\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
+configuration file:
+
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Frontend/API Server Configuration},captionpos=b]
   "start": "react-scripts start",
@@
 the instantiation of \lstinline{TranslationPodManager} and
-\lstinline{UserManager}.\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={API Server New RR Debug Session Event Handler},captionpos=b]
+\lstinline{UserManager}.
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={API Server New RR Debug Session Event Handler},captionpos=b]
 tpm = podmanager.TranslationPodManager(
@@
 
-% \begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
+% \begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Frontend/API Server Configuration},captionpos=b]
 % users:            pods:               examples:
@@
 binds users to the Pod, and returns the channel the Pod was created
-with.  The first part of the function is shown below:\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 1},captionpos=b]
+with.  The first part of the function is shown below:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={Pod Creation 1},captionpos=b]
 def create_pod(self, names, program):
@@
 passed a name when there are no images, an exception is raised.
-Otherwise, a RR Debug Session Pod schema is created using the image:\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 2},captionpos=b]
+Otherwise, a RR Debug Session Pod schema is created using the image:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={Pod Creation 2},captionpos=b]
   dep = {'apiVersion': 'v1',
@@
 \lstinline{create_pod} does the
-work necessary to dynamically create a Pod:\\
-
-\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 3},captionpos=b]
+work necessary to dynamically create a Pod:
+
+\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={Pod Creation 3},captionpos=b]
   channel = self.dbc.add_pod(

### 2020-12-04 example_programs/memoization/memo.c

@@ -1,0 +1,85 @@
+#include <stdlib.h>
+#include <stdio.h>
+
+#define HASHSIZE 100
+
+struct pair {
+  struct pair *next;
+  int key;
+  int value;
+};
+
+static struct pair *hashtable[HASHSIZE];
+
+unsigned hash (int k)
+{
+  return k % HASHSIZE;
+}
+
+struct pair *search(int k)
+{
+  struct pair *pp;
+
+  for (pp = hashtable[hash(k)]; pp != NULL; pp = pp->next)
+    if (k == pp->key)
+      return pp;
+  return NULL;
+}
+
+/* The bug here is that they forget to use strdup */
+struct pair *insert(int key, int value)
+{
+  struct pair *pp;
+  unsigned hashval;
+
+  if ((pp = search(key)) == NULL) {
+    pp = (struct pair *) malloc(sizeof(*pp));
+    if (pp == NULL)
+      return NULL;
+    pp->key = key;
+    hashval = hash(key);
+    pp->next = hashtable[hashval];
+    hashtable[hashval] = pp;
+  } 
+  pp->value = value;
+  return pp;
+}
+
+int fib_rec (int n){
+  if (n < 0)
+    return -1;
+  if (n == 0 || n == 1)
+    return n;
+  return fib_rec(n - 1) + fib_rec (n - 2);
+}
+
+int fib_itr (int n) {
+  int a = 0;
+  int b = 1;
+  for (int x = 0; x < n; x ++){
+    int t = a;
+    a = b;
+    b = t + b;
+  }
+  return a;
+}
+
+int fib_memo (int n) {
+  if (n < 0)
+    return -1;
+  if (n == 0 || n == 1)
+    return n;
+  struct pair *pp = search(n);
+  if (pp != NULL){
+    return pp->value;
+  }
+  pp = insert(n, fib_memo(n - 1) + fib_memo (n - 2));
+  return pp->value;
+}
+
+int main(){
+  int term = 30;
+  
+  printf("Fib recursive: %d\nFib iterative: %d\nFib memoization: %d\n",
+	 fib_rec(term), fib_itr(term), fib_memo(term));
+}

### 2020-12-04 example_programs/memoization/startup.sh

@@ -1,0 +1,3 @@
+gcc -g -o memo memo.c
+rr record ./memo
+python3 rrtranslation.py $1

### 2020-12-04 example_programs/smashing/greet.c

@@ -1,0 +1,18 @@
+#include <stdio.h>
+
+int main (int argc, char **argv) {
+  char x [16];
+
+  if (argc < 2){
+    printf("Usage: ./greet <name>\n");
+    return 1;
+  }
+
+  // Input given is "will_this_crash_the_code?"
+  sscanf(argv[1], "%s", x);
+
+  // Break at line 15 and examine the stack using the command:
+  // x/50cb $sp
+  printf("Hello %s!", x);
+  return 0;
+}

### 2020-12-04 example_programs/smashing/startup.sh

@@ -1,0 +1,3 @@
+gcc -g -o greet greet.c
+rr record ./greet will_this_crash_the_code?
+python3 rrtranslation.py $1

### 2020-12-04 paper/paper.tex
Lots and lots of paper reorganization, section on the frontend started, intro modified

@@
 
+Programmers spend large amounts of time debugging.  They often use
+debuggers to assist them in this process of diagnosing sometimes
+inexplicable problems with their code.  Though debuggers are primarily
+used to hunt down hiden mistakes, they are also powerful tools to
+understand program execution.
+\par
+
 The goal of this project is to create a collaborative debugger to aid
-in the teaching of debugging.  The collaborative debugger aims to help
-teachers integrate this undertaught skill into their classes.  It
+in the teaching of debugging.  The collaborative debugger hopes to
+help teachers integrate this undertaught skill into their classes.  It
 realizes the benefits of collaborative programming and teaching
-debugging by providing a platform that is genuinely useful to students
-and teachers.
+debugging by providing a platform that aims to be useful to both
+students and teachers.
 
@@
 
-\subsection{Encouraging Collaboration}
+\subsection{Encouraging Collaboration} \label{exisitingcollab}
 
@@
 The collaborative debugger aims to provide a similar experience to
-Replit.  Students should be able to start a debugging session, easily
-invite other students, and interact with the debugger in a way that
-makes it easy to share their knowledge.  By allowing multiple users to
-interact with the same debugging session together in real-time, the
-collaborative debugger lets students share experience and work through
-problems together.  The debugger also makes it easy for teachers to
-demonstrate debugging techniques remotely to multiple students, any of
-whom can also interact with the debugger.  Since debugging sessions
-are hosted remotely, users can start a session in one location and
-resume it later from a different computer.
+tools like Replit.  Students should be able to start a debugging
+session, easily invite other students, and interact with the debugger
+in a way that makes it easy to share their knowledge.  By allowing
+multiple users to interact with the same debugging session together in
+real-time, the collaborative debugger lets students share experience
+and work through problems together.  The debugger also makes it easy
+for teachers to demonstrate debugging techniques remotely to multiple
+students, any of whom can also interact with the debugger.  Since
+debugging sessions are hosted remotely, users can start a session in
+one location and resume it later from a different computer.
 \par
@@
 
-The aim is to provide students with an experience they can directly
-reference when using rr or gdb in the future, as well as one which is
-seamless to encourage the acquisition of debugging skills.
-
-\subsection{Existing Tools that Enable Collaborative Programming} \label{exisitingcollab}
-
-\section{Tools Used}
+Currently, there are three example debug sessions included with the
+collaborative debugger:
+
+\begin{enumerate}
+\item \lstinline{hash}: an example of classic pointer confusion.  The
+  hashtable implementation included with the program is slightly
+  bugged, where failing to call \lstinline{strncpy} results in the
+  value for each key-value pair in the hash pointing to the same
+  location in memory.  This is particuarly useful example for learning
+  about setting watchpoints and breakpoints, as well as basic
+  procedures to step through code.
+\item \lstinline{smash}: a brief example of stack smashing.  A badly
+  designed call to \lstinline{sscanf} results in a portion of the call
+  stack being overwritten, and the process is killed as a result of
+  stack smashing.  This example focuses on printing and examining
+  memory locations.
+\item \lstinline{memoization}: a program that demonstrates the
+  difference between a recursive, iterative, and memoized recursive
+  implimentation of the classic nth fibonacci number function.  This
+  example focuses on examining the stack under conditions produced by
+  different types of functions.
+\end{enumerate}
+
+\lstinline{memoization} will be covered in detail in the following
+section on the frontend of the collaborative debugger (\ref{webapp}).
+All three examples show how teachers can design debugging sessions
+around specific concepts.  The collaborative debugger should make it
+easy to distribute these sessions and for students to learn from them
+collaboratively.
+\par
+
+Students using all three examples would benefit greatly from
+visualization tools to help understand pointer locations, the process
+space, and data structures.  The steps necessary to implement
+visualization tools in the collaborative debugger are discussed
+further in (\ref{next}).
+\par
+
+The aim of the collaborative debugger is to provide students with an
+experience they can directly reference when using similar debuggers in
+the future, as well as one which is seamless to encourage the
+acquisition of debugging skills.  Creating such a platform will
+hopefully encourage both students and teachers to explore the
+undertaught area of debugging.
+
+\section{Frontend Webapp}\label{webapp}
+
+Before diving into the construction of the frontend webapp, it's
+helpful to look at the steps two students would take to debug the
+\lstinline{memoization} example program remotely.
 
@@
 
-  \includegraphics[scale=.8]{overall_system}
+  \includegraphics[width=\textwidth]{memoization}
   \centering
-  \caption{Overview of the Collaborative Debugger}
+  \caption{Debugging the Memoization Example Session}
   \label{debugger:overview}
@@
 
-The next sections give an overview of the various tools used to create
-the collaborative debugger.  The debugger consists of:
-
 \begin{enumerate}
-\item A frontend web app built using React (\ref{react}) that presents
-  a debugging interface to the end user.
-\item A distributed backend managed by Kubernetes (\ref{k8s}), split
-  into three parts:
-  \begin{enumerate}
-  \item A Pod for each debugging instance which runs the rr debugger
-    (\ref{rr}).  These communicate directly with users through
-    WebSocket server Pods. (\ref{socketio}).
-  \item Pods running a frontend server written in Node.js which works
-    in tandem with an API server created using Flask
-    (\ref{flask/node}).  The API server manages creation and
-    destruction of debug sessions, as well as authentication.
-  \item A MongoDB (\ref{mongodb}) database.
+\item Students A and B both log into the collaborative debugger from
+  different computers.
+\item Student A creates a new Debug Session by clicking the button for
+  \lstinline{memoization} under ``New Session''.  A few moments later,
+  they're presented with the main screen of the debugger.
+\item Using the session code (17734 in the example above), Student B
+  joins the Debug Session.
+\item Both students can now interact with the same debugger remotely.
+  Input and output from the debugger and the current source file are
+  synced between all students.  Students can work together to
+  understand the reasons why memoization often improves the execution
+  time of recursive programs.
 \end{enumerate}
-\end{enumerate}
-
-\subsection{Kubernetes}\label{k8s}
+
+Notice that the debugging window in the example shown above is called
+``rr''.  rr is an enhancment to gdb that is foundational in the
+collaborative debugger's ability to deterministically record and
+replay program execution. rr and its role in the collaborative
+debugger will be covered extensively in (\ref{rrpods}).  Users
+interact with rr through the collaborative debugger's frontend.
+
+\subsection{Tools Used}
+
+The frontend of the collaborative debugger must handle constant
+changes in state as users participate in the process outlined above.
+The frontend uses the following tools to manage state and display
+output.
+
+\subsubsection{React}\label{react}
+
+React is JavaScript library that simplifies creating user interfaces
+and managing state \cite{react}.  React's state management is of
+particular importance to the collaborative debugger's frontend.  State
+constantly changes as users create/delete debugging sessions, join
+existing sessions, and communicate with rr.  React allows classes and
+functions to encapsulate components such as a list of existing
+debugging sessions, a view of the current program's source code, and
+the terminal interface with rr.  Instances of these classes maintain
+state and update efficiently.
+\par
+
+The frontend makes extensive use of JSX, syntax which allows the
+inclusion of segments of HTML code within a React app written in
+JavaScript.  This makes it easy for each component of the one-page
+webapp to hide/show subcomponents as state changes.
+
+\subsubsection{Monaco and Xterm.js}\label{xtermjs/monaco}
+
+After joining or creating a debugging session, users spend most of
+their time interacting collaboratively with rr.  Their primary
+interface to rr is through Xterm.js, a frontend component that makes
+it easy to emulate terminal behavior in the browser \cite{xtermjs}.
+With a few control methods, it is simple to provide a terminal
+interface to rr that is virtually indistinguishable from a local
+session.  By using the Xterm.js based interface, students can learn to
+use rr (and by extension gdb) collaboratively, and directly translate
+that knowledge to individual work.
+\par
+
+In addition to the terminal interface, the frontend shows a view of
+the current source file being debugged.  The Monaco Editor
+\cite{monaco} is used to display this source view.  Though more
+complex than is strictly necessary to display code, Monaco makes it
+easy to format and syntax-highlight.  Using Monaco also simplifies the
+future addition of editing source code, should the need arise.
+React's state management allows updating text in the editor as
+efficiently as possible.
+
+\section{Backend Design Overview}
+
+The design of the collaborative debugger's backend is heavily
+distributed, allowing individual components to be modified without the
+whole system needing to be reconfigured.  Kubernetes in conjunction
+with Docker are used to create the backend of the collaborative
+debugger.
+
+\subsection{Tools Used}
+
+Kubernetes and Docker were chosen largely so that new Debug Sessions
+could be easily and securely created on demand.
+
+\subsubsection{Docker}\label{docker}
+
+Docker is containerization software.  Docker containers encapsulate
+applications and their dependencies while sharing their host's kernel.
+This allows for lightweight, portable, and secure execution of complex
+programs \cite{docker}.  All programs in the collaborative debugger's
+backend run in Docker containers.  Docker containers' lightweight
+nature and security are instrumental when trying to quickly create
+containers to run user-defined code.
+
+\subsubsection{Kubernetes}\label{k8s}
 
@@
 
-\subsection{Mozilla's rr}\label{rr}
-
-\subsubsection{Overview}
-
-rr is ``a lightweight tool for recording, replaying and debugging
-execution of applications''\cite{rr-repo}. rr allows a programmer to
-record the execution of a program on any compatible machine and replay
-the execution later.  This enhances GDB's ability to ``time-travel''
-when debugging, using commands such as \lstinline{reverse-continue}
-and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
-forwards through a program's execution.  Through a novel encapsulation
-of the execution space, rr is able to deterministically record and
-replay the execution of syscalls and other process behavior that
-differs run-to-run.  This is invaluable when trying to debug behavior
-that is not entirely dependent on the code being debugged.  A typical
-workflow in rr consists of recording an inexplicable error, replaying
-execution to find the area in which the error occurs, and then
-narrowing in on the bug not by re-running the entire program, but by
-progressing back and forth through execution in the problem area.
-\par
-
-rr is an ideal tool for teaching debugging because it allows
-instructors to record execution of a program and design a debugging
-example with the knowledge that normally non-deterministic events will
-be repeatable, and that any input they provide to the program will be
-exactly replicated.  With the collaborative debugger, teachers can
-record a program's execution and design a debugging lesson which
-students can work on together.  The repeatability of rr means that
-students can focus on debugging, and teachers can create as specific
-examples as they please.  The use of rr is the most significant step
-the collaborative debugger takes to addressing the lack of
-back-tracing ability/coverage found in existing tools for teaching
-debugging\cite{10.1145/3286960.3286970}.
-
-\subsubsection{Limitations}
-
-In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
-that rely on capturing the entire state of of a virtual machine to
-replay execution, rr records and replays faster, produces far smaller
-files, and doesn't force execution inside of a
-VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The trade off for these
-benefits are two major system limitations: rr is only compatible with
-the Linux kernel, and it's deterministic recording and replay relies
-on a feature that is only found on modern \textit{Intel} x86 CPUs.
-These limitations influenced the development of this project as a
-webapp similar to existing tools for collaborative programming.
-\par
-
-Luckily, the speed and size benefits of rr lend themselves well to
-non-local execution.  In conjunction with Kubernetes, it takes a few
-seconds to create a new container running rr and connect to web
-clients.
-
-\subsection{pygdbmi}
-
-In order to ``support the development of systems which use the
-debugger as just one small component of a larger system'', GDB
-provides a machine-oriented interface called GDB/MI \cite{gdbman}. rr
-supports interaction through GDB/MI, and using the interface was a
-natural choice for the collaborative debugger.  In addition to being
-far easier to interact with from within a program, the structured,
-machine-friendly output of GDB/MI lends itself in particular to future
-development of visualization aids in the collaborative debugger.
-\par
-To parse rr output into Python dictionaries and to easily control rr
-as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
-pygdbmi's abstraction simplifies programatically controlling rr.  A
-Pod can receive a command from the client, pass it to rr, and respond
-without having to deal with parsing GDB/MI output or with directly
-managing the rr process.
-
-\subsection{Socket.IO}\label{socketio}
-
-To speed communication, the collaborative debugger uses WebSockets to
-directly connect web clients and the Pods running rr.  Socket.IO is a
-library that extends WebSockets.  It provides backup incase a
-WebSocket connection cannot be established, enables automatic
-reconnection and disconnection detection, and adds support for
-namespaces \cite{socketio}.  The collaborative debugger uses the
-standard JavaScript implementation of Socket.IO on the client side.
-Messages are passed through a server to individual debugging Pods,
-both of which use the Python implementation of Socket.IO,
-python-socket.io \cite{python_socketio}.
-
-\subsection{MongoDB}\label{mongodb}
-
-The collaborative debugger uses a database to store information about
-users, Pods, and example debugging sessions.  Due to it's speed of
-deployment and natural interaction with the object-oriented languages
-used to create the project, MongoDB was chosen as database software
-\cite{mongodb}.
-
-\subsection{Flask and Node.js}\label{flask/node}
-
-The primary server for the collaborative debugger is split into two
-sections: a simple Node.js \cite{node} server that serves the frontend
-webapp, and an API server created using Flask \cite{flask}.  While in
-development, the builtin React (\ref{react}) development server
-is used to serve the frontend.  This makes debugging the frontend far
-easier.
-\par
-
-An API server is necessary to authenticate users and to provide a
-means to create/delete debugging sessions.  Since the rest of the
-backend was created using Python, Flask was chosen to create the API
-server.  Flask is a lightweight web application framework which lends
-itself perfectly to interacting with the Python MongoDB and Kubernetes
-APIs.
-
-\subsection{React}\label{react}
-
-React is JavaScript library that simplifies creating user interfaces
-and managing state \cite{react}.  React's state management is of
-particular importance to the collaborative debugger's frontend.  State
-constantly changes as users create/delete debugging sessions, join
-existing sessions, and communicate with rr.  React allows classes to
-encapsulate components such as a list of existing debugging sessions,
-a view of the current program's source code, and the terminal
-interface with rr.  Instances of these classes maintain state
-and update efficiently.
-\par
-The frontend makes extensive use of JSX, syntax which allows the
-inclusion of segments of HTML code within a React app written in
-JavaScript.  This makes it easy for each component of the one-page
-webapp to hide/show subcomponents as state changes.
-
-\subsection{Monaco and Xterm.js}\label{xtermjs/monaco}
-
-After joining or creating a debugging session, users spend most of
-their time interacting collaboratively with rr.  Their primary
-interface to rr is through Xterm.js, a frontend component that makes
-it easy to emulate terminal behavior in the browser \cite{xtermjs}.
-With a few control methods, it is simple to provide a terminal
-interface to rr that is virtually indistinguishable from a local
-session.  By using the Xterm.js based interface, students can learn to
-use rr (and by extension gdb) collaboratively, and directly translate
-that knowledge to individual work.
-\par
-
-In addition to the terminal interface, the frontend shows a view of
-the current source file being debugged.  The Monaco Editor
-\cite{monaco} is used to display this source view.  Though more
-complex than is strictly necessary to display code, Monaco makes it
-easy to format and syntax-highlight.  Using Monaco also simplifies the
-future addition of editing source code, should the need arise.
-React's state management allows updating text in the editor as
-efficiently as possible.
-
-\section{Design}
+\subsection{Overview}
 
@@
 backend primarally so that a Pod could be created dynamically for each
-debugging session.  The design of the backend is heavily distributed,
-allowing individual components to be modified without the whole system
-needing to be reconfigured.
+debugging session.
 
@@
   Message Server Pods and Frontend/API Server Pods may be created
-  using Deployments \cite{k8s_docs} to allow later scaling, where
-  multiple Pods running the same application may be created to
-  facilitate increased load.
+  using Deployments \cite{k8s_docs} to allow scaling, where multiple
+  Pods running the same application may be created to facilitate
+  increased load.
 \item Dynamically created Pods.  This class of Pod contains the
@@
 
-Each component of the backend and frontend will be discussed in depth
-in the following sections.
+Each component of the backend will be discussed in depth in the
+following sections.
 
@@
 
-\subsection{RR Debug Sessions \& RR Message Server}
+\section{RR Debug Sessions \& RR Message Server}
 
@@
 
-\subsubsection{RR Message Server}
+\subsection{RR Message Server}
+
+\subsubsection{Tools Used}
+
+The collaborative debugger needs an effecient way for clients to
+communicate with running RR Debug Sessions. The Socket.IO library was
+selected to construct this efficient communication pathway.
+
+\paragraph{Socket.IO}\label{socketio}
+
+To speed communication, the collaborative debugger uses WebSockets to
+directly connect web clients and the Pods running rr.  Socket.IO is a
+library that extends WebSockets.  It provides backup incase a
+WebSocket connection cannot be established, enables automatic
+reconnection and disconnection detection, and adds support for
+namespaces \cite{socketio}.  The collaborative debugger uses the
+standard JavaScript implementation of Socket.IO on the client side.
+Messages are passed through a server to individual debugging Pods,
+both of which use the Python implementation of Socket.IO,
+python-socket.io \cite{python_socketio}.
+
+\subsubsection{Overview}
 
@@
 
-\subsubsection{RR Debug Session}
-
-\paragraph{Building Example Pods} \label{buildingchannel}
+\subsection{RR Debug Session} \label{rrpods}
+
+\subsubsection{Tools Used}
+
+\paragraph{Mozilla's rr}\label{rr}
+
+rr is ``a lightweight tool for recording, replaying and debugging
+execution of applications''\cite{rr-repo}. rr allows a programmer to
+record the execution of a program on any compatible machine and replay
+the execution later.  This enhances GDB's ability to ``time-travel''
+when debugging, using commands such as \lstinline{reverse-continue}
+and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
+forwards through a program's execution.  Through a novel encapsulation
+of the execution space, rr is able to deterministically record and
+replay the execution of syscalls and other process behavior that
+differs run-to-run.  This is invaluable when trying to debug behavior
+that is not entirely dependent on the code being debugged.  A typical
+workflow in rr consists of recording an inexplicable error, replaying
+execution to find the area in which the error occurs, and then
+narrowing in on the bug not by re-running the entire program, but by
+progressing back and forth through execution in the problem area.
+\par
+
+rr is an ideal tool for teaching debugging because it allows
+instructors to record execution of a program and design a debugging
+example with the knowledge that normally non-deterministic events will
+be repeatable, and that any input they provide to the program will be
+exactly replicated.  With the collaborative debugger, teachers can
+record a program's execution and design a debugging lesson which
+students can work on together.  The repeatability of rr means that
+students can focus on debugging, and teachers can create as specific
+examples as they please.  The use of rr is the most significant step
+the collaborative debugger takes to addressing the lack of
+back-tracing ability/coverage found in existing tools for teaching
+debugging\cite{10.1145/3286960.3286970}.
+
+In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
+that rely on capturing the entire state of of a virtual machine to
+replay execution, rr records and replays faster, produces far smaller
+files, and doesn't force execution inside of a
+VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The trade off for these
+benefits are two major system limitations: rr is only compatible with
+the Linux kernel, and it's deterministic recording and replay relies
+on a feature that is only found on modern \textit{Intel} x86 CPUs.
+These limitations influenced the development of this project as a
+webapp similar to existing tools for collaborative programming.
+\par
+
+Luckily, the speed and size benefits of rr lend themselves well to
+non-local execution.  In conjunction with Kubernetes, it takes a few
+seconds to create a new container running rr and connect to web
+clients.
+
+\paragraph{pygdbmi}
+
+In order to ``support the development of systems which use the
+debugger as just one small component of a larger system'', gdb
+provides a machine-oriented interface called gdb/mi \cite{gdbman}. rr
+supports interaction through gdb/mi, and using the interface was a
+natural choice for the collaborative debugger.  In addition to being
+far easier to interact with from within a program, the structured,
+machine-friendly output of gdb/mi lends itself in particular to future
+development of visualization aids in the collaborative debugger.
+\par
+To parse rr output into Python dictionaries and to easily control rr
+as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
+pygdbmi's abstraction simplifies programatically controlling rr.  A
+Pod can receive a command from the client, pass it to rr, and respond
+without having to deal with parsing gdb/mi output or with directly
+managing the rr process.
+
+\subsubsection{Building Example Pods} \label{buildingchannel}
 
@@
 
-\begin{lstlisting}[language=Dockerfile,basicstyle=,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
+\begin{lstlisting}[language=Dockerfile,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
 FROM example-container-registry.com/sproj/...
@@
 
-\paragraph{The RR Translation Program} \label{joiningchannel}
+\subsubsection{The RR Translation Program} \label{joiningchannel}
 
@@
 \lstinline{sio}.  \lstinline{RRInterface} contains an instance of the
-\lstinline{GDBController} class from pygdbmi to interface with rr and
+\lstinline{gdbController} class from pygdbmi to interface with rr and
 control the rr subprocess, as well as a collection of methods to parse
@@
 
-\subsection{Frontend/API Server}\label{api}
+% \section{Tools Used}
+
+% \begin{figure}[h!]
+
+%   \includegraphics[scale=.8]{overall_system}
+%   \centering
+%   \caption{Overview of the Collaborative Debugger}
+%   \label{debugger:overview}
+% \end{figure}
+
+% The next sections give an overview of the various tools used to create
+% the collaborative debugger.  The debugger consists of:
+
+% \begin{enumerate}
+% \item A frontend web app built using React (\ref{react}) that presents
+%   a debugging interface to the end user.
+% \item A distributed backend managed by Kubernetes (\ref{k8s}), split
+%   into three parts:
+%   \begin{enumerate}
+%   \item A Pod for each debugging instance which runs the rr debugger
+%     (\ref{rr}).  These communicate directly with users through
+%     WebSocket server Pods. (\ref{socketio}).
+%   \item Pods running a frontend server written in Node.js which works
+%     in tandem with an API server created using Flask
+%     (\ref{flask/node}).  The API server manages creation and
+%     destruction of debug sessions, as well as authentication.
+%   \item A MongoDB (\ref{mongodb}) database.
+%   \end{enumerate}
+% \end{enumerate}
+
+\section{Frontend/API Server}\label{api}
 
@@
 webapp, as well as processing all API requests, such as those to
-authenticate users and create new RR Debug Sessions.  The builtin
-React development server is used to serve the frontend webapp and
-redirect API requests to the Flask API server during development.  For
-both security and performance reasons this should be changed to a
-combination of a more robust solution like Nginx \cite{nginx} and a
-production suitable web server for production.  The relationships
-between components and overall process will remain unchanged after
-this migration.
+authenticate users and create new RR Debug Sessions.
+
+\subsection{Tools Used}
+
+\subsubsection{MongoDB}\label{mongodb}
+
+The collaborative debugger uses a database to store information about
+users, Pods, and example debugging sessions.  Due to it's speed of
+deployment and natural interaction with the object-oriented languages
+used to create the project, MongoDB was chosen as database software
+\cite{mongodb}.
+
+\subsubsection{Flask and Node.js}\label{flask/node}
+
+The primary server for the collaborative debugger is split into two
+sections: a simple Node.js \cite{node} server that serves the frontend
+webapp, and an API server created using Flask \cite{flask}.  While in
+development, the builtin React (\ref{react}) development server
+is used to serve the frontend.  This makes debugging the frontend far
+easier.
+\par
+
+An API server is necessary to authenticate users and to provide a
+means to create/delete debugging sessions.  Since the rest of the
+backend was created using Python, Flask was chosen to create the API
+server.  Flask is a lightweight web application framework which lends
+itself perfectly to interacting with the Python MongoDB and Kubernetes
+APIs.
+
+\subsection{Overview}
+
+The builtin React development server is used to serve the frontend
+webapp and redirect API requests to the Flask API server during
+development.  For both security and performance reasons this should be
+changed to a combination of a more robust solution like Nginx
+\cite{nginx} and a production suitable web server for production.  The
+relationships between components and overall process will remain
+unchanged after this migration.
 \par
@@
 
-\subsubsection{Frontend/API Server Configuration}
+\subsection{Frontend/API Server Configuration}
 
@@
 
-\subsubsection{The API Server}
+\subsection{The API Server}
 
@@
 
-\paragraph{Processing a POST Request---Extracting Information}
+\subsubsection{Processing a POST Request---Extracting Information}
 
@@
 
-\paragraph{The Database} Before drilling into \lstinline{create_pod},
+\subsubsection{The Database} Before drilling into \lstinline{create_pod},
 it may be helpful to examine the MongoDB database which
@@
 
-\paragraph{Processing a POST Request---Dynamic Pod Creation}
+\subsubsection{Processing a POST Request---Dynamic Pod Creation}
 
@@
 
-\paragraph{Processing a POST Request---Returning a Response}
+\subsubsection{Processing a POST Request---Returning a Response}
 
@@
 
-\subsection{Webapp}\label{webapp}
-
-TODO
-
-\section{Next Steps}
+\section{Next Steps} \label{next}
 

### 2020-12-04 example_programs/memoization/memo.c

@@ -1,5 +1,6 @@
 #include <stdlib.h>
 #include <stdio.h>
+#include <time.h>
 
 #define HASHSIZE 100
 
@@ -26,7 +27,6 @@ struct pair *search(int k)
   return NULL;
 }
 
-/* The bug here is that they forget to use strdup */
 struct pair *insert(int key, int value)
 {
   struct pair *pp;
@@ -77,9 +77,20 @@ int fib_memo (int n) {
   return pp->value;
 }
 
+double timer(int (f)(int), int i, int *result){
+  clock_t start = clock();
+  *result = f(i);
+  clock_t finish = clock();
+  return (double)(finish - start) / CLOCKS_PER_SEC;
+}
+
 int main(){
   int term = 30;
-  
-  printf("Fib recursive: %d\nFib iterative: %d\nFib memoization: %d\n",
-	 fib_rec(term), fib_itr(term), fib_memo(term));
+  int fr, fi, fm;
+  double tfr, tfi, tfm;
+  tfr = timer(&fib_rec, term, &fr);
+  tfi = timer(&fib_itr, term, &fi);
+  tfm = timer(&fib_memo, term, &fm);
+  printf("Fib recursive: time=%f result=%d\nFib iterative: time=%f result=%d\nFib memoization: time=%f result=%d\n",
+	 tfr, fr, tfi, fi, tfm, fm);
 }

### 2020-12-05 paper/paper.tex
Final Draft Finished

@@
 
+% From https://tex.stackexchange.com/questions/89574/language-option-supported-in-listings
+
+\definecolor{lightgray}{rgb}{.9,.9,.9}
+\definecolor{darkgray}{rgb}{.4,.4,.4}
+\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
+
+\lstdefinelanguage{JavaScript}{
+  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break},
+  keywordstyle=\color{blue}\bfseries,
+  ndkeywords={class, export, boolean, throw, implements, import, this},
+  ndkeywordstyle=\color{darkgray}\bfseries,
+  identifierstyle=\color{black},
+  sensitive=false,
+  comment=[l]{//},
+  morecomment=[s]{/*}{*/},
+  commentstyle=\color{purple}\ttfamily,
+  stringstyle=\color{red}\ttfamily,
+  morestring=[b]',
+  morestring=[b]"
+}
+
 \graphicspath{{./images/}}
 
-\author{Sammy Furr}
+\author{A Senior Project submitted to\\The Division of Science, Mathematics, and Computing\\of\\Bard College\\\\by\\Sammy Furr}
 \title{The Development of a Collaborative Tool to Teach Debugging}
@@
 \begin{titlepage}
-  \maketitle
+  \linespread{1}
+    \begin{center}
+
+        \Huge
+        The Development of a Collaborative Tool to Teach Debugging
+
+        \vspace{4cm}
+
+        \large
+        A Senior Project submitted to\\
+        The Division of Science, Mathematics, and Computing\\
+        of\\
+        Bard College\\
+        \vspace{1cm}
+        by\\
+        Sammy Furr
+        \vfill
+
+        Annandale-on-Hudson, New York\\
+        December, 2020
+    \end{center}
 \end{titlepage}
@@
 \begin{abstract}
-  TODO: write an abstract
+  \linespread{1}
+  Debugging is rarely formally taught, despite being used by
+  programmers every day.  Research indicates that it is valuable to
+  teach debugging, and suggests that teaching it collaboratively may
+  be maximally effective.  The goal of this project is to create a
+  collaborative debugger.  The debugger aims to be the ideal platform
+  to teach and learn debugging.  This paper briefly reviews relevant
+  literature which covers teaching debugging and that which focuses on
+  teaching programming collaboratively.  Most of the paper is devoted
+  to the design of the collaborative debugger.
 \end{abstract}
+
+\section{Acknowledgments}
+
+\begin{itemize}
+\item Keith O'Hara, my senior project advisor, for his constant
+  thoughtfulness about my project, his advice about the future, and
+  for giving me the oppurtunity to tutor for his classes.
+\item Mar Hicks and Kyle Hale, professors whose classes at IIT changed
+  my perspective on computer science.
+\item All my professors at Bard, particuarly Olga Voronina, Wakako
+  Suzuki, and Nate Shockey, whose encouragement of my interests
+  outside of computer science has been instrumental in continuing my
+  passion for it.
+\item My family and friends, for their constant love and support.
+\end{itemize}
 
@@
 visualization tools to help understand pointer locations, the process
-space, and data structures.  The steps necessary to implement
-visualization tools in the collaborative debugger are discussed
-further in (\ref{next}).
+space, and data structures.  As discussed in \ref{vistools}, adding
+vizualization tools to the collaborative debugger should be relatively
+easy.
 \par
@@
 
-Before diving into the construction of the frontend webapp, it's
+Before diving into the construction of the frontend webapp, it may be
 helpful to look at the steps two students would take to debug the
-\lstinline{memoization} example program remotely.
+\lstinline{memoization} example program remotely:
 
@@
   \caption{Debugging the Memoization Example Session}
-  \label{debugger:overview}
+  \label{frontend:rrterm}
 \end{figure}
@@
 debugger will be covered extensively in (\ref{rrpods}).  Users
-interact with rr through the collaborative debugger's frontend.
+interact with rr through the collaborative debugger's frontend.  Care
+has been taken to make the frontend as representative of rr as
+possible so that users can learn to use rr/gdb by using the
+collaborative debugger.
 
@@
 The frontend uses the following tools to manage state and display
-output.
+output:
 
@@
 efficiently as possible.
+
+In order to easily integrate Monaco and Xterm.js into the React
+frontend, React wrapper libraries were used.  These are
+react-xterm\cite{reactxterm} and
+react-monaco-editor\cite{reactmonaco}.
+
+\subsection{Overview}
+
+React makes it easy to create simple webapps with no build
+configuration necessary through the use of the
+\lstinline{create-react-app} command.  This creates a structure and
+toolchain for a basic one-page application---suitable for the
+collaborative debugger.  Once this structure was created and any
+superfluous files/code were removed, it's wasn't necessary for the
+purposes of the collaborative debugger to directly edit any files
+except two: \lstinline{App.js} and \lstinline{App.css}.  The
+pre-defined build process takes care of transforming the React code
+defined in \lstinline{App.js} and the styles defined in
+\lstinline{App.css} into portable Javascript, HTML, and CSS.
+\par
+
+\lstinline{App.js} consists of a primary React Component class,
+\lstinline{Debugger}, as well as many secondary classes and functions.
+More complex components of the frontend, such as the Xterm.js
+interface to rr, are created as classes, while simple components like
+buttons are defined as functions.  All classes contain a
+\lstinline{render()} method responsible for returning HTML elements to
+display based on the state contained in the class.
+\lstinline{Debugger}'s \lstinline{render()} method is responsible for
+displaying all other components, based on whether the user is logged
+in, has selected a new session, etc.
+\par
+
+Many components ``lift state up'' to \lstinline{Debugger}.  This is
+accomplished by passing methods from \lstinline{Debugger} in as
+``props'' when elements are created in \lstinline{Debugger}'s
+\lstinline{render()} method:
+
+\begin{lstlisting}[language=Javascript,basicstyle=\linespread{0.5}\ttfamily,caption={Lifting State Up},captionpos=b]
+  onLogin(name) {
+    this.setState({user: name})
+    .
+    .
+    .
+  }
+  // In Debugger.render().  This is JSX.
+  login = <LoginForm onLogin={this.onLogin}/>;
+
+  // In LoginForm.
+  this.props.onLogin(data.name);
+\end{lstlisting}
+
+\lstinline{Debugger}'s \lstinline{onLogin(name)} method changes the
+state of debugger to indicate that the user is logged in.  When an
+instance of \lstinline{LoginForm} is created in
+\lstinline{Debugger.render()}, \lstinline{onLogin(name)} is passed to
+it as a ``prop''.  The code for logging in the user can be contained
+in \lstinline{LoginForm}, but when it becomes necessary to change the
+state of the whole application to indicate that the user is logged in,
+\lstinline{LoginForm} calls \lstinline{Debugger.onLogin(name)},
+lifting state up to \lstinline{Debugger}.
+\par
+
+Other than methods to handle state and display components,
+\lstinline{Debugger} also contains the Socket.IO client that is used
+to communicate with the actual debugger running in the collaborative
+debugger's backend.  Socket.IO is a library that extends WebSockets.
+It and this communication process will be discussed in detail in
+(\ref{rrmessageserver}).  Similarly to how \lstinline{onLogin(name)}
+is passed as a property to \lstinline{LoginForm}, the Socket.IO client
+is passed to the most important component in the frontend,
+\lstinline{RRTerm}, when it is instantiated.
+
+\subsection{The Collaborative Interface to rr}
+
+As shown in Figure \ref{frontend:rrterm}, the primary piece of the
+collaborative debugger's frontend is an interface to the rr debugger
+backend.  This interface currently consists of a view of the user's
+source code and a terminal interface to rr.  These are contained
+within the most important class in the frontend: \lstinline{RRTerm}.
+This section will outline the process of sending an rr command to the
+backend and displaying a response.
+\par
+
+In debugging the \lstinline{memoization} example program, users will
+want to display the call stack from within each implementation of the
+nth fibonacci number function.  To stop in \lstinline{fib_rec}, the
+recursive implementation, one user will want to set a breakpoint at
+line 52 in the source file.  To do this they will type the command
+\lstinline{break 52} and hit enter.  The command will be sent to the
+backend, and all users in the same debug session (17734 in Figure
+\ref{frontend:rrterm}) will see the debuggers response:\\
+\lstinline{Breakpoint 1 at 0x561c8fede303: file memo.c, line 52.}\\
+Users who did not send the command will see it displayed along with
+the response.
+
+\subsubsection{Sending a Command}
+
+\lstinline{RRTerm}'s state is instantiated as such:
+
+\begin{lstlisting}[language=Javascript,basicstyle=\linespread{0.5}\ttfamily,caption={RRTerm's State},captionpos=b]
+  this.state = {command: '',
+                  name: '',
+                  linum: 0,
+                  code: ''};
+\end{lstlisting}
+
+\lstinline{name}, \lstinline{linum}, and \lstinline{code} store
+information about the current source file being debugged.
+\lstinline{command} is the most important field for examing how
+commands are sent to the backend, as it holds the current command.
+\textit{Every time} the user types a character, \lstinline{command} is
+updated to reflect the current state of the Xterm.js component.  The
+code to handle keystrokes is shown below:
+
+\begin{lstlisting}[language=Javascript,basicstyle=\linespread{0.5}\ttfamily,caption={Handling Keystrokes},captionpos=b]
+  this.term.on('key', (key, ev) => {
+    const printable = !ev.altKey && !ev.ctrlKey && !ev.metaKey;
+    if (ev.key === 'Enter') {
+      // Disable input while we wait for the response from rr
+      this.term.disableStdin = true;
+      const c = this.state.command;
+      this.sendCommand(c);
+      this.setState({command: ''});
+    } else if (ev.key === 'Backspace') {
+      // Do not delete the prompt
+      if (this.term.buffers._activeBuffer.x > prompt_length) {
+        this.term.write('\b \b');
+        this.updateCommand(this.state.command, 'delete');
+      }
+    } else if (ev.key === 'ArrowUp') {
+    } else if (ev.key === 'ArrowDown') {
+    } else if (printable) {
+      this.term.write(key);
+      this.updateCommand(this.state.command, key);
+    }
+  });
+\end{lstlisting}
+
+Whenever the user types a printable key, the key is written to the
+terminal and \lstinline{command} is updated using the
+\lstinline{updateCommand} method.  If a user types \lstinline{brk},
+\lstinline{updateCommand} would be called three times, once for each
+keystroke.  Upon realizing they made a mistake when typing
+\lstinline{break 52}, the user would likely delete a character.  As
+long as the prompt (\lstinline{(rr)} followed by a space) would not be
+overwritten, the terminal is updated and \lstinline{updateCommand} is
+called again.  When the user hits the enter key the command is finally
+sent to the backend using \lstinline{sendCommand}.  To mimick the
+behavior of a local rr debug session and to prevent garbled input,
+writing to the terminal is disabled until a response is recieved.
+
+\subsubsection{Receiving a Response}
+
+All users connected to the same debug session will recieive responses
+to the commands sent by one user.  This is how collaboration is
+achieved in the collaborative debugger.  The JSON response object to
+\lstinline{break 52} recieved by all users is shown below:
+
+\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={JSON Response},captionpos=b]
+{
+  channel: ``11734''
+  command: ``break 52''
+  from: ``1ed644ad87d44cd68121de3daa65adb4''
+  response: {
+    output: ``Breakpoint 2 at 0x561c8fede303: file memo.c, line 52.''
+    source: {
+      file_name: null
+      current_line: ``87''
+      contents: null
+    }
+  }
+}
+\end{lstlisting}
+
+The command has been issued from a breakpoint set at \lstinline{main},
+which is at line 87 in \lstinline{memo.c}.  Since the source file
+hasn't changed, \lstinline{source.file_name} and
+\lstinline{source.contents} are set to \lstinline{null}.
+\lstinline{output} contains the response from rr, while
+\lstinline{channel}, \lstinline{command}, and \lstinline{from} contain
+information necessary to display commands correctly for all users.
+
+\begin{lstlisting}[language=Javascript,basicstyle=\linespread{0.5}\ttfamily,caption={Processing a Response},captionpos=b]
+  this.props.socket.on('rr_response', (data) => {
+    if(data.from !== this.props.socket.id){
+      // Display commands from other clients
+      this.term.disableStdin = true;
+      this.term.clear();
+      this.term.write(data.command);
+    };
+    // Write a newline, but no prompt
+    this.term.writeln('');
+    // Write each line of the output to the terminal
+    data.response.output.split(/\n+/).forEach((l) => {
+      this.term.writeln(l);
+    });
+    this.term.prompt();
+    // Re-enable input.  See above and below.
+    this.term.disableStdin = false;
+    // If an error has occured, there wont be any new
+    // information about the location in the source file.
+    if(data.response.source != null) {
+      this.updateSource(data.response.source);
+    }
+  });
+\end{lstlisting}
+
+When a response is received, \lstinline{RRTerm} first checks to see if
+the response has originated from a different user.  If so, input is
+disabled, the current line is cleared, and the command that
+corresponds to the response is written to the terminal.  Each line of
+the response is then written to the terminal.  Finally, input is
+renabled, having been disabled either to write commands received from
+another user, or when the command was first sent.
+\lstinline{updateSource} is called to update the Monaco editor source
+file view.  Since \lstinline{break 52} was sent from a location
+already inside \lstinline{memo.c}, the source file name and contents
+are not updated, though the line number is changed.
+
+\subsection{Considerations} \label{vistools}
+
+The intention of much of the code described above, such as that to
+disable input after a command has been sent, is to mimic rr in a way
+that is invisible to the user.  The user experience is intended to be
+that of using a fully-fledged debugger collaboratively.  The source
+code view allows users to easily see what code they are working on
+together, and the terminal updates quickly whenever a user enters a
+command.
+\par
+
+Care has been taken to design the frontend in a modular fashion so
+that it can be easily extended in the future.  By sending additional
+data about program state in the response and creating new componenets,
+visualization tools could be easily added.  If further authentication
+features or the ability to edit source code are desired in the future,
+relevant components can be updated without the need to change the
+entire application.  Since the collaborative debugger depends heavily
+on interactions between it's frontend and backend, similar care has
+been taken in designing the backend, which will be described in the
+following sections.
 
@@
 
-\subsection{RR Message Server}
+\subsection{RR Message Server}\label{rrmessageserver}
 
@@
 \par
+
+\subsection{Considerations}
 
@@
 
+\subsection{Considerations}
+
+The Frontend/API server connects all components of the collaborative
+debugger together.  It authenticates users through the database, joins
+the frontend and Kubernetes cluster through the dynamic creation of
+Pods, and starts the interaction between users and the programs they
+wish to debug.  Care has been taken to make it modular and extensible
+within reason, and additional authenication or pod creation features
+should be easy to add.
+\par
+
+The experience provided by invisible components such as the
+Frontend/API Server is just as important as that provided by the
+frontend webapp.  All components of the collaborative debugger
+hopefully come together to provide an experience that facilitates the
+teaching and learning of debugging.
+
 \section{Next Steps} \label{next}
 
-TODO
+\subsection{Features to Add}
+
+The current iteration of the collaborative debugger succesfully
+enables collaboration between students.  Significant effort has been
+taken to design the debugger in a way that makes collaboration
+seemless.  The collaborative debugger hopefully fulfills it's first
+requirment---to encourage collaboration.
+\par
+
+Though the collaborative debugger makes it easy for teachers to design
+and distribute debugging lessons, more work is necessary to fully
+fulfill the second goal---that of enhancing students' understanding of
+debugging and of programs being debugged.
+\par
+
+The most significant addition towards this goal would be that of
+frontend vizualization tools.  A simple visualization of the stack and
+registers would greatly expand students' understanding of the program
+being debugged in all three example sessions currently included with
+the collaborative debugger.  gdb/mi supports commands which give
+structured output of both these aspects of the execution space, and
+the frontend has been designed to make adding visualizations of them
+simple.  Adding visualization features will bring the collaborative
+debugger much closer towards hopefully fulfilling it's second
+requirement.
+
+\subsection{Testing Efficacy}
+
+Though a review of the research seems to indicate that a tool like the
+collaborative debugger should help make students more effective at
+debugging, further study is needed to draw conclusions.  First, more
+research is necessary to determine definitively if teaching debugging
+collaboratively is more effective than teaching it individually.
+Second, the collaborative debugger needs to be thoroughly tested to
+determine whether it is an ideal platform for collaborative debugging.
+\par
+
+A study similar to those undertaken in \cite{10.1145/1026487.1008043},
+\cite{10.1145/1145287.1145293}, and \cite{10.1145/3361721.3361724}
+could be done to test both these aspects.  Students participating in
+an introductory systems programming class could be split into three
+groups.  Group one could study debugging individually, using rr by
+themselves to progress through debugging sessions designed by the
+instructor.  Group two could study debugging in pairs, relying on
+existing tools such as video-conferencing software to facilitate
+collaboration.  In the final group, pairs would use the collaborative
+debugger to work together on the same debugging sessions as groups one
+and two.  Students' individual and pair performance in debugging both
+unseen and self-generated programs could then be tested.
+
+\subsection{Conclusions}
+
+The collaborative debugger aims to enourage collaboration and build
+students knowledge of debugging through careful design.  Many
+components of a distributed system come together to create a platform
+that hopefully benefits both teachers and students.  Through
+deliberate design, the collaborative debugger can conceivably be the
+ideal platform to teach and learn debugging, both remotely and
+in-person.
 

### 2020-12-06 paper/paper.tex
Paper before final read through

@@
   thoughtfulness about my project, his advice about the future, and
-  for giving me the oppurtunity to tutor for his classes.
+  for giving me the opportunity to tutor for his classes.
 \item Mar Hicks and Kyle Hale, professors whose classes at IIT changed
   my perspective on computer science.
-\item All my professors at Bard, particuarly Olga Voronina, Wakako
+\item All my professors at Bard, particularly Olga Voronina, Wakako
   Suzuki, and Nate Shockey, whose encouragement of my interests
@@
 
-Programmers spend large amounts of time debugging.  They often use
-debuggers to assist them in this process of diagnosing sometimes
-inexplicable problems with their code.  Though debuggers are primarily
-used to hunt down hiden mistakes, they are also powerful tools to
-understand program execution.
+Programmers spend large amounts of time debugging---the process of
+searching for and correcting errors in code that are not immediately
+apparent.  They often use debuggers---tools which help programmers to
+inspect a running program---to assist them in this process of
+diagnosing sometimes inexplicable problems with their code.  Though
+debuggers are primarily used to hunt down hidden mistakes, they are
+also powerful tools to understand program execution.
 \par
@@
 a recent rise in the inclusion of debugging in ``computational
-thinking'' curriculums\cite{10.1145/3361721.3361724}.  These
-curriculums attempt to teach skills in computer science classes that
+thinking'' curricula\cite{10.1145/3361721.3361724}.  These
+curricula attempt to teach skills in computer science classes that
 translate into other subject areas: the UK's computer science
@@
 
-Research into the value of collaborative programming is overwhelmingly
-positive.  McDowell et. al. found that not only does collaborative
-programming significantly boost student confidence and the retention
-of students in computer science majors, but that it demonstrably
-improves student's work \cite{10.1145/1145287.1145293}.  These
-benefits of confidence and correctness are present when paired and
-non-paired students are given identical assignments
+Collaborative programming, where multiple programmers work together to
+write and test code, is popular in both industry and computer science
+education\cite{10.1145/1145287.1145293}. Collaborative programming
+most often manifests as pair programming---two programmers working
+together on the same program.  Research into the value of pair
+programming is overwhelmingly positive.  McDowell et. al. found that
+not only does pair programming significantly boost student confidence
+and the retention of students in computer science majors, but that it
+demonstrably improves student's work \cite{10.1145/1145287.1145293}.
+These benefits of confidence and correctness are present when paired
+and non-paired students are given identical assignments
 \cite{10.1145/1026487.1008043}, further indicating that the simple act
@@
 There are similarities between the introductory nature of teaching
-fundemental debugging skills and the nature of teaching fundemental
+fundamental debugging skills and the nature of teaching fundamental
 programming skills covered in the classes studied in
@@
 confidence \textit{and} correctness that they do from pair programming
-in introductory comptuer science classes.
+in introductory computer science classes.
 \par
@@
 
-\section{Requirments}
+\section{Requirements}
 
 In order to provide the best platform for both teachers and students,
-the collaborative debugger must fufill two key requirements:
+the collaborative debugger must fulfill two key requirements:
 
@@
   value for each key-value pair in the hash pointing to the same
-  location in memory.  This is particuarly useful example for learning
+  location in memory.  This is particularly useful example for learning
   about setting watchpoints and breakpoints, as well as basic
@@
   difference between a recursive, iterative, and memoized recursive
-  implimentation of the classic nth fibonacci number function.  This
+  implementation of the classic nth Fibonacci number function.  This
   example focuses on examining the stack under conditions produced by
@@
 space, and data structures.  As discussed in \ref{vistools}, adding
-vizualization tools to the collaborative debugger should be relatively
+visualization tools to the collaborative debugger should be relatively
 easy.
@@
 
-Notice that the debugging window in the example shown above is called
-``rr''.  rr is an enhancment to gdb that is foundational in the
-collaborative debugger's ability to deterministically record and
-replay program execution. rr and its role in the collaborative
-debugger will be covered extensively in (\ref{rrpods}).  Users
-interact with rr through the collaborative debugger's frontend.  Care
-has been taken to make the frontend as representative of rr as
-possible so that users can learn to use rr/gdb by using the
-collaborative debugger.
-
 \subsection{Tools Used}
@@
 changes in state as users participate in the process outlined above.
-The frontend uses the following tools to manage state and display
-output:
+The frontend uses React, Monaco, and Xterm.js to manage state and
+display output.  It primarily displays the output from rr, an
+enhancement of gdb that the collaborative debugger uses to enable
+collaboration.
+
+\subsubsection{Mozilla's rr}\label{rr}
+
+Notice that the debugging window in the above example is called
+``rr''. rr is ``a lightweight tool for recording, replaying and
+debugging execution of applications''\cite{rr-repo}. rr allows a
+programmer to record the execution of a program on any compatible
+machine and replay the execution later.  This enhances GDB's ability
+to ``time-travel'' when debugging, using commands such as
+\lstinline{reverse-continue} and
+\lstinline{reverse-stepi}\cite{gdbman} to step backwards and forwards
+through a program's execution.  Through a novel encapsulation of the
+execution space, rr is able to deterministically record and replay the
+execution of syscalls and other process behavior that differs
+run-to-run.  This is invaluable when trying to debug behavior that is
+not entirely dependent on the code being debugged.  A typical workflow
+in rr consists of recording an inexplicable error, replaying execution
+to find the area in which the error occurs, and then narrowing in on
+the bug not by re-running the entire program, but by progressing back
+and forth through execution in the problem area.
+\par
+
+rr is an ideal tool for teaching debugging because it allows
+instructors to record execution of a program and design a debugging
+example with the knowledge that normally non-deterministic events will
+be repeatable, and that any input they provide to the program will be
+exactly replicated.  With the collaborative debugger, teachers can
+record a program's execution and design a debugging lesson which
+students can work on together.  The repeatability of rr means that
+students can focus on debugging, and teachers can create as specific
+examples as they please.  The use of rr is the most significant step
+the collaborative debugger takes to addressing the lack of
+back-tracing ability/coverage found in existing tools for teaching
+debugging\cite{10.1145/3286960.3286970}.
+
+In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
+that rely on capturing the entire state of of a virtual machine to
+replay execution, rr records and replays faster, produces far smaller
+files, and doesn't force execution inside of a
+VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The trade off for these
+benefits are two major system limitations: rr is only compatible with
+the Linux kernel, and it's deterministic recording and replay relies
+on a feature that is only found on modern \textit{Intel} x86 CPUs.
+These limitations influenced the development of this project as a
+webapp similar to existing tools for collaborative programming.
+\par
+
+Luckily, the speed and size benefits of rr lend themselves well to
+non-local execution.  In conjunction with Kubernetes (\ref{k8s}), it
+takes a few seconds to create a new container (\ref{docker}) running
+rr and connect to web clients.
 
@@
 defined in \lstinline{App.js} and the styles defined in
-\lstinline{App.css} into portable Javascript, HTML, and CSS.
+\lstinline{App.css} into portable JavaScript, HTML, and CSS.
 \par
@@
 want to display the call stack from within each implementation of the
-nth fibonacci number function.  To stop in \lstinline{fib_rec}, the
+nth Fibonacci number function.  To stop in \lstinline{fib_rec}, the
 recursive implementation, one user will want to set a breakpoint at
@@
 information about the current source file being debugged.
-\lstinline{command} is the most important field for examing how
+\lstinline{command} is the most important field for examining how
 commands are sent to the backend, as it holds the current command.
@@
 called again.  When the user hits the enter key the command is finally
-sent to the backend using \lstinline{sendCommand}.  To mimick the
+sent to the backend using \lstinline{sendCommand}.  To mimic the
 behavior of a local rr debug session and to prevent garbled input,
-writing to the terminal is disabled until a response is recieved.
+writing to the terminal is disabled until a response is received.
 
@@
 
-All users connected to the same debug session will recieive responses
+All users connected to the same debug session will receive responses
 to the commands sent by one user.  This is how collaboration is
 achieved in the collaborative debugger.  The JSON response object to
-\lstinline{break 52} recieved by all users is shown below:
+\lstinline{break 52} received by all users is shown below:
 
@@
 that it can be easily extended in the future.  By sending additional
-data about program state in the response and creating new componenets,
+data about program state in the response and creating new components,
 visualization tools could be easily added.  If further authentication
@@
 
-Kubernetes is the defacto standard in container orchestration
+Kubernetes is the de facto standard in container orchestration
 software.  It provides a layer of abstraction on top of normal
@@
 instance in the collaborative debugger.  This allows the debugger to
-provide a similar level of convinience to exisiting collaborative
-tools, such repl.it (\ref{exisitingcollab}).
+provide a similar level of convenience to existing collaborative
+tools, such Replit (\ref{exisitingcollab}).
 \par
@@
 backend and frontend React webapp.  Kubernetes was chosen for the
-backend primarally so that a Pod could be created dynamically for each
+backend primarily so that a Pod could be created dynamically for each
 debugging session.
@@
   exposes the RR Message Server Pods to the outside world.  Using
-  Socket.IO, clients send commands to and recieve responses from
+  Socket.IO, clients send commands to and receive responses from
   individual RR Debug Session Pods through the RR Message Server Pods.
@@
   which exposes the Frontend/API Server Pods to the outside world.
-  Clients request the frontend webapp and send api requests/recieve
-  api responses through this Load Balancer.
+  Clients request the frontend webapp and send API requests/receive
+  API responses through this Load Balancer.
 \end{enumerate}
@@
   Session images are based on the latest Ubuntu image.  This is
-  particuarly necessary for the RR Debug Session image, as rr's
+  particularly necessary for the RR Debug Session image, as rr's
   low-level nature necessitates somewhat frequent updates as changes
@@
 Services are the first components of the collaborative debugger
-created after the cluster is initiallized.  The three Services used by
+created after the cluster is initialized.  The three Services used by
 the debugger are all statically created.  Like statically created
@@
 This five digit number, is used as the channel for the RR Debug
-Session, seperating its communication from other RR Debug Session
+Session, separating its communication from other RR Debug Session
 Pods.  Clients, RR Message Server Pods, and the RR Debug Session Pod
@@
   example), sends an rr command.
-\item The RR Message Server recieves the command and emits a message
-  that RR Debug Session Pods are equiped to recieve on the same
+\item The RR Message Server receives the command and emits a message
+  that RR Debug Session Pods are equipped to receive on the same
   channel.  In practice, since only one RR Debug Session Pod is ever
-  on a channel, this equates to emmiting a message to the Pod
+  on a channel, this equates to emitting a message to the Pod
   directly.
-\item The RR Debug Session Pod recieves the message, and passes the
-  command to its instance of RRInterface.  When it recieves a
+\item The RR Debug Session Pod receives the message, and passes the
+  command to its instance of RRInterface.  When it receives a
   response, it emits the output from rr along with other debugging
   information and the channel.
-\item The RR Message Server recieves the response, and emits a
-  response message that clients are equipped to recieve on the
+\item The RR Message Server receives the response, and emits a
+  response message that clients are equipped to receive on the
   corresponding channel.
 \item Clients A and B are connected to the corresponding channel, so
-  they both recieve the response.
+  they both receive the response.
 \end{enumerate}
@@
 
-The collaborative debugger needs an effecient way for clients to
+The collaborative debugger needs an efficient way for clients to
 communicate with running RR Debug Sessions. The Socket.IO library was
@@
 directly connect web clients and the Pods running rr.  Socket.IO is a
-library that extends WebSockets.  It provides backup incase a
+library that extends WebSockets.  It provides backup in case a
 WebSocket connection cannot be established, enables automatic
@@
 \lstinline{on_rr_response}, are equally simple.  When the RR Message
-Server recieves an \lstinline{'rr_command'} message, it passes the
+Server receives an \lstinline{'rr_command'} message, it passes the
 message data along to the corresponding RR Debug Session Pod by
@@
 \lstinline{'rr_command'}s but frontend clients do not, only RR Message
-Server Pods recieve the command.  The same process happens in reverse
+Server Pods receive the command.  The same process happens in reverse
 for \lstinline{'rr_response'}s, with only the frontend socket.io
@@
 \subsubsection{Tools Used}
-
-\paragraph{Mozilla's rr}\label{rr}
-
-rr is ``a lightweight tool for recording, replaying and debugging
-execution of applications''\cite{rr-repo}. rr allows a programmer to
-record the execution of a program on any compatible machine and replay
-the execution later.  This enhances GDB's ability to ``time-travel''
-when debugging, using commands such as \lstinline{reverse-continue}
-and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
-forwards through a program's execution.  Through a novel encapsulation
-of the execution space, rr is able to deterministically record and
-replay the execution of syscalls and other process behavior that
-differs run-to-run.  This is invaluable when trying to debug behavior
-that is not entirely dependent on the code being debugged.  A typical
-workflow in rr consists of recording an inexplicable error, replaying
-execution to find the area in which the error occurs, and then
-narrowing in on the bug not by re-running the entire program, but by
-progressing back and forth through execution in the problem area.
-\par
-
-rr is an ideal tool for teaching debugging because it allows
-instructors to record execution of a program and design a debugging
-example with the knowledge that normally non-deterministic events will
-be repeatable, and that any input they provide to the program will be
-exactly replicated.  With the collaborative debugger, teachers can
-record a program's execution and design a debugging lesson which
-students can work on together.  The repeatability of rr means that
-students can focus on debugging, and teachers can create as specific
-examples as they please.  The use of rr is the most significant step
-the collaborative debugger takes to addressing the lack of
-back-tracing ability/coverage found in existing tools for teaching
-debugging\cite{10.1145/3286960.3286970}.
-
-In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
-that rely on capturing the entire state of of a virtual machine to
-replay execution, rr records and replays faster, produces far smaller
-files, and doesn't force execution inside of a
-VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The trade off for these
-benefits are two major system limitations: rr is only compatible with
-the Linux kernel, and it's deterministic recording and replay relies
-on a feature that is only found on modern \textit{Intel} x86 CPUs.
-These limitations influenced the development of this project as a
-webapp similar to existing tools for collaborative programming.
-\par
-
-Luckily, the speed and size benefits of rr lend themselves well to
-non-local execution.  In conjunction with Kubernetes, it takes a few
-seconds to create a new container running rr and connect to web
-clients.
 
@@
 as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging Pod.
-pygdbmi's abstraction simplifies programatically controlling rr.  A
+pygdbmi's abstraction simplifies programmatically controlling rr.  A
 Pod can receive a command from the client, pass it to rr, and respond
@@
 slightly from other images used by the collaborative debugger.
-Currently, three example RR Debug Session images are availiable for
+Currently, three example RR Debug Session images are available for
 use.  Example container images are based on the primary RR Debug
@@
 After joining the appropriate channel, \lstinline{sio} waits to
-recieve an \lstinline{'rr_command'}.  The first part of
+receive an \lstinline{'rr_command'}.  The first part of
 \lstinline{sio}'s \lstinline{'rr_command'} event handler is shown below:
@@
 for the RR Message Server and client later.  Next, it calls the
-function necessary to pass a message to rr and recieve a response,
+function necessary to pass a message to rr and receive a response,
 \lstinline{rri.write()}.
@@
 of all commands the debug session has executed.  Though unused at the
-moment, it can serve in the future to impliment a shared history
+moment, it can serve in the future to implement a shared history
 between all clients and to facilitate the saving of debug sessions
@@
 identical state to a previous RR Debug Session Pod.  This appears to
-clients as a seemless restoration of a previous Debug Session.
+clients as a seamless restoration of a previous Debug Session.
 \par
@@
 the collaborative debugger that support visualization tools could use
-gdbmi commands such as \lstinline{-stack-list-frames} to retrieve
+gdb/mi commands such as \lstinline{-stack-list-frames} to retrieve
 information to be used by frontend visualization tools.
@@
 method.  The \lstinline{get_full_rr_response} method ensures that all
-relavant output has been recieved from the rr subprocess before
+relevant output has been received from the rr subprocess before
 returning.  Write reduces the lengthy dictionary that would be
@@
 
-After the \lstinline{'rr_command'} event handler has recieved a
+After the \lstinline{'rr_command'} event handler has received a
 response from \lstinline{rri}, it calls \lstinline{RRInterface}'s
@@
 name, and contents.  If an error occurs at any point in the process, a
-detailed trace is emmitted for debugging purposes.  The trace should
-be ommited in production.
+detailed trace is emitted for debugging purposes.  The trace should
+be omitted in production.
 \par
@@
 parsing rr output.  The flow of data in the program is quite simple: a
-command is recieved, it is passed to rr, rr's response is processed,
-and a response message is emmited.  This flow would remain unchanged
+command is received, it is passed to rr, rr's response is processed,
+and a response message is emitted.  This flow would remain unchanged
 even with the addition of functions to retrieve information for data
@@
 Effort is taken in the design of the RR Debug Session and RR Message
-Server to ensure students recieve a pleasant and near-identical
+Server to ensure students receive a pleasant and near-identical
 experience to debugging locally in the future.  The system of
@@
 Since the entire webapp is one page, the process for serving it is
-completely standard.  The server recieves a request for the index page
+completely standard.  The server receives a request for the index page
 of the website, and returns the compiled React webapp that makes up
@@
 
-The API server is implimented using Flask.  The server program
+The API server is implemented using Flask.  The server program
 consists of handlers for the various API routes and instances of two
@@
 would be to break the \lstinline{UserManager} and
-\lstinline{TranslationPodManager} instances out into seperate programs
+\lstinline{TranslationPodManager} instances out into separate programs
 running on their own Pods in the cluster.  This would allow changes to
@@
 All the API request handlers listed above return an error message in
-the case of an error occuring during the processing of an API request.
+the case of an error occurring during the processing of an API request.
 Since most errors that could occur would either be the result of
-unanticipated issues on the part of the user, or of some unforseen
-catostrophic internal error, informing users of error specifics would
+unanticipated issues on the part of the user, or of some unforeseen
+catastrophic internal error, informing users of error specifics would
 be unhelpful.  Some \lstinline{TranslationPodManager} and
@@
 All API routes only support POST requests.  The route handler first
-extracts the relavant information from the POST request body, in this
+extracts the relevant information from the POST request body, in this
 case \lstinline{name} and \lstinline{program}.  \lstinline{name}
@@
 \lstinline{examples} are independent of \lstinline{users} and
-\lstinline{pods}.  The continer registry URLs of RR Debug Session
+\lstinline{pods}.  The container registry URLs of RR Debug Session
 example images are stored in the database so that new images can be
@@
 example image names and container registry URLs from the database.  In
-the event that the user has somehow pased a spurious image name or has
+the event that the user has somehow passed a spurious image name or has
 passed a name when there are no images, an exception is raised.
@@
 slowing as more channels are used.  Given the few current users, the
-likelyhood of a collision when generating a new channel is low enough
+likelihood of a collision when generating a new channel is low enough
 that a more advanced method is unnecessary.
@@
 isolate dynamically and statically created Pods with minimal hassle,
-since Kubernetes' DNS resolution can connect pods accros namespaces.
+since Kubernetes' DNS resolution can connect pods across namespaces.
 If Pod creation fails, an error will be thrown.  Otherwise, a loop is
@@
 Most request handlers, \lstinline{'/new'} included, simply return a
-JSON serializable dicitionary containing whatever results the API
+JSON serializable dictionary containing whatever results the API
 Server's instance of \lstinline{TranslationPodManager} or
@@
 wish to debug.  Care has been taken to make it modular and extensible
-within reason, and additional authenication or pod creation features
+within reason, and additional authentication or pod creation features
 should be easy to add.
@@
 
+\subsection{Current Functionality}
+
+The following section details the first few steps students take to
+explore the \lstinline{memoization} example using the collaborative
+debugger.  Students A and B begin by examining the call stack at its
+deepest point in the recursive nth Fibonacci number function.  The
+terminal interface to rr and the source code view have been
+color-inverted to improve readability in this paper.
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session1i}
+  \centering
+  \caption{Creating a new Memoization Example Session}
+  \label{session1i}
+\end{figure}
+
+To start a new session, Student A selects ``memoization'' from the
+list of example sessions.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session2i}
+  \centering
+  \caption{Joining the Memoization Example Session}
+  \label{session2i}
+\end{figure}
+
+In order to join Student A's session, Student B uses session code
+``71825''.  Meanwhile, Student A is presented with the the main view
+of the collaborative debugger.  On the right is a view of the source
+code for the \lstinline{memoization} example.  On the left is the
+terminal interface to rr.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session3i}
+  \centering
+  \caption{Observing Output from the Memoization Example Session}
+  \label{session3i}
+\end{figure}
+
+Student B now types \lstinline{c} to continue execution.  Both
+students see the command and output: execution times and results for
+three different implementations of nth Fibonacci.  An advantage of
+rr's recording model is that these numbers will be the same every time
+the students run the program.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session4i}
+  \centering
+  \caption{Setting a Breakpoint in the Memoization Example Session}
+  \label{session4i}
+\end{figure}
+
+Student A enters \lstinline{break 52} to set a breakpoint at line 52,
+the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
+and then \lstinline{c} to restart the program and continue to the
+breakpoint.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session5i}
+  \centering
+  \caption{Examining the Stack in the Memoization Example
+    Session}
+  \label{session5i}
+\end{figure}
+
+The students examine the stack resulting from a call of
+\lstinline{fib_rec(30)}.  Students can inspect different areas of the
+program and output while collaborating---Student A can look at
+\lstinline{fib_rec}, while Student B scrolls up for a preview of the
+memoized implementation.
+\par
+
+The students will continue to inspect the stack for other functions,
+exploring the reasons for the vast performance disparity between
+\lstinline{fib_memo} and \lstinline{fib_rec}.  Students A and B can
+not only help each other to learn about the stack and memoization, but
+can also teach one another how to use a debugger.  Because the
+frontend produces near-identical output to rr, students can directly
+translate their knowledge to it or gdb in the future.  Paradigms, such
+as examining the stack to understand execution, can be expanded to all
+future debugging.
+
 \subsection{Features to Add}
 
-The current iteration of the collaborative debugger succesfully
+The current iteration of the collaborative debugger successfully
 enables collaboration between students.  Significant effort has been
 taken to design the debugger in a way that makes collaboration
-seemless.  The collaborative debugger hopefully fulfills it's first
-requirment---to encourage collaboration.
+seamless.  The collaborative debugger hopefully fulfills it's first
+requirement---to encourage collaboration.
 \par
@@
 The most significant addition towards this goal would be that of
-frontend vizualization tools.  A simple visualization of the stack and
+frontend visualization tools.  A simple visualization of the stack and
 registers would greatly expand students' understanding of the program
@@
 
-The collaborative debugger aims to enourage collaboration and build
+The collaborative debugger aims to encourage collaboration and build
 students knowledge of debugging through careful design.  Many

### 2020-12-06 paper/paper.tex
Final updates to paper

@@
 \begin{abstract}
-  \linespread{1}
-  Debugging is rarely formally taught, despite being used by
-  programmers every day.  Research indicates that it is valuable to
-  teach debugging, and suggests that teaching it collaboratively may
-  be maximally effective.  The goal of this project is to create a
-  collaborative debugger.  The debugger aims to be the ideal platform
-  to teach and learn debugging.  This paper briefly reviews relevant
-  literature which covers teaching debugging and that which focuses on
+  \linespread{1} Debugging is rarely formally taught, despite being
+  used by programmers every day.  Research indicates that it is
+  valuable to teach debugging, and suggests that teaching it
+  collaboratively may be maximally effective.  The goal of this
+  project is to create a collaborative debugger.  The debugger aims to
+  be the ideal platform to teach and learn debugging.  This paper
+  briefly reviews relevant literature covering teaching debugging and
   teaching programming collaboratively.  Most of the paper is devoted
@@
 \item All my professors at Bard, particularly Olga Voronina, Wakako
-  Suzuki, and Nate Shockey, whose encouragement of my interests
+  Suzuki, and Nathan Shockey, whose encouragement of my interests
   outside of computer science has been instrumental in continuing my
@@
 Debugging is invaluable in writing and understanding code, yet it is
-rarely formally taught\cite{doi:10.1080/08993400802114581}.  We
-typically teach students programming structures, concepts, and
-languages, but leave them to learn the tools they use to write code by
-themselves.  This approach often works well---a programmer's choice of
-tools is often \textit{very} personal and students figure out how to
-configure an individualized workflow.  Perhaps because debuggers are
-tools, students are often expected to learn them with minimal
-guidance.  Unlike editors or reference guides however, effectively
-using a debugger requires a set of high-level, platform agnostic,
-teachable skills.  Teaching these skills is effective, and translates
-into better, faster, debugging and
+rarely formally taught\cite{doi:10.1080/08993400802114581}.  Students
+are typically taught programming structures, concepts, and languages,
+but are left to learn the tools they use to write code alone.  This
+approach often works well---a programmer's choice of tools is often
+\textit{very} personal and students figure out how to configure an
+individualized workflow.  Perhaps because debuggers are tools,
+students are often expected to learn them with minimal guidance.
+Unlike editors or reference guides however, effectively using a
+debugger requires a set of high-level, platform agnostic, teachable
+skills.  Teaching these skills is effective, and translates into
+better and faster debugging and
 programming\cite{10.1145/3286960.3286970}\cite{10.1145/3361721.3361724}.
@@
 debugging sessions are hosted remotely, users can start a session in
-one location and resume it later from a different computer.
+one location and resume it later from a different computer.  This is
+particularly beneficial for students who may not always have access to
+a computer or a machine capable of running the programs they wish to
+debug.
 \par
@@
   value for each key-value pair in the hash pointing to the same
-  location in memory.  This is particularly useful example for learning
-  about setting watchpoints and breakpoints, as well as basic
+  location in memory.  This is a particularly useful example for
+  learning about setting watchpoints and breakpoints, as well as basic
   procedures to step through code.
@@
 \item \lstinline{memoization}: a program that demonstrates the
-  difference between a recursive, iterative, and memoized recursive
+  difference between a recursive, iterative, and memoized-recursive
   implementation of the classic nth Fibonacci number function.  This
@@
 the future, as well as one which is seamless to encourage the
-acquisition of debugging skills.  Creating such a platform will
-hopefully encourage both students and teachers to explore the
+acquisition of general debugging skills.  Creating such a platform
+will hopefully encourage both students and teachers to explore the
 undertaught area of debugging.
@@
 helpful to look at the steps two students would take to debug the
-\lstinline{memoization} example program remotely:
+\lstinline{memoization} example program remotely.  This example is
+covered in more detail in \ref{current}.
 
@@
 Luckily, the speed and size benefits of rr lend themselves well to
-non-local execution.  In conjunction with Kubernetes (\ref{k8s}), it
-takes a few seconds to create a new container (\ref{docker}) running
-rr and connect to web clients.
+non-local execution.  In conjunction with tools used to create the
+backend of the collaborative debugger it takes a few seconds to create
+a new debug session running rr and connect it to the frontend webapp
+discussed in this section.
 
@@
 
+% The frontend webapp communicates with a distributed backend.  A simple
+% diagram of the entire system is below.
+
+% \begin{figure}[h!]
+
+%   \includegraphics[scale=.8]{overall_system}
+%   \centering
+%   \caption{Overview of the Collaborative Debugger}
+%   \label{debugger:overview}
+% \end{figure}
+
+% The collaborative debugger consists of:
+
+% \begin{enumerate}
+% \item A frontend web app that presents a debugging interface to the
+%   end user.
+% \item A distributed backend split into three parts:
+%   \begin{enumerate}
+%   \item 
+%   \item Pods running a frontend server written in Node.js which works
+%     in tandem with an API server created using Flask
+%     (\ref{flask/node}).  The API server manages creation and
+%     destruction of debug sessions, as well as authentication.
+%   \item A MongoDB (\ref{mongodb}) database.
+%   \end{enumerate}
+% \end{enumerate}
+
 React makes it easy to create simple webapps with no build
@@
 collaborative debugger.  Once this structure was created and any
-superfluous files/code were removed, it's wasn't necessary for the
+superfluous files/code were removed, it wasn't necessary for the
 purposes of the collaborative debugger to directly edit any files
@@
 This section will outline the process of sending an rr command to the
-backend and displaying a response.
+backend and displaying a response.  Apart from sending simple API
+requests to start sessions or log users in and out, this communication
+process is where the frontend webapp spends most of its time.
 \par
@@
 \lstinline{Breakpoint 1 at 0x561c8fede303: file memo.c, line 52.}\\
-Users who did not send the command will see it displayed along with
-the response.
+Users who did not send the command \lstinline{c} will see it displayed
+along with the response.
 
@@
 Care has been taken to design the frontend in a modular fashion so
-that it can be easily extended in the future.  By sending additional
-data about program state in the response and creating new components,
-visualization tools could be easily added.  If further authentication
-features or the ability to edit source code are desired in the future,
-relevant components can be updated without the need to change the
-entire application.  Since the collaborative debugger depends heavily
-on interactions between it's frontend and backend, similar care has
-been taken in designing the backend, which will be described in the
-following sections.
+that it can be easily extended in the future.  Visualization tools
+could be easily added by sending additional data about program state
+in the response and creating new components.  If further
+authentication features or the ability to edit source code are desired
+in the future, relevant components can be updated without the need to
+change the entire application.  Since the collaborative debugger
+depends heavily on interactions between its frontend and backend,
+similar care has been taken in designing the backend, which will be
+described in the following sections.
 
@@
 deployment and re-deployment of applications running inside
-containers.  It is trivial to create new Pods, or to create multiple
-Pods running the same application as needed within a Kubernetes
-cluster \cite{k8s}.  The speed at which even relatively large Pods can
-be created and the inherent security provided by containerization
-drove the decision to create a new Pod on the fly for each debugging
+containers.  It is trivial to create new Pods (or multiple Pods
+running the same application) as needed within a Kubernetes cluster
+\cite{k8s}.  The speed at which even relatively large Pods can be
+created and the inherent security provided by containerization drove
+the decision to create a new Pod on the fly for each debugging
 instance in the collaborative debugger.  This allows the debugger to
@@
 
-\subsubsection{Cluster Selection and Configuration}
+\subsubsection{Cluster Requirements and Configuration}
 
@@
 rr to be able to work efficiently, the
-\lstinline{kernel.perf_event_paranoid} parameter must be set to 1
-\cite{rr-repo}.  This should be done on every node in the cluster
-which will run RR Debug Sessions.  For the purposes of development, it
-has been set to 1 on all nodes in the collaborative debugger cluster.
+\lstinline{kernel.perf_event_paranoid} parameter must be set to
+\lstinline{1} \cite{rr-repo}.  This should be done on every node in
+the cluster which will run RR Debug Sessions.  For the purposes of
+development, it has been set to \lstinline{1} on all nodes in the
+collaborative debugger cluster.
 
@@
 \item Second, any necessary packages are installed.  For the RR Debug
-  Pod image, rr is compiled from source and installed.
+  Session image, rr is compiled from source and installed.
 \item A non-root user is created if necessary.
 \item Program files are copied over and a file structure is created.
-  Packages that don't rely on the base images builtin package manager
-  are installed at this time.  In the example above, these include
-  Python packages.
+  Packages that don't rely on the base image's built in package
+  manager are installed at this time.  In the example above, these
+  include Python packages.
 \item A startup command is defined.
@@
 assigned a unique five digit identification number when it is created.
-This five digit number, is used as the channel for the RR Debug
+This five digit number is used as the channel for the RR Debug
 Session, separating its communication from other RR Debug Session
@@
 
-  \includegraphics[scale=1]{rr_detailed}
+  \includegraphics[scale=.8]{rr_detailed}
   \centering
@@
 \item The RR Debug Session Pod receives the message, and passes the
-  command to its instance of RRInterface.  When it receives a
-  response, it emits the output from rr along with other debugging
-  information and the channel.
+  command to its instance of RRInterface, the class which controls rr.
+  When it receives a response, it emits the output from rr along with
+  other debugging information and the channel.
 \item The RR Message Server receives the response, and emits a
@@
 
+RR Debug Sessions primarily manage and communicate with rr, which has
+already been covered in detail in \ref{rr}.  pygdbmi is used for this
+process.
+
 \paragraph{pygdbmi}
@@
 Pod can receive a command from the client, pass it to rr, and respond
-without having to deal with parsing gdb/mi output or with directly
-managing the rr process.
+without having handle directly parsing gdb/mi output or managing the
+rr process.
 
@@
 Currently, three example RR Debug Session images are available for
-use.  Example container images are based on the primary RR Debug
-Session image, RR Translation.  The build process for this image,
-based on the lasted official Ubuntu container image, installs rr,
-python, and all necessary packages as well as the app that will run on
-the final image, \lstinline{rrtranslation.py}.
+use: \lstinline{hash}, \lstinline{smash}, and \lstinline{memoization}.
+Example container images are based on the primary RR Debug Session
+image, RR Translation.  The build process for this image, based on the
+lasted official Ubuntu container image, installs rr, python, and all
+necessary packages as well as the app that will run on the final
+image, \lstinline{rrtranslation.py}.
 
@@
 
-The Dockerfile shown above is for a simple hash program RR Debug
+The Dockerfile shown above is for the \lstinline{hash} RR Debug
 Session example.  The build process copies over any necessary files,
@@
 Current example pod startup scripts compile the program to be debugged
-and record execution when the pod is created.  It is easy to create a
-more specific example to be replayed, if exact reproduction of
-syscalls or other non-deterministic behavior is desired.  The
-example's creator can record an rr session locally ahead of building
-the example Pod and copy the recording into the appropriate folder
-during the build process.  Since rr recordings are portable, the
-example will replay without issue in the RR Debug Session Pod.
+and record execution when the pod is created.
 
@@
 
-The RR Translation program consists of two main components: a socket.io
-client, \lstinline{sio}, and an instance of the
+The RR Translation program consists of two main components: a
+Socket.IO client, \lstinline{sio}, and an instance of the
 \lstinline{RRInterface} class, \lstinline{rri}, that controls the rr
@@
 connecting to the RR Message Server.  It immediately emits a
-\lstinline{'join_channel'} message, using the channel passed in from
-the Kubernetes API call on it's creation.
+\lstinline{'join_channel'} message, using the channel number passed in
+by the Kubernetes API on Pod creation (\ref{dynamicc}).
 
@@
 \textit{recordings}, not currently executing programs that
-\lstinline{shell} could effect are being debugged, and it opens the
+\lstinline{shell} could effect, are being debugged.  It also opens the
 door to security issues.  Though containerization means that the most
@@
 Since the frontend currently consists of a terminal interface to rr
-and a view of the current source code, write does not need to return
-any information from rr other than console output.  Future versions of
-the collaborative debugger that support visualization tools could use
-gdb/mi commands such as \lstinline{-stack-list-frames} to retrieve
-information to be used by frontend visualization tools.
+and a view of the current source code, \lstinline{write} does not need
+to return any information from rr other than console output.  Future
+versions of the collaborative debugger that support visualization
+tools could use gdb/mi commands such as \lstinline{-stack-list-frames}
+to retrieve information to be used by frontend visualization tools.
 \lstinline{console_output} extracts user-relevant output from the
@@
 relevant output has been received from the rr subprocess before
-returning.  Write reduces the lengthy dictionary that would be
-produced by a command as simple as \lstinline{break main} into a few
-lines of user-relevant output.
+returning.  \lstinline{write} reduces the lengthy dictionary that
+would be produced by a command as simple as \lstinline{break main}
+into a few lines of user-relevant output.
 \par
@@
 be as simple as possible, though a more fully featured frontend could
-simplify this process further in the future.
+streamline this process further in the future.
 
@@
 
-  \includegraphics[scale=1]{login_request}
+  \includegraphics[scale=.9]{login_request}
   \centering
@@
 All the API request handlers listed above return an error message in
-the case of an error occurring during the processing of an API request.
-Since most errors that could occur would either be the result of
-unanticipated issues on the part of the user, or of some unforeseen
-catastrophic internal error, informing users of error specifics would
-be unhelpful.  Some \lstinline{TranslationPodManager} and
-\lstinline{UserManager} functions throw specific errors in the event
-of duplicate usernames, channels that do not exist, etc.  These errors
-are caught and meaningful error messages are returned to the frontend
-webapp, which can then pass them on to the user.
+the case of an error occurring during the processing of an API
+request.  Since most errors that could occur would either be the
+result of unanticipated issues on the part of the user, or of some
+unforeseen catastrophic internal error, informing users of error
+specifics is often unhelpful.  Some \lstinline{TranslationPodManager}
+and \lstinline{UserManager} functions throw specific errors in the
+event of duplicate usernames, channels that do not exist, etc.  These
+errors are caught and meaningful error messages are returned to the
+frontend webapp, which can then pass them on to the user.
 \par
@@
 MongoDB keeps \lstinline{_id}s unique.  In addition, the collaborative
-debugger defines uniqueness constrains on \lstinline{users:name},
-\lstinline{pods:channel}, and \lstinline{examples:name} when the
+debugger defines uniqueness constrains on \lstinline{users.name},
+\lstinline{pods.channel}, and \lstinline{examples.name} when the
 database is created.
@@
 
-\subsubsection{Processing a POST Request---Dynamic Pod Creation}
+\subsubsection{Processing a POST Request---Dynamic Pod Creation} \label{dynamicc}
 
@@
 \lstinline{channel} for the Pod, using the uniqueness constraint
-imposed on \lstinline{pods:channel} to ensure an unused
+imposed on \lstinline{pods.channel} to ensure an unused
 \lstinline{channel} is generated.  If no channels are open an error is
@@
 This process has the effect of limiting the number of simultaneous RR
-Debug Sessions that can be in use at the same time to 89999, and of
+Debug Sessions that can be in use at the same time to 89,999, and of
 slowing as more channels are used.  Given the few current users, the
@@
 
-\subsection{Current Functionality}
+\subsection{Current Functionality} \label{current}
 
@@
 the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
-and then \lstinline{c} to restart the program and continue to the
+and then \lstinline{c} to restart the recording and continue to the
 breakpoint.\pagebreak

### 2020-12-07 paper/paper.tex
final paper

@@
 A large number of tools exist to facilitate collaborative programming.
-COVID19 has greatly increased the demand for tools that not only make
+COVID-19 has greatly increased the demand for tools that not only make
 collaboration easier, but make it easy remotely.  The tool that was
@@
 
-\lstinline{memoization} will be covered in detail in the following
-section on the frontend of the collaborative debugger (\ref{webapp}).
-All three examples show how teachers can design debugging sessions
-around specific concepts.  The collaborative debugger should make it
-easy to distribute these sessions and for students to learn from them
-collaboratively.
-\par
-
-Students using all three examples would benefit greatly from
-visualization tools to help understand pointer locations, the process
-space, and data structures.  As discussed in \ref{vistools}, adding
-visualization tools to the collaborative debugger should be relatively
-easy.
+% \lstinline{memoization} will be covered in detail in the following
+% section on the frontend of the collaborative debugger (\ref{webapp}).
+% All three examples show how teachers can design debugging sessions
+% around specific concepts.
+The collaborative debugger should make it easy to distribute these
+sessions and for students to learn from them collaboratively.
+\par
+
+\subsection{Current Functionality} \label{current}
+
+The following section details the first few steps students take to
+explore the \lstinline{memoization} example using the current version
+of the collaborative debugger.  This example, which focuses more on
+learning about how functions execute than on fixing bugs, highlights
+the ways in which the collaborative debugger enables pair programming.
+\par
+
+Students A and B begin by examining the call stack at its deepest
+point in the recursive nth Fibonacci number function.  The terminal
+interface to rr and the source code view have been color-inverted to
+improve readability in this paper.
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session1i}
+  \centering
+  \caption{Creating a new Memoization Example Session}
+  \label{session1i}
+\end{figure}
+
+Students A and B both log into the collaborative debugger.  To start a
+new session, Student A selects ``memoization'' from the list of
+example sessions.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session2i}
+  \centering
+  \caption{Joining the Memoization Example Session}
+  \label{session2i}
+\end{figure}
+
+In order to join Student A's session, Student B uses session code
+``71825''.  Meanwhile, Student A is presented with the the main view
+of the collaborative debugger.  On the left is a view of the source
+code for the \lstinline{memoization} example.  On the right is the
+terminal interface to rr, the enhancement to gdb used by the
+collaborative debugger.  Students debug recordings of program
+execution, which can be replayed deterministically. rr is covered in
+detail in \ref{rr}.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session3i}
+  \centering
+  \caption{Observing Output from the Memoization Example Session}
+  \label{session3i}
+\end{figure}
+
+Student B now types \lstinline{c} to continue execution.  Both
+students see the command and output: execution times and results for
+three different implementations of nth Fibonacci.  An advantage of
+rr's recording model is that these numbers will be the same every time
+the students run the program.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session4i}
+  \centering
+  \caption{Setting a Breakpoint in the Memoization Example Session}
+  \label{session4i}
+\end{figure}
+
+Student A enters \lstinline{break 52} to set a breakpoint at line 52,
+the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
+and then \lstinline{c} to restart the recording and continue to the
+breakpoint.\pagebreak
+
+\begin{figure}[h!]
+
+  \includegraphics[width=\textwidth]{session5i}
+  \centering
+  \caption{Examining the Stack in the Memoization Example
+    Session}
+  \label{session5i}
+\end{figure}
+
+The students examine the stack resulting from a call of
+\lstinline{fib_rec(30)}.  Students can inspect different areas of the
+program and output while collaborating---Student A can look at
+\lstinline{fib_rec}, while Student B scrolls up for a preview of the
+memoized implementation.
+\par
+
+The students will continue to inspect the stack for other functions,
+exploring the reasons for the vast performance disparity between
+\lstinline{fib_memo} and \lstinline{fib_rec}.  Students A and B can
+not only help each other to learn about the stack and memoization, but
+can also teach one another how to use a debugger.  Because the
+frontend produces near-identical output to rr, students can directly
+translate their knowledge to it or gdb in the future.  Paradigms, such
+as examining the stack to understand execution, can be expanded to all
+future debugging.
+
+\subsection{Considerations}
+
+Though the collaborative debugger makes it easy for students in the
+above example to practice using a debugger together, more features
+could be added to help improve students' understanding of the program
+being debugged. Students using all three examples would benefit
+greatly from visualization tools to help understand pointer locations,
+the process space, and data structures.
+\par
+
+In \lstinline{memoization}, a simple stack visualization that updated
+with each command sent to the debugger would allow students to more
+easily see changes in the process space.  Rather than having to
+repeatedly \lstinline{continue} and \lstinline{backtrace} to visualize
+the stack, students could simply step through relevant portions of the
+program and watch as the stack updated.  This sort of visualization
+pairs particularly well with \lstinline{reverse-continue} or
+\lstinline{reverse-step}(\lstinline{i}), allowing students to move back and forth
+over particularly significant stack changes.  Visualization tools also
+aid students' understanding of debugging---students can see how typing
+\lstinline{bt} to backtrace or \lstinline{info frame} to view
+information about the current stack frame provide the same information
+as the visualization tools they are already familiar with.  As
+discussed in \ref{vistools}, adding visualization tools to the
+collaborative debugger should be relatively easy.
 \par
@@
 
-Before diving into the construction of the frontend webapp, it may be
-helpful to look at the steps two students would take to debug the
-\lstinline{memoization} example program remotely.  This example is
-covered in more detail in \ref{current}.
+An overview of the \lstinline{memoization} example discussed in
+\ref{current} is below.  This pattern is followed by all debugging
+sessions in the frontend of the collaborative debugger.
 
@@
 \lstinline{reverse-continue} and
-\lstinline{reverse-stepi}\cite{gdbman} to step backwards and forwards
-through a program's execution.  Through a novel encapsulation of the
-execution space, rr is able to deterministically record and replay the
-execution of syscalls and other process behavior that differs
-run-to-run.  This is invaluable when trying to debug behavior that is
-not entirely dependent on the code being debugged.  A typical workflow
-in rr consists of recording an inexplicable error, replaying execution
-to find the area in which the error occurs, and then narrowing in on
-the bug not by re-running the entire program, but by progressing back
-and forth through execution in the problem area.
+\lstinline{reverse-step}(\lstinline{i})\cite{gdbman} to step backwards
+and forwards through a program's execution.  Through a novel
+encapsulation of the execution space, rr is able to deterministically
+record and replay the execution of syscalls and other process behavior
+that differs run-to-run.  This is invaluable when trying to debug
+behavior that is not entirely dependent on the code being debugged.  A
+typical workflow in rr consists of recording an inexplicable error,
+replaying execution to find the area in which the error occurs, and
+then narrowing in on the bug not by re-running the entire program, but
+by progressing back and forth through execution in the problem area.
 \par
@@
 \lstinline{Breakpoint 1 at 0x561c8fede303: file memo.c, line 52.}\\
-Users who did not send the command \lstinline{c} will see it displayed
-along with the response.
+Users who did not send the command \lstinline{break 52} will see it
+displayed along with the response.
 
@@
 backend primarily so that a Pod could be created dynamically for each
-debugging session.
+debugging session. \pagebreak
 
@@
 \lstinline{image} sourced from the container registry, and in the case
-of pods that need to interact with a load balancer, an
+of Pods that need to interact with a load balancer, an
 \lstinline{app}.
@@
 in the schema from the container registry and starts the Pod with any
-necessary commands.  The database Pod is connected to a long-term
+necessary commands.  The Database Pod is connected to a long-term
 storage volume at this time.  Upon successful creation, the Pod is
@@
 connect through the RR Message Load Balancer.  Clients and RR Debug
-Sessions send messages using the channel assigned to the pod at
+Sessions send messages using the channel assigned to the Pod at
 creation time.  A diagram of one full communication cycle between two
@@
 
-Current example pod startup scripts compile the program to be debugged
-and record execution when the pod is created.
+Current example Pod startup scripts compile the program to be debugged
+and record execution when the Pod is created.
 
@@
 experience to debugging locally in the future.  The system of
-inheritance used to create example debug session pods is designed to
+inheritance used to create example Debug Session Pods is designed to
 be as simple as possible, though a more fully featured frontend could
@@
 consists of handlers for the various API routes and instances of two
-classes which communicate with the database/Kubernetes API:
+classes which communicate with the database and/or Kubernetes API:
 \lstinline{TranslationPodManager} and \lstinline{UserManager}.  The
@@
 instantiates the Kubernetes Core V1 Python API to communicate with the
-cluster in order to create/delete pods.  This process of abstraction
+cluster in order to create/delete Pods.  This process of abstraction
 ensures that the database, underlying cluster structure, and
@@
 isolate dynamically and statically created Pods with minimal hassle,
-since Kubernetes' DNS resolution can connect pods across namespaces.
+since Kubernetes' DNS resolution can connect Pods across namespaces.
 If Pod creation fails, an error will be thrown.  Otherwise, a loop is
@@
 wish to debug.  Care has been taken to make it modular and extensible
-within reason, and additional authentication or pod creation features
+within reason, and additional authentication or Pod creation features
 should be easy to add.
@@
 \section{Next Steps} \label{next}
-
-\subsection{Current Functionality} \label{current}
-
-The following section details the first few steps students take to
-explore the \lstinline{memoization} example using the collaborative
-debugger.  Students A and B begin by examining the call stack at its
-deepest point in the recursive nth Fibonacci number function.  The
-terminal interface to rr and the source code view have been
-color-inverted to improve readability in this paper.
-
-\begin{figure}[h!]
-
-  \includegraphics[width=\textwidth]{session1i}
-  \centering
-  \caption{Creating a new Memoization Example Session}
-  \label{session1i}
-\end{figure}
-
-To start a new session, Student A selects ``memoization'' from the
-list of example sessions.\pagebreak
-
-\begin{figure}[h!]
-
-  \includegraphics[width=\textwidth]{session2i}
-  \centering
-  \caption{Joining the Memoization Example Session}
-  \label{session2i}
-\end{figure}
-
-In order to join Student A's session, Student B uses session code
-``71825''.  Meanwhile, Student A is presented with the the main view
-of the collaborative debugger.  On the right is a view of the source
-code for the \lstinline{memoization} example.  On the left is the
-terminal interface to rr.\pagebreak
-
-\begin{figure}[h!]
-
-  \includegraphics[width=\textwidth]{session3i}
-  \centering
-  \caption{Observing Output from the Memoization Example Session}
-  \label{session3i}
-\end{figure}
-
-Student B now types \lstinline{c} to continue execution.  Both
-students see the command and output: execution times and results for
-three different implementations of nth Fibonacci.  An advantage of
-rr's recording model is that these numbers will be the same every time
-the students run the program.\pagebreak
-
-\begin{figure}[h!]
-
-  \includegraphics[width=\textwidth]{session4i}
-  \centering
-  \caption{Setting a Breakpoint in the Memoization Example Session}
-  \label{session4i}
-\end{figure}
-
-Student A enters \lstinline{break 52} to set a breakpoint at line 52,
-the base case of \lstinline{fib_rec}.  Student B types \lstinline{run}
-and then \lstinline{c} to restart the recording and continue to the
-breakpoint.\pagebreak
-
-\begin{figure}[h!]
-
-  \includegraphics[width=\textwidth]{session5i}
-  \centering
-  \caption{Examining the Stack in the Memoization Example
-    Session}
-  \label{session5i}
-\end{figure}
-
-The students examine the stack resulting from a call of
-\lstinline{fib_rec(30)}.  Students can inspect different areas of the
-program and output while collaborating---Student A can look at
-\lstinline{fib_rec}, while Student B scrolls up for a preview of the
-memoized implementation.
-\par
-
-The students will continue to inspect the stack for other functions,
-exploring the reasons for the vast performance disparity between
-\lstinline{fib_memo} and \lstinline{fib_rec}.  Students A and B can
-not only help each other to learn about the stack and memoization, but
-can also teach one another how to use a debugger.  Because the
-frontend produces near-identical output to rr, students can directly
-translate their knowledge to it or gdb in the future.  Paradigms, such
-as examining the stack to understand execution, can be expanded to all
-future debugging.
 
@@
 simple.  Adding visualization features will bring the collaborative
-debugger much closer towards hopefully fulfilling it's second
+debugger much closer towards hopefully fulfilling its second
 requirement.
@@
 \cite{10.1145/1145287.1145293}, and \cite{10.1145/3361721.3361724}
-could be done to test both these aspects.  Students participating in
-an introductory systems programming class could be split into three
+could be done to test both of these aspects.  Students participating
+in an introductory systems programming class could be split into three
 groups.  Group one could study debugging individually, using rr by
@@
 The collaborative debugger aims to encourage collaboration and build
-students knowledge of debugging through careful design.  Many
+students' knowledge of debugging through careful design.  Many
 components of a distributed system come together to create a platform
